### 2019-12-02 blackholes/pavt/article.tex
Article skeleton

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
\usepackage{marvosym}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\usepackage{graphicx}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{hyperref}
\def\UrlFont{\rmfamily}
\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}

\def\orcidID#1{\unskip$^{[#1]}$}
\def\letter{$^{\textrm{(\Letter)}}$}

\begin{document}
\mainmatter              % start of a contribution
%
\title{Topological approach to finding blackholes in directed networks}
%
\titlerunning{Topological black holes mining}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Denis Ivanov\inst{1} \and Alexander Semenov\inst{2}}
%
\authorrunning{Ivanov, Semenov} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Denis Ivanov and Alexander Semenov}
%
\institute{Lomonosov Moscow State University, Moscow, Russian Federation \\
\email{mr.salixnew@gmail.com}
\and
JSC NICEVT, Moscow, Russian Federation\\
\email{semenov@nicevt.ru}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
In this paper we consider the task of finding so called Blackhole pattern in directed unweighed graphs.
Firstly, we analyse already existing algorithms and approaches. Secondly, we describe how structure of a graph
affects their efficiency. Then we introduce our approach to graph preprocessing. Finally, we describe topological sort based heuristic
and provide results of experimental comparison between our approach and previously developed algorithm.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{Directed networks $\cdot$ Subgraph mining}
\end{abstract}
%
\section{Blackhole search in directed graph}
%

%
\subsection{Task statement}
%

%
\subsubsection{Subsection example}
%

%
\subsection{Related work}
%

%
\subsection{Known issues}
%

%
\section{Graph perprocessing}
%

%
\section{Graph topology}
%

%
\section{Algorithm design}
%

%
\section{Experimental part}
%

%
\section{Conclusion}
%

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\begin{thebibliography}{6}
%

\bibitem {a1:a2}
Example
\doi{doi:number}

\bibitem {onlyurl}
NCBI: National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}
\end{document}

### 2019-12-02 blackholes/pavt/article.tex
Algorithm design in general

@@
 %
-\section{Graph perprocessing}
+\section{Algorithm design}
+We propose to look at the black hole search task from the following perspective: it would be divided into two independent 
+tasks. 
+The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to
+restricted time span. Decreased graph scale would definitely speed up the rest of the calculation, as it is result in brute force search in most cases.
+The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in restricted time.
 %
@@
 %
-\section{Graph topology}
+\subsection{Graph perprocessing}
 %
@@
 %
-\section{Algorithm design}
+\subsection{Graph topology}
 %

### 2019-12-02 blackholes/pavt/article.tex
Biblography is added
Graph preprocessing

@@
 %
-\subsection{Graph perprocessing}
+\subsection{Graph preprocessing}
+As it was described in the Known issues section, large SCCs waste lots of computational time without adding any new blackholes.
+For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's vertices. 
+In such a case, it would be nice to consider large SCC as a single vertex, aggregating all incoming and outcoming peripheral edges
+of original SCC.
+On the preprocessing stage of algrithm, we reduce every SCC to a single vertex. The trick is not new and known as graph condensation.
+More formally:\par
+\begin{definition}
+TODO: Graph condensation definition
+\end{definition}
+
+In this article we use to use the following algorithm for finding graph condensation.\par
+
+First, we find topological sort of the initial graph. It takes O(E + V).
+\begin{definition}
+TODO: Topolorgical sort
+\end{definition}
+
+\begin{definition}
+TODO: Closure
+\end{definition}
+
+\begin{definition}
+TODO: Reachable
+\end{definition}
+
+When we finally have an array, containing graph vertices in the topologically sorted order, we do the following:
+
+Define global array of used vertices. Initially, all vertices are unused.
+Then for every vertex from bottom to top:
+\begin{enumerate}
+    \item Acquire closure of unused vertices reachable from current vertex; 
+    \item Mark every vertex in the closure as used;
+    \item All the acquired vertices belong to the same SCC.
+\end{enumerate}
+
+The fact given in the last step can be shortly explained in the following way:
+topological sort guarantees that any vertex has index larger than indeces of her children.
+That is why, given that we process vertices from bottom to top, by the moment we process the vertex, all it's children are already marked used and no longer available for search.
+As all the ways down are prohibited and there's no way up, we will acquire an SCC.\par
+
+As a result of the algorithm, described above, we have each vertex associated with certain SCC.
+SCCs itself form a graph with number of vertices less than or equal to one of the initial graph.
+
+Finally, we need to remove edges duplicates and the graph condesation is built.
+
 %

### 2019-12-02 blackholes/pavt/article.tex
Blackholw search

@@
 \subsection{Related work}
-%
+\cite{li2010detecting,li2012mining,li2014mining,hong2015detecting}
 
@@
 
-In this article we use to use the following algorithm for finding graph condensation.\par
-
+In this article we use the following algorithm for finding graph condensation.
+It was independently proposed by Kosaraju and Sharir back in 1979.\par 
 First, we find topological sort of the initial graph. It takes O(E + V).
 \begin{definition}
-TODO: Topolorgical sort
+TODO: Topological sort
 \end{definition}
@@
 \begin{definition}
-TODO: Closure
+Closure of vertex v is a set of all nodes reachable from v, including v.
 \end{definition}
@@
 %
-\subsection{Graph topology}
+\subsection{Blackholes search}
+When the graph is prepared, we can move on to searching blackholes. The task is combinatorial in its nature, therefore, we aim to
+decrease the number of potential black holes.
+To understand the issue, we first describe BruteForce approach. \par
+
+\begin{definition}
+TODO: simple blackhole
+\end{definition}
+
+\begin{definition}
+TODO: root of blackhole
+\end{definition}
+
+\begin{definition}
+TODO: basis of blackhole
+\end{definition}
+
+\begin{definition}
+TODO: complex blackhole
+\end{definition}
+
+\begin{definition}
+TODO:
+\end{definition}
+
+To BruteForce the task we iterate over all possible unordered sets of nodes. 
+Then for every node we acquire its closure. Union of all such closures is a candidate blackhole.
+If candidate blackhole is a weakly connected subgraph, then it is a blackhole.
+
+With BruteForce algorithm it is possible to find the same blackhole several times.
+We aim to avoid at least some of the duplicates and introduce heuristic for it.
+
+According to the definitions above if some node is not a root of blackhole, it can be omitted, as set of roots
+uniquely identify a blackhole.
+Every non-root node in blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for
+the given graph, we can skip bloated non-basis combinations of nodes.
+Of course, we could straightforward build the reachability matrix. But, it takes O(V \^ 3) time. Even though we decreased graph size
+in preprocessing step, it still can result in lots of computation. That is why, we will only try to partially skip duplicates and the rest will
+be filtered later.\par
+
+Let's consider a closure of a single blackhole root. Let this closure to be acquired by depth-first search, which, basically means, that
+we can lay those nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
+in the given roots closure. Finally, if some node lays on the i'th position on the topological sort and its closure size is i, than we can
+be conclude, that every node with index less than i in the topological sort will be reachable from v(i).
+
 %
@@
 \bibliographystyle{spmpsci}
-\begin{thebibliography}{6}
-%
+\bibliography{biblio}
 
-\bibitem {a1:a2}
-Example
-\doi{doi:number}
-
-\bibitem {onlyurl}
-NCBI: National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}
-
-\end{thebibliography}
 \end{document}

### 2019-12-02 blackholes/pavt/article.tex
Minor modification

--- 2019-12-02 blackholes/pavt/article.tex
@@
 \def\letter{$^{\textrm{(\Letter)}}$}
+
+%%%%%% DEBUG section %%%%%%
+\newcommand*{\DEBUG}{}
+
+\ifdefined\DEBUG
+\usepackage[T2A]{fontenc}
+\usepackage[utf8]{inputenc}
+%\usepackage[russian]{babel}
+\usepackage[usenames]{color}
+%\usepackage{colortbl}
+\newcommand{\FIXME}[1]{ % описание
+	\colorbox{yellow}{#1}
+}
+\else
+\newcommand{\FIXME}[1]{ % описание
+}
+\fi
 
@@
 %
-\subsection{Related work}
+\subsection{Related Work}
+\cite{li2010detecting,li2012mining,li2014mining,hong2015detecting}
+
+%
+\subsection{Known Issues}
 %
@@
 %
-\subsection{Known issues}
+\section{Algorithm Design}
+We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
+The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
+Decreased graph scale would definitely speed up the the calculation, as it can fall into a brute-force search in most cases.
+The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in a restricted time.
 %
@@
 %
-\section{Algorithm design}
-We propose to look at the black hole search task from the following perspective: it would be divided into two independent 
-tasks. 
-The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to
-restricted time span. Decreased graph scale would definitely speed up the rest of the calculation, as it is result in brute force search in most cases.
-The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in restricted time.
-%
+\subsection{Graph Preprocessing}
+\FIXME{не забыть описать в KnownIssues}
 
-%
-\subsection{Graph preprocessing}
-As it was described in the Known issues section, large SCCs waste lots of computational time without adding any new blackholes.
+\FIXME{не забыть вставить ссылки на RMAT \cite{chakrabarti2004r}, SSCA2 \cite{bader2005design}, \cite{random-uniform}, SmallWorld \cite{watts1999networks}}
+
+\FIXME{не понял логику абзаца: большая SCC может жрать много времени, граф может из нее состоять, и лучше сжать ее до 1 вершины}
+
+\FIXME{Хорошо бы ссылку на graph condensation}
+
+As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
 For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's vertices. 
@@
 of original SCC.
-On the preprocessing stage of algrithm, we reduce every SCC to a single vertex. The trick is not new and known as graph condensation.
+On the preprocessing stage of the algorithm we reduce every SCC to a single vertex. The algorithm is not new and known as a graph condensation.
+
 More formally:\par
@@
 \bibliographystyle{spmpsci}
-\begin{thebibliography}{6}
-%
+\bibliography{biblio}
 
-\bibitem {a1:a2}
-Example
-\doi{doi:number}
+%\begin{thebibliography}{6}
+%\bibitem {a1:a2}
+%Example
+%\doi{doi:number}
 
-\bibitem {onlyurl}
-NCBI: National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}
+%\bibitem {onlyurl}
+%\NCBI: National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}
 
-\end{thebibliography}
+%\end{thebibliography}
 \end{document}

### 2019-12-02 blackholes/pavt/article.tex

--- 2019-12-02 blackholes/pavt/article.tex
@@
 %
-\subsection{Related work}
+\subsection{Related Work}
 \cite{li2010detecting,li2012mining,li2014mining,hong2015detecting}
@@
 %
-\subsection{Known issues}
+\subsection{Known Issues}
 %
@@
 %
-\section{Algorithm design}
-We propose to look at the black hole search task from the following perspective: it would be divided into two independent 
-tasks. 
-The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to
-restricted time span. Decreased graph scale would definitely speed up the rest of the calculation, as it is result in brute force search in most cases.
-The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in restricted time.
+\section{Algorithm Design}
+We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
+The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
+Decreased graph scale would definitely speed up the the calculation, as it can fall into a brute-force search in most cases.
+The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in a restricted time.
 %
@@
 %
-\subsection{Graph preprocessing}
-As it was described in the Known issues section, large SCCs waste lots of computational time without adding any new blackholes.
+\subsection{Graph Preprocessing}
+\FIXME{не забыть описать в KnownIssues}
+
+\FIXME{не забыть вставить ссылки на RMAT \cite{chakrabarti2004r}, SSCA2 \cite{bader2005design}, \cite{random-uniform}, SmallWorld \cite{watts1999networks}}
+
+\FIXME{не понял логику абзаца: большая SCC может жрать много времени, граф может из нее состоять, и лучше сжать ее до 1 вершины}
+
+\FIXME{Хорошо бы ссылку на graph condensation}
+
+As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
 For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's vertices. 

### 2019-12-02 blackholes/pavt/article.tex
Minor modification
Minor modifications

@@
 \section{Blackhole search in directed graph}
-%
+applications Anti-Money Laundering \cite{semenov2017survey}
 
@@
 
-\FIXME{Хорошо бы ссылку на graph condensation}
-
 As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
-For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's vertices. 
-In such a case, it would be nice to consider large SCC as a single vertex, aggregating all incoming and outcoming peripheral edges
+For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
+In such a case, it would be nice to consider large SCC as a single node, aggregating all incoming and outcoming peripheral edges
 of original SCC.
-On the preprocessing stage of the algorithm we reduce every SCC to a single vertex. The algorithm is not new and known as a graph condensation.
+On the preprocessing stage of the algorithm we reduce every SCC to a single node. The algorithm is not new and known as a graph condensation.
 
@@
 In this article we use the following algorithm for finding graph condensation.
-It was independently proposed by Kosaraju and Sharir back in 1979.\par 
+It was independently proposed by Kosaraju and Sharir back in 1979. \FIXME{Ссылку на эту статью} \par
+
 First, we find topological sort of the initial graph. It takes O(E + V).
@@
 \begin{definition}
-Closure of vertex v is a set of all nodes reachable from v, including v.
+Closure of node v is a set of all nodes reachable from v, including v.
 \end{definition}
@@
 
-When we finally have an array, containing graph vertices in the topologically sorted order, we do the following:
+When we finally have an array, containing graph nodes in the topologically sorted order, we do the following:
 
-Define global array of used vertices. Initially, all vertices are unused.
-Then for every vertex from bottom to top:
+Define global array of used nodes. Initially, all nodes are unused.
+Then for every node from bottom to top:
 \begin{enumerate}
-    \item Acquire closure of unused vertices reachable from current vertex; 
-    \item Mark every vertex in the closure as used;
-    \item All the acquired vertices belong to the same SCC.
+    \item Acquire closure of unused nodes reachable from current node; 
+    \item Mark every node in the closure as used;
+    \item All the acquired nodes belong to the same SCC.
 \end{enumerate}
@@
 The fact given in the last step can be shortly explained in the following way:
-topological sort guarantees that any vertex has index larger than indeces of her children.
-That is why, given that we process vertices from bottom to top, by the moment we process the vertex, all it's children are already marked used and no longer available for search.
+topological sort guarantees that any node has index larger than indeces of her children.
+That is why, given that we process nodes from bottom to top, by the moment we process the node, all it's children are already marked used and no longer available for search.
 As all the ways down are prohibited and there's no way up, we will acquire an SCC.\par
 
-As a result of the algorithm, described above, we have each vertex associated with certain SCC.
-SCCs itself form a graph with number of vertices less than or equal to one of the initial graph.
+As a result of the algorithm, described above, we have each node associated with certain SCC.
+SCCs itself form a graph with number of nodes less than or equal to one of the initial graph.
 
@@
 \subsection{Blackholes search}
-When the graph is prepared, we can move on to searching blackholes. The task is combinatorial in its nature, therefore, we aim to
-decrease the number of potential black holes.
-To understand the issue, we first describe BruteForce approach. \par
+When the graph is prepared, we can move on to searching blackholes. The problem is combinatorial in its nature, therefore, we aim to
+decrease the number of potential blackholes.
+To understand the issue, we first describe a brute-force approach.
 
@@
 
-To BruteForce the task we iterate over all possible unordered sets of nodes. 
-Then for every node we acquire its closure. Union of all such closures is a candidate blackhole.
-If candidate blackhole is a weakly connected subgraph, then it is a blackhole.
+%TODO: осознать
+In the BruteForce algorithm we iterate over all possible unordered sets of nodes. 
+Then for every node we acquire its closure. Union of all such closures is a potential blackhole.
+If the candidate blackhole is a weakly connected subgraph, then it is a blackhole.
 
-With BruteForce algorithm it is possible to find the same blackhole several times.
+With the BruteForce algorithm it is possible to find the same blackhole several times.
 We aim to avoid at least some of the duplicates and introduce heuristic for it.
 
-According to the definitions above if some node is not a root of blackhole, it can be omitted, as set of roots
-uniquely identify a blackhole.
-Every non-root node in blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for
-the given graph, we can skip bloated non-basis combinations of nodes.
+According to the definitions above if some node is not a root of a blackhole, it can be omitted, as set of roots uniquely identify a blackhole.
+Every non-root node in a blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.
 Of course, we could straightforward build the reachability matrix. But, it takes O(V \^ 3) time. Even though we decreased graph size

### 2019-12-02 blackholes/pavt/article.tex
Results and experimental part

--- 2019-12-02 blackholes/pavt/article.tex
@@
 
-\FIXME{не понял логику абзаца: большая SCC может жрать много времени, граф может из нее состоять, и лучше сжать ее до 1 вершины}
-
-\FIXME{Хорошо бы ссылку на graph condensation}
-
 As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
@@
 we can lay those nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
-in the given roots closure. Finally, if some node lays on the i'th position on the topological sort and its closure size is i, than we can
-be conclude, that every node with index less than i in the topological sort will be reachable from v(i).
+in the given roots closure. Finally, if some node lays on the i'th position on the topological sort and its closure size is i, then we can 
+conclude, that every node with index less than i in the topological sort will be reachable from v(i).
 
@@
 \section{Experimental part}
+To demonstrate efficiency of our approach, we conducted a series of experiments. We ran both algorithms (New Jersey team and ours) on a 
+set of graphs. \par
+For these experiments we chose RMAT, Uniform Random (UR) and SSCA2 graph types. Those were generated for scales from 4 to 22
+with step 2. Size i means that graph has 2 pow i nodes and 32 * V edges. Actual count of edges to process can differ from these ideal values, as duplicate edges 
+and self-loops are omitted when reading graph from file. \par
+
+\begin{table}[]
+\caption{Configurations of experimental graphs}
+\label{tabular:tableconfigurations}
+\begin{center}
+\begin{tabular}{ccc}
+Graph Type & Nodes Count & Edges Count \\
+RMAT.04 & 16 & 177 \\
+RMAT.06 & 64 & 1270 \\
+RMAT.08 & 256 & 6663 \\
+RMAT.10 & 1024 & 30186 \\
+RMAT.12 & 4096 & 126988 \\
+RMAT.14 & 16384 & 517554 \\
+RMAT.16 & 65536 & 2087072 \\
+RMAT.18 & 262144 & 7892410 \\
+RMAT.20 & 1048576 & 32993233 \\
+RMAT.22 & 4194304 & 67099679 \\
+SSCA2.04 & 16 & 120 \\
+SSCA2.06 & 64 & 722 \\
+SSCA2.08 & 256 & 2668 \\
+SSCA2.10 & 1024 & 10529 \\
+SSCA2.12 & 4096 & 43603 \\
+SSCA2.14 & 16384 & 173477 \\
+SSCA2.16 & 65536 & 677649 \\
+SSCA2.18 & 262144 & 2724762 \\
+SSCA2.20 & 1048576 & 10902287 \\
+SSCA2.22 & 4194304 & 43580024 \\
+UR.04 & 16 & 207 \\
+UR.06 & 64 & 1614 \\
+UR.08 & 256 & 7652 \\
+UR.10 & 1024 & 32201 \\
+UR.12 & 4096 & 130486 \\
+UR.14 & 16384 & 523740 \\
+UR.16 & 65536 & 2096570 \\
+UR.18 & 262144 & 8388074 \\
+UR.20 & 1048576 & 33553908 \\
+UR.22 & 4194304 & 134217174
+\end{tabular}
+\end{center}
+\end{table}
+
+We've been conducting our experiments on a Linux machine with the following specifications:
+\begin{itemize}
+    \item OS: Ubuntu 16.04 xenial
+    \item Kernel: x86\_64 Linux 4.15.0-46-generic
+    \item Shell: bash 4.3.48
+    \item CPU: Intel Core i7-8550U CPU @ 4GHz
+    \item RAM: 15802MiB
+\end{itemize}
+
+All runs were performed in a single-thread mode. We allowed up to 20 minutes of working time for every run. In order to save time,
+algorithms did not print out found blackholes, but printed counter on every found and not filtered (by any reason) blackhole.
+
+\begin{table}[H]
+\caption{Experimental results}
+\label{tabular:tableresults}
+\begin{center}
+\begin{tabular}{ccccccccc}
+Graph & Type & Algorithm & Blackholes & Time & Algorithm & Blackholes & Time \\
+RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
+RMAT & 06 & NewJersey & 1 & 0.0121347 & TopSort & 1 & 0.00034812 \\
+RMAT & 08 & NewJersey & 1 & 1.66685 & TopSort & 1 & 0.00146486 \\
+RMAT & 10 & NewJersey & 6 & 1200 & TopSort & 10 & 0.00684045 \\
+RMAT & 12 & NewJersey & 29 & 1200 & TopSort & 971738 & 1200 \\
+RMAT & 14 & NewJersey & 187 & 1200 & TopSort & 735128 & 1200 \\
+RMAT & 16 & NewJersey & 1192 & 1200 & TopSort & 373705 & 1200 \\
+RMAT & 18 & NewJersey & 5176 & 1200 & TopSort & 10286 & 1200 \\
+RMAT & 20 & NewJersey & 617 & 1200 & TopSort & 48867 & 1200 \\
+RMAT & 22 & NewJersey & 98 & 1200 & TopSort & 0 & 1200 \\
+SSCA2 & 04 & NewJersey & 16 & 0.154763 & TopSort & 16 & 0.000189925 \\
+SSCA2 & 06 & NewJersey & 11 & 1200 & TopSort & 70 & 0.00121528 \\
+SSCA2 & 08 & NewJersey & 42 & 1200 & TopSort & 263314 & 1200 \\
+SSCA2 & 10 & NewJersey & 85 & 1200 & TopSort & 114411 & 1200 \\
+SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
+SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
+SSCA2 & 16 & NewJersey & 5829 & 1200 & TopSort & 893376 & 1200 \\
+SSCA2 & 18 & NewJersey & 23179 & 1200 & TopSort & 224567 & 1200 \\
+SSCA2 & 20 & NewJersey & 92347 & 1200 & TopSort & 24199 & 1200 \\
+SSCA2 & 22 & NewJersey & 370770 & 1200 & TopSort & 0 & 1200 \\
+UR & 04 & NewJersey & 1 & 0.000223768 & TopSort & 1 & 0.000115029 \\
+UR & 06 & NewJersey & 1 & 0.0170713 & TopSort & 1 & 0.000331235 \\
+UR & 08 & NewJersey & 1 & 2.30256 & TopSort & 1 & 0.00147483 \\
+UR & 10 & NewJersey & 1 & 240.822 & TopSort & 1 & 0.00670724 \\
+UR & 12 & NewJersey & 0 & 1200 & TopSort & 1 & 0.034011 \\
+UR & 14 & NewJersey & 0 & 1200 & TopSort & 1 & 0.47257 \\
+UR & 16 & NewJersey & 0 & 1200 & TopSort & 1 & 3.05003 \\
+UR & 18 & NewJersey & 0 & 1200 & TopSort & 1 & 14.5788 \\
+UR & 20 & NewJersey & 0 & 1200 & TopSort & 1 & 67.7176 \\
+UR & 22 & NewJersey & 0 & 1200 & TopSort & 1 & 373.244
+\end{tabular}
+\end{center}
+\end{table}
+
 %

### 2019-12-02 blackholes/pavt/article.tex

@@
 As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
-For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's vertices. 
-In such a case, it would be nice to consider large SCC as a single vertex, aggregating all incoming and outcoming peripheral edges
+For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
+In such a case, it would be nice to consider large SCC as a single node, aggregating all incoming and outcoming peripheral edges
 of original SCC.
-On the preprocessing stage of the algorithm we reduce every SCC to a single vertex. The algorithm is not new and known as a graph condensation.
+On the preprocessing stage of the algorithm we reduce every SCC to a single node. The algorithm is not new and known as a graph condensation.
 

### 2019-12-03 blackholes/pavt/article.tex
Fix build
Algo example is added, tables are edited

--- 2019-12-02 blackholes/pavt/article.tex
@@
 \usepackage{graphicx}
+\usepackage{float}
+\usepackage[boxed,vlined,ruled]{algorithm2e}
 
@@
 
-\FIXME{не забыть вставить ссылки на RMAT \cite{chakrabarti2004r}, SSCA2 \cite{bader2005design}, \cite{random-uniform}, SmallWorld \cite{watts1999networks}}
-
-\FIXME{не понял логику абзаца: большая SCC может жрать много времени, граф может из нее состоять, и лучше сжать ее до 1 вершины}
+\FIXME{не забыть вставить ссылки SmallWorld \cite{watts1999networks}}
 
@@
 That is why, given that we process nodes from bottom to top, by the moment we process the node, all it's children are already marked used and no longer available for search.
-As all the ways down are prohibited and there's no way up, we will acquire an SCC.\par
+As all the ways down are prohibited and there's no way up, we will acquire an SCC.
 
@@
 %
-\subsection{Blackholes search}
+\subsection{Blackholes Search}
 When the graph is prepared, we can move on to searching blackholes. The problem is combinatorial in its nature, therefore, we aim to
@@
 
+\begin{figure}[H]
+	\begin{center}
+		\begin{algorithm}[H]
+			\SetAlgoLined
+			\SetKwInOut{Input}{Input}
+			\SetKwInOut{Output}{Output}
+			\Input{$CDG(C, E)$ -- channel dependency graph for a system topology with 
+				direction order routing, \\
+				min\_dir -- directions, which are less than this will not be considered}
+			\Output{$CDG$ with added edges}
+			// build the used direction sets\\ 
+			\ForEach{$v \in C$}{
+				$B(v) = \{s_i: \alpha_{u_i, s_i} \in BFS(CDG(C, E), v)\}$
+			}
+			\ForEach{$\alpha_{v_1, s_1}, \alpha_{v_2, s_2} \in C: v_2 = v_1 + s_1, (\alpha_{v_1, s_1}, \alpha_{v_2, s_2}) \notin E, s_1 \ge min\_dir$}{
+				\If{$s_1 \notin B(\alpha_{v_2,s_2})$}{
+					add $(\alpha_{v_1, s_1}, \alpha_{v_2, s_2})$ to $E$\\
+					// rebuild used direction set for $\alpha_{v_1, s_1}$\\
+					\ForEach{$v \in BFS(CDG(C, \bar{E}), \alpha_{v_1, s_1})$}{
+						$B(v) = B(v) \cup B(\alpha_{v_2, s_2})$
+					}
+				}
+			}
+			\label{alg:add_additional_edge}
+			\caption{Adding edges to CDG.}
+		\end{algorithm}
+	\end{center}
+\end{figure}
+
 %TODO: осознать
@@
 Every non-root node in a blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.
-Of course, we could straightforward build the reachability matrix. But, it takes O(V \^ 3) time. Even though we decreased graph size
-in preprocessing step, it still can result in lots of computation. That is why, we will only try to partially skip duplicates and the rest will
-be filtered later.\par
+Of course, we could straightforward build the reachability matrix.
+But it takes $O(V^3)$ time.
+\FIXME{определить $G=(V, E)$} Even though we decreased graph size in preprocessing step, it still can result in lots of computation.
+That is why we will only try to partially skip duplicates and the rest will be filtered later.
 
-Let's consider a closure of a single blackhole root. Let this closure to be acquired by depth-first search, which, basically means, that
-we can lay those nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
-in the given roots closure. Finally, if some node lays on the i'th position on the topological sort and its closure size is i, than we can
-be conclude, that every node with index less than i in the topological sort will be reachable from v(i).
+Let's consider a closure of a single blackhole root. Let this closure to be acquired by depth-first search, which basically means, that
+we can iterate over these nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
+in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
+conclude, that every node with index less than $i$ in the topological sort will be reachable from $v[i]$. \FIXME{определить v[i]}
+
+%
+\section{Experimental Results}
+To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey and TopSort) on a 
+set of graphs.
+
+For these experiments we chose Uniform Random (UR) \cite{random-uniform}, RMAT \cite{chakrabarti2004r},  and SSCA2 \cite{bader2005design} graph types. Those were generated for scales from 4 to 22
+with step 2.
+Size $i$ means that graph has $2^i$ nodes and $32*V$ edges.
+Actual count of edges to process can differ from these ideal values, as duplicate edges and self-loops are omitted when reading graph from file.
+Configuration of the considered graphs is presented in Table \ref{tabular:graphs}.
+
+\begin{table}[]
+\caption{Configuration of experimental graphs}
+\label{tabular:graphs}
+\begin{center}
+\begin{tabular}{c|c|c}
+Graph Type & Nodes Count & Edges Count \\
+\hline
+RMAT.04 & 16 & 177 \\
+RMAT.06 & 64 & 1270 \\
+RMAT.08 & 256 & 6663 \\
+RMAT.10 & 1024 & 30186 \\
+RMAT.12 & 4096 & 126988 \\
+RMAT.14 & 16384 & 517554 \\
+RMAT.16 & 65536 & 2087072 \\
+RMAT.18 & 262144 & 7892410 \\
+RMAT.20 & 1048576 & 32993233 \\
+RMAT.22 & 4194304 & 67099679 \\
+SSCA2.04 & 16 & 120 \\
+SSCA2.06 & 64 & 722 \\
+SSCA2.08 & 256 & 2668 \\
+SSCA2.10 & 1024 & 10529 \\
+SSCA2.12 & 4096 & 43603 \\
+SSCA2.14 & 16384 & 173477 \\
+SSCA2.16 & 65536 & 677649 \\
+SSCA2.18 & 262144 & 2724762 \\
+SSCA2.20 & 1048576 & 10902287 \\
+SSCA2.22 & 4194304 & 43580024 \\
+UR.04 & 16 & 207 \\
+UR.06 & 64 & 1614 \\
+UR.08 & 256 & 7652 \\
+UR.10 & 1024 & 32201 \\
+UR.12 & 4096 & 130486 \\
+UR.14 & 16384 & 523740 \\
+UR.16 & 65536 & 2096570 \\
+UR.18 & 262144 & 8388074 \\
+UR.20 & 1048576 & 33553908 \\
+UR.22 & 4194304 & 134217174
+\end{tabular}
+\end{center}
+\end{table}
+
+We have been conducting our experiments on a Linux machine with the following specifications:
+\begin{itemize}
+    \item CPU: Intel Core i7-8550U CPU @ 4GHz
+    \item RAM: 15802 MiB
+    \item OS: Ubuntu 16.04 xenial
+    \item Kernel: x86\_64 Linux 4.15.0-46-generic
+    \item Shell: bash 4.3.48
+\end{itemize}
+
+All runs were performed in a single-thread mode. We allowed up to 20 minutes of working time for every run. In order to save time,
+algorithms did not print out found blackholes, but printed counter on every found and not filtered (by any reason) blackhole.
+
+\begin{table}[]
+\caption{Experimental results}
+\label{tabular:tableresults}
+\begin{center}
+\begin{tabular}{c|c|c|c|c|c|c|c|c}
+Graph & Type & Algorithm & Blackholes & Time & Algorithm & Blackholes & Time \\
+\hline
+RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
+RMAT & 06 & NewJersey & 1 & 0.0121347 & TopSort & 1 & 0.00034812 \\
+RMAT & 08 & NewJersey & 1 & 1.66685 & TopSort & 1 & 0.00146486 \\
+RMAT & 10 & NewJersey & 6 & 1200 & TopSort & 10 & 0.00684045 \\
+RMAT & 12 & NewJersey & 29 & 1200 & TopSort & 971738 & 1200 \\
+RMAT & 14 & NewJersey & 187 & 1200 & TopSort & 735128 & 1200 \\
+RMAT & 16 & NewJersey & 1192 & 1200 & TopSort & 373705 & 1200 \\
+RMAT & 18 & NewJersey & 5176 & 1200 & TopSort & 10286 & 1200 \\
+RMAT & 20 & NewJersey & 617 & 1200 & TopSort & 48867 & 1200 \\
+RMAT & 22 & NewJersey & 98 & 1200 & TopSort & 0 & 1200 \\
+SSCA2 & 04 & NewJersey & 16 & 0.154763 & TopSort & 16 & 0.000189925 \\
+SSCA2 & 06 & NewJersey & 11 & 1200 & TopSort & 70 & 0.00121528 \\
+SSCA2 & 08 & NewJersey & 42 & 1200 & TopSort & 263314 & 1200 \\
+SSCA2 & 10 & NewJersey & 85 & 1200 & TopSort & 114411 & 1200 \\
+SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
+SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
+SSCA2 & 16 & NewJersey & 5829 & 1200 & TopSort & 893376 & 1200 \\
+SSCA2 & 18 & NewJersey & 23179 & 1200 & TopSort & 224567 & 1200 \\
+SSCA2 & 20 & NewJersey & 92347 & 1200 & TopSort & 24199 & 1200 \\
+SSCA2 & 22 & NewJersey & 370770 & 1200 & TopSort & 0 & 1200 \\
+UR & 04 & NewJersey & 1 & 0.000223768 & TopSort & 1 & 0.000115029 \\
+UR & 06 & NewJersey & 1 & 0.0170713 & TopSort & 1 & 0.000331235 \\
+UR & 08 & NewJersey & 1 & 2.30256 & TopSort & 1 & 0.00147483 \\
+UR & 10 & NewJersey & 1 & 240.822 & TopSort & 1 & 0.00670724 \\
+UR & 12 & NewJersey & 0 & 1200 & TopSort & 1 & 0.034011 \\
+UR & 14 & NewJersey & 0 & 1200 & TopSort & 1 & 0.47257 \\
+UR & 16 & NewJersey & 0 & 1200 & TopSort & 1 & 3.05003 \\
+UR & 18 & NewJersey & 0 & 1200 & TopSort & 1 & 14.5788 \\
+UR & 20 & NewJersey & 0 & 1200 & TopSort & 1 & 67.7176 \\
+UR & 22 & NewJersey & 0 & 1200 & TopSort & 1 & 373.244
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 %
-\section{Experimental part}
-%
+\section{Conclusion}
 
-%
-\section{Conclusion}
-%
+\textit{Comment for the reviewer}. In the final paper revision we add parallel implementation of the TopSort algorithm and present much detailed evaluation on the larger graphs.
 

### 2019-12-03 blackholes/pavt/article.tex
Changes

--- 2019-12-02 blackholes/pavt/article.tex
@@
 
-More formally:\par
+More formally:
+
 \begin{definition}
@@
 In this article we use the following algorithm for finding graph condensation.
-It was independently proposed by Kosaraju and Sharir back in 1979. \FIXME{Ссылку на эту статью} \par
+It was independently proposed by Kosaraju and Sharir back in 1979. \FIXME{Ссылку на эту статью}
 
@@
 Of course, we could straightforward build the reachability matrix. But, it takes O(V \^ 3) time. Even though we decreased graph size
-in preprocessing step, it still can result in lots of computation. That is why, we will only try to partially skip duplicates and the rest will
-be filtered later.\par
+in preprocessing step, it still can result in lots of computation. It's worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
+Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
+be filtered later in lazy manner.
 
@@
 in the given roots closure. Finally, if some node lays on the i'th position on the topological sort and its closure size is i, then we can 
-conclude, that every node with index less than i in the topological sort will be reachable from v(i).
+conclude, that every node with index less than i in the topological sort will be reachable from topsort[i].
+
+TopSortFilter algorithm:
+\begin{enumerated}
+    \item For every node 
+    \item Calculate topological sort order
+    \item Calculate closure size for every vertex
+    \item For every node
 
@@
 To demonstrate efficiency of our approach, we conducted a series of experiments. We ran both algorithms (New Jersey team and ours) on a 
-set of graphs. \par
+set of graphs.
+
+iBlackhole algorithm was implemented in accordance with the pseudocode in th original article. Divide and Conquer heuristic were not used for it.
+Our algorithm first applies condensation-based preprocessing and then uses BruteForce with partial reachability heuristic implemented via topsort.
+Both algorithms filter duplicates
+
 For these experiments we chose RMAT, Uniform Random (UR) and SSCA2 graph types. Those were generated for scales from 4 to 22
 with step 2. Size i means that graph has 2 pow i nodes and 32 * V edges. Actual count of edges to process can differ from these ideal values, as duplicate edges 
-and self-loops are omitted when reading graph from file. \par
+and self-loops are omitted when reading graph from file.
 
@@
 
-\begin{table}[H]
+Please, mention, that "blackhole found" means that we have a set of vertex numbers in memory ready to be printed out and it's already validated.
+
+\begin{table}[]
 \caption{Experimental results}
@@
 \begin{tabular}{ccccccccc}
-Graph & Type & Algorithm & Blackholes & Time & Algorithm & Blackholes & Time \\
+Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
 RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
@@
 SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}[]
+\caption{Experimental results. Part 2}
+\label{tabular:tableresults2}
+\begin{center}
+\begin{tabular}{ccccccccc}
+Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
 SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
@@
 
+\pagebreak
 %

### 2019-12-03 blackholes/pavt/article.tex

--- 2019-12-03 blackholes/pavt/article.tex
@@
 Every non-root node in a blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.
-Of course, we could straightforward build the reachability matrix.
-But it takes $O(V^3)$ time.
-\FIXME{определить $G=(V, E)$} Even though we decreased graph size in preprocessing step, it still can result in lots of computation.
-That is why we will only try to partially skip duplicates and the rest will be filtered later.
+
+\FIXME{определить $G=(V, E)$}
+\FIXME{определить v[i]}
+
+Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
+in preprocessing step, it still can result in lots of computation. It's worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
+Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
+be filtered later in lazy manner.
 
@@
 in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
-conclude, that every node with index less than $i$ in the topological sort will be reachable from $v[i]$. \FIXME{определить v[i]}
-
+conclude, that every node with index less than $i$ in the topological sort will be reachable from $v[i]$. 
+TopSortFilter algorithm:
+\begin{enumerate}
+    \item For every node 
+    \item Calculate topological sort order
+    \item Calculate closure size for every vertex
+    \item For every node
+\end{enumerate}
 %
@@
 set of graphs.
+
+iBlackhole algorithm was implemented in accordance with the pseudocode in the original article. Divide and Conquer heuristic were not used for it.
+Our algorithm first applies condensation-based preprocessing and then uses BruteForce with partial reachability heuristic implemented via topsort.
+Both algorithms filter duplicates
 
@@
 
+Please, mention, that "blackhole found" means that we have a set of vertex numbers in memory ready to be printed out and it's already validated.
+
 \begin{table}[]
@@
 \begin{tabular}{c|c|c|c|c|c|c|c|c}
-Graph & Type & Algorithm & Blackholes & Time & Algorithm & Blackholes & Time \\
+Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
 \hline
@@
 SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}[]
+\caption{Experimental results. Part 2}
+\label{tabular:tableresults2}
+\begin{center}
+\begin{tabular}{c|c|c|c|c|c|c|c|c}
+Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
 SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\

### 2019-12-03 blackholes/pavt/article.tex
heuristic description

@@
 in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
-conclude, that every node with index less than $i$ in the topological sort will be reachable from $v[i]$. 
+conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsort[i]$. We will call such nodes special.
+
+If there are two special nodes under the same root, union of their closures will be equal to the closure of topologically higher node.
+Hence, union of closures for any subset of special nodes under the same root will result into closure of the highest special node.
+
+Let's consider two nodes in the closure of global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
+special one should be topologically lower than non-special. Otherwise it would be one node basis.
+
+To sum up: if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.
+
+Imagine, we are given a graph and a set of nodes. What is the algorithm to decide whether these nodes form the basis of a blackhole? We shall decide if there's at least
+one non-basic node.
+First, we need to know what special nodes there are. To find out, we will take every global root of condensated graph and find topoological sort of its closure.
+Those vertices, which suffice the rule above, we will mark as "special". 
+This step can be precalculated.
+Now, we shall calculate how many special nodes for every root there is in the given set. If there is more than one special node for root - the set is to be called non-basis and
+omitted. Also we shall check if there are any non-special nodes lower than the special ones, and if those found - omit the node set.
+
+\begin{figure}[H]
+	\begin{center}
+		\begin{algorithm}[H]
+			\SetAlgoLined
+			\SetKwInOut{Input}{Input}
+			\SetKwInOut{Output}{Output}
+			\Input{$G(V,E)$ -- directed acyclic graph (Graph condensation)}
+			\Output{$Special$ -- }
+			// build the used direction sets\\ 
+			\ForEach{$v \in C$}{
+				$B(v) = \{s_i: \alpha_{u_i, s_i} \in BFS(CDG(C, E), v)\}$
+			}
+			\ForEach{$\alpha_{v_1, s_1}, \alpha_{v_2, s_2} \in C: v_2 = v_1 + s_1, (\alpha_{v_1, s_1}, \alpha_{v_2, s_2}) \notin E, s_1 \ge min\_dir$}{
+				\If{$s_1 \notin B(\alpha_{v_2,s_2})$}{
+					add $(\alpha_{v_1, s_1}, \alpha_{v_2, s_2})$ to $E$\\
+					// rebuild used direction set for $\alpha_{v_1, s_1}$\\
+					\ForEach{$v \in BFS(CDG(C, \bar{E}), \alpha_{v_1, s_1})$}{
+						$B(v) = B(v) \cup B(\alpha_{v_2, s_2})$
+					}
+				}
+			}
+			\label{alg:add_additional_edge}
+			\caption{Adding edges to CDG.}
+		\end{algorithm}
+	\end{center}
+\end{figure}
+
 TopSortFilter algorithm:

### 2019-12-03 blackholes/pavt/article.tex
Algorithms description and pseudocode

@@
 
-\begin{figure}[H]
-	\begin{center}
-		\begin{algorithm}[H]
-			\SetAlgoLined
-			\SetKwInOut{Input}{Input}
-			\SetKwInOut{Output}{Output}
-			\Input{$CDG(C, E)$ -- channel dependency graph for a system topology with 
-				direction order routing, \\
-				min\_dir -- directions, which are less than this will not be considered}
-			\Output{$CDG$ with added edges}
-			// build the used direction sets\\ 
-			\ForEach{$v \in C$}{
-				$B(v) = \{s_i: \alpha_{u_i, s_i} \in BFS(CDG(C, E), v)\}$
-			}
-			\ForEach{$\alpha_{v_1, s_1}, \alpha_{v_2, s_2} \in C: v_2 = v_1 + s_1, (\alpha_{v_1, s_1}, \alpha_{v_2, s_2}) \notin E, s_1 \ge min\_dir$}{
-				\If{$s_1 \notin B(\alpha_{v_2,s_2})$}{
-					add $(\alpha_{v_1, s_1}, \alpha_{v_2, s_2})$ to $E$\\
-					// rebuild used direction set for $\alpha_{v_1, s_1}$\\
-					\ForEach{$v \in BFS(CDG(C, \bar{E}), \alpha_{v_1, s_1})$}{
-						$B(v) = B(v) \cup B(\alpha_{v_2, s_2})$
-					}
-				}
-			}
-			\label{alg:add_additional_edge}
-			\caption{Adding edges to CDG.}
-		\end{algorithm}
-	\end{center}
-\end{figure}
-
 %TODO: осознать
@@
 special one should be topologically lower than non-special. Otherwise it would be one node basis.
-
 To sum up: if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.
@@
 one non-basic node.
-First, we need to know what special nodes there are. To find out, we will take every global root of condensated graph and find topoological sort of its closure.
+\begin{enumerate}
+    \item First, we need to know what special nodes there are. This step can be precalculated. 
+    To find out, we will take every global root of condensated graph and find topoological sort of its closure.
     Those vertices, which suffice the rule above, we will mark as "special". 
-This step can be precalculated.
-Now, we shall calculate how many special nodes for every root there is in the given set. If there is more than one special node for root - the set is to be called non-basis and
-omitted. Also we shall check if there are any non-special nodes lower than the special ones, and if those found - omit the node set.
+    \item Now, we shall calculate how many special nodes for every root there is in the given set. If there is more than one special node for root - the set is to be called non-basis and
+    omitted. 
+    \FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }
+    \item Also we shall check if there are any non-special nodes lower than the special ones in related root topsort order, and if those found - omit the node set.
+    When we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, omit this set.
+    \FIXME{partially implemented: здесь условие выглядит несколько иначе: в упорядоченном наборе кандидатов специальная вершина может быть только первой. Впрочем, возможно, что тут баг. }
+    \item Otherwise, we have a valid basis of a blackhole.
+    To acquire blackhole we should, according to definition of basis, unite all the closures of basic nodes.
+    Finally, we check for candidate set weak connectivity and if it is present - print blackhole.
+\end{enumerate}
 
@@
 			\SetKwInOut{Output}{Output}
-			\Input{$G(V,E)$ -- directed acyclic graph (Graph condensation)}
-			\Output{$Special$ -- }
-			// build the used direction sets\\ 
-			\ForEach{$v \in C$}{
-				$B(v) = \{s_i: \alpha_{u_i, s_i} \in BFS(CDG(C, E), v)\}$
-			}
-			\ForEach{$\alpha_{v_1, s_1}, \alpha_{v_2, s_2} \in C: v_2 = v_1 + s_1, (\alpha_{v_1, s_1}, \alpha_{v_2, s_2}) \notin E, s_1 \ge min\_dir$}{
-				\If{$s_1 \notin B(\alpha_{v_2,s_2})$}{
-					add $(\alpha_{v_1, s_1}, \alpha_{v_2, s_2})$ to $E$\\
-					// rebuild used direction set for $\alpha_{v_1, s_1}$\\
-					\ForEach{$v \in BFS(CDG(C, \bar{E}), \alpha_{v_1, s_1})$}{
-						$B(v) = B(v) \cup B(\alpha_{v_2, s_2})$
+			\Input{$G(V,E)$ -- directed acyclic graph (Graph condensation) \\
+                                $Cand(G)$ -- set of nodes; candidate to be basis}
+			\Output{$True/False$}
+                        // (1) precalc special nodes for every root \\
+                        \ForEach{$r \in Roots(G)$} {
+                            $Special_r = \emptyset$ \\
+                            $tsOrder_r = TopSort(r)$ \\
+                            \For{$i = 0; i<|tsOrder_r|; i=i + 1$} {
+                                $v = tsOrder_r[i]$ \\
+                                $C_v = Closure(v)$ \\
+                                \If{$|C_v| == i$} {
+                                    $Special_r = Special_r \cup v$ \\
                                 }
@@
                         }
-			\label{alg:add_additional_edge}
-			\caption{Adding edges to CDG.}
+			// (2) check for excluding specials \\
+			\ForEach{$v \in Cand(G)$} {
+                            \ForEach{$r \in Roots(G)$} {
+                                \If{$v \in Special_r$} {
+                                    $Spec_r = Spec_r \cup v$ \\
+                                    \If{$|Spec_r| > 1$} {
+                                        \Return $False$ \\
+                                    }
+                                }
+                            }
+			}
+                        // (3) Check for candidates reachable from one another \\
+                        \ForEach{$(v, u) \in Cand(G) \times Cand(G)$} {
+                            \If{v is reachable from u} {
+                                \Return $False$ \\
+                            }
+                        }
+                        \Return $True$
+			\label{alg:isbasis}
+			\caption{BasisOrNot}
 		\end{algorithm}
@@
 
-TopSortFilter algorithm:
-\begin{enumerate}
-    \item For every node 
-    \item Calculate topological sort order
-    \item Calculate closure size for every vertex
-    \item For every node
-\end{enumerate}
+\begin{figure}[H]
+	\begin{center}
+		\begin{algorithm}[H]
+			\SetAlgoLined
+			\SetKwInOut{Input}{Input}
+			\SetKwInOut{Output}{Output}
+			\Input{$G(V,E)$ -- directed graph}
+			\Output{$Blackhole$ -- set of blackholes of different sizes}
+
+                        $Blackhole = \emptyset$\\
+                        $Build\ graph\ condensation\ Cond(V',E')$\\
+                        \For{$i = 1\ to\ |V'|$} {
+                            \ForEach{$B_i \in V'(i)$} {
+                                \If{$BasisOrNot(Cond(V',E'),B_i)$} {
+                                    \If{$B_i$\ is\ weakly\ connected} {
+                                        $BlackHole = Blackhole \cup B_i$ \\
+                                    }
+                                }
+                            }
+                        }
+                        \Return $Blackhole$
+			\label{alg:topsort}
+			\caption{TopSort}
+		\end{algorithm}
+	\end{center}
+\end{figure}
+
+\begin{figure}[H]
+	\begin{center}
+		\begin{algorithm}[H]
+			\SetAlgoLined
+			\SetKwInOut{Input}{Input}
+			\SetKwInOut{Output}{Output}
+			\Input{$G$ -- directed graph \\
+                                $V$ -- set of all nodes \\
+                                $E$ -- set of all edges \\
+                                $n$ -- max number of nodes each blackhole can contain}
+			\Output{$Blackhole$ -- 1 to n-node blackhole set of G}
+
+                        $Blackhole = \emptyset$ \\
+                        \For{$i = 1\ to\ n$} {
+                            $P_i = \{v | d_{out}(v) < i\}$ \\
+                            \ForEach{$v \in P_i$} {
+                                \If{$v \notin C_{i-1}$} {
+                                    \If{$at\ least\ one\ of\ v's\ directed\ successors\ are\ not\ in\ P_i$} {
+                                        $remove\ v\ from\ P_i$\\
+                                        $remove\ all\ v's\ predecessors\ from\ P_i$\\
+                                    }
+                                }
+                            }
+                        }
+                        $C_i = P_i$\\
+                        \ForEach{$v \in C_i$} {
+                            \If{$|Closure(v)| == i$} {
+                                $Blackhole = Blackhole \cup Closure(v)$
+                            }
+                            \If{$|Closure(v)| >= i$} {
+                                $remove\ v\ from\ C_i$ \\
+                                $remove\ all\ v's\ predecessors\ from\ C_i$ \\
+                            }
+                        }
+                        $F_i = C_i$\\
+                        \ForEach{$B \in F_i(i)$} {
+                            \If{$G(B)\ is\ weakly\ connected$} {
+                                \If{$d_{out}(B) == 0$} {
+                                    $Blackhole = Blackhole \cup B$ \\
+                                }
+                            }
+                        }
+                        \Return $Blackhole$
+			\label{alg:iblackhole}
+			\caption{iBlackHole}
+		\end{algorithm}
+	\end{center}
+\end{figure}
 %
 \section{Experimental Results}
-To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey and TopSort) on a 
+To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey \ref{alg:iblackhole} and TopSort ) on a 
 set of graphs.

### 2019-12-04 blackholes/pavt/article.tex
Algos and definitions

@@
 %
-\subsubsection{Subsection example}
-%
-
-%
 \subsection{Related Work}
@@
 As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
-For certain graph types (RMAT, SmallWorld) it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
+For certain graph types (RMAT, SmallWorld\cite{watts1999networks}) it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
 In such a case, it would be nice to consider large SCC as a single node, aggregating all incoming and outcoming peripheral edges
@@
 \begin{definition}
-TODO: Graph condensation definition
+Set of nodes in a directed graph is called a Strongly Connected Component (SCC) if every node is reachable from every other node.
 \end{definition}
 
-In this article we use the following algorithm for finding graph condensation.
+\begin{definition}
+If every SCC of directed graph is contracted to a single node, the resulting graph is directed acyclic graph, which is call the condensation of G.
+\end{definition}
+
+In this article we use the following algorithm for finding graph condensation. \ref{alg:condensation}
 It was independently proposed by Kosaraju and Sharir back in 1979. \FIXME{Ссылку на эту статью}
 
-First, we find topological sort of the initial graph. It takes O(E + V).
+First, we find topological sort of the initial graph..
+
 \begin{definition}
-TODO: Topological sort
+Topological sort or topological ordering of a directed graph is a linear ordering of 
+its vertices such that for every directed edge $(u,v)$ from node $u$ to node $v$,
+$v$ comes before $u$ in the ordering.
 \end{definition}
@@
 \begin{definition}
-TODO: Reachable
+
 \end{definition}
@@
 
+\begin{figure}[H]
+	\begin{center}
+		\begin{algorithm}[H]
+			\SetAlgoLined
+			\SetKwInOut{Input}{Input}
+			\SetKwInOut{Output}{Output}
+			\Input{$G(V,E)$ -- directed graph}
+			\Output{$G'(V',E')$ -- directed acyclic graph }
+                        $tsOrder = TopSort(G)$ \\
+                        $scc = [-1\ for\ i = 1\ to\ |V|] $ \\ 
+                        $component = 0$ \\
+                        \ForEach{$v \in tsOrder$} {
+                            \If{$scc[v] == -1$} {
+                                $C_v = Closure(v)$ \\
+                                \ForEach{$u \in C_v$} {
+                                    $scc[u] = component$ \\
+                                }
+                                $component = component + 1$ \\
+                            }
+                        }
+
+                        $E' = \emptyset$ \\
+                        $V' = \emptyset$ \\
+                        \ForEach{$(u,v) \in V \times V$} {
+                            \If{$scc[u] \neq scc[v]$} {
+                               $V' = V' \cup \{scc[u], scc[v]\}$ \\ 
+                               $E' = E' \cup (scc[u], scc[v])$ \\ 
+                            }
+                        }
+                        \Return $G'$
+			\label{alg:condensation}
+			\caption{Graph condensation. Kosarayu, Sharir}
+		\end{algorithm}
+	\end{center}
+\end{figure}
+
 As a result of the algorithm, described above, we have each node associated with certain SCC.
 SCCs itself form a graph with number of nodes less than or equal to one of the initial graph.
-
 Finally, we need to remove edges duplicates and the graph condesation is built.
@@
 
-\begin{definition}
-TODO: simple blackhole
+\begin{definition}{Root of blackhole.\\}
+Root of blackhole(subgraph) is a node, which has no incoming edges from other nodes in blackhole.
 \end{definition}
 
-\begin{definition}
-TODO: root of blackhole
+\begin{definition}{Blackhole basis. \\}
+Set of all roots of blackhole is called balckhole basis
 \end{definition}
 
-\begin{definition}
-TODO: basis of blackhole
+\begin{definition}{Simple blackhole. \\}
+A blackhole is called simple if its basis has consist of a single node, i.e. it has single root.
 \end{definition}
 
-\begin{definition}
-TODO: complex blackhole
-\end{definition}
-
-\begin{definition}
-TODO:
+\begin{definition}{Comples blackhole. \\}
+A blackhole is called complex if its basis consists of more than one node.
 \end{definition}
@@
 Imagine, we are given a graph and a set of nodes. What is the algorithm to decide whether these nodes form the basis of a blackhole? We shall decide if there's at least
-one non-basic node.
+one non-basic node. It's done by \ref{alg:isbasis}
 \begin{enumerate}
@@
     \item Also we shall check if there are any non-special nodes lower than the special ones in related root topsort order, and if those found - omit the node set.
-    When we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, omit this set.
+    \item When we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, omit this set. 
     \FIXME{partially implemented: здесь условие выглядит несколько иначе: в упорядоченном наборе кандидатов специальная вершина может быть только первой. Впрочем, возможно, что тут баг. }
@@
     To acquire blackhole we should, according to definition of basis, unite all the closures of basic nodes.
-    Finally, we check for candidate set weak connectivity and if it is present - print blackhole.
+    Finally, we check for weak connectivity and if it is present - print blackhole.
 \end{enumerate}
+
+\begin{remark}
+Of course, we still have to spend some computation time in order to decide, whether
+the candidate is worth checking or not, and we understand, that sometimes it can make process even more slower than before, but experimental results demostrate its efficiency.
+Our goal is to avoid this costly procedure of acquiring blackhole if possible.
+\end{remark}
 

### 2019-12-04 blackholes/pavt/article.tex
All section are done, except for introductin. Lacks accuracy in rendering. Results are only in tables. Some inaccuracies are possible

@@
 \end{abstract}
-%
+
 \section{Blackhole search in directed graph}
@@
 
-%
-\subsection{Task statement}
-%
+\subsection{Task statement and Preliminaries}
+In this section we provide some basic notations and definitions, which will be used in this paper.
 
-%
+Previously team of University of New Jersey gave \cite{li2010detecting} general and simplified problem formulation for blackhole search
+in directed graph. In this paper we consider simplified problem formulation, applied to directed unweighted graph.
+
+First, consider a directed graph G = G(V,E), where $V$ is the set of all nodes and E is the set if all edges.
+Assume, that we deal with the general case of graph, without any special restrictions, unless the opposite is stated.
+
+\begin{definition}{Blackhole}
+Subset of nodes $B \in V$ is called a blackhole if there is no such pair of nodes $(u,v)$ that $u \in B$ and $v \in V \backslash B$
+and $(u,v) \in E$. 
+$G'(B, E')$ must form a weakly connected subgraph of G(V,E).
+\end{definition}
+
+The task is to find in unweighted directed graph as many blackholes as possible in restricted time.
+
+To support further discussion, we provide several more definitions.
+
+\begin{definition}
+Given a directed graph $G(V,E),\ v_i \in V,\ e_i \in E$, where $e_i = (v_{i-1}, v_i)$. We say that sequence $v_0e_1v_1e_2v_2...e_{k}v_{k}$
+forms a directed path from $v_0$ to $v_k$, if $v_i \neq v_j$ for all $0 <= i, j <= k, i \neq j$. The length of this directed path is $k$.
+\end{definition}
+
+\begin{definition}
+Given a directed graph $G(V,E), u,v \in V$. We say that $v$ is reachable from $u$ if there is a directed path that starts from $u$ and ends at $v$.
+\end{definition}
+
+\begin{definition}
+Given a directed graph $G(V,E), u,v \in V$. If $v$ is reachable from $u$, then we say $u$ is a predecessor of $v$ and $v$ is a successor if $u$.
+If there is an edge from $u$ to $v$, then $u$ is a direct predecessor of $v$, and $v$ is a direct successor if $u$.
+\end{definition}
+
+\begin{definition}
+Set of nodes in a directed graph is called a Strongly Connected Component(SCC) if every node is reachable from every other node.
+\end{definition}
+
+\begin{definition}
+Given a directed graph $G(V,E), v \in V$. The closure of $v$, denoted as $Closure(v)$, is defined as $Closure(v) = \{u\ |\ there\ is\ a\ directed\ path\ from\ v\ to\ s\} \cup \{v\}$.
+In other words: closure of node $v$ is a set of all nodes reachable from $v$, including $v$.
+\end{definition}
+
+The following lemmas will be given without proof. It is provided in \cite{li2010detecting}.
+
+\begin{lemma}
+If a node $v \in B$, where $B \subseteq V$ is a blackhole, then all the successors of $v$ are all in $B$.
+\end{lemma}
+
+\begin{lemma}
+If a node $v \in B$, where $B \subseteq V$ is a blackhole, then closure of $v$ is a subset of $B$.
+\end{lemma}
+
+\begin{lemma}
+Closure of $v$ forms a blackhole.
+\end{lemma}
+
+We will need the following statement to discuss issues of known approaches.
+\begin{lemma}
+Given graph $G(V,E)$, blackhole $B \subseteq V$, SCC $S \subseteq V$ and $v \in V : v \in S$. Then $\forall s \in S$ will be true $s \in B$.
+If any node in SCC belongs to blackhole, then all the nodes of this SCC belong to the same blackhole.
+\end{lemma}
+\begin{proof}
+    In SCC any node is reachable from any other node. This means, that any node in SCC contains the whole SCC in its closure.
+\end{proof}
+
 \subsection{Related Work}
-\cite{li2010detecting,li2012mining,li2014mining,hong2015detecting}
+The main contributor to the topic is the New Jersey University. They both formulated the task and provided an algorithm for
+blackhole detection in case of unweighted directed graph \cite{li2010detecting}. Two years later they published an algorithm
+for approximate blackholes search in general case (weighted graph)\cite{li2012mining,li2014mining}. Hong et al \cite{hong2015detecting} had
+an unusual application for blackhole mining in the city environment. They introduced original algorithm. Semenov et al. \cite{semenov2017survey}
+in 2017 provided a survey of common approaches to anti - money laundering by graph mining.
 
-%
-\subsection{Known Issues}
-%
+\subsection{Known issues}
+In this section we shortly discuss applicability of iBlackhole algorithm.
+iBlackhole \ref{alg:iblackhole} algorithm is designed for searching blackholes of fixed size. It has a great resource of parallelism, which is demonstrated
+by tests. However, survey by Li et al, doesn't cover case of large scale graphs, as they only run their algorithm on 1500 node networks, which seem quite
+a small size.
+Also, it is worth mentioning that there are graphs, which has a small amount of blackholes in relation to number of nodes. They consist
+of several large SCCs and relatively small amount of independent leaves (or roots). Here arise some questions considering applicability of
+iBlackhole algorithm:
+\begin{itemize}
+\item How do we know if there are blackholes of certain size? We have to check all the possible sizes, which can take too long even in parallel.
+\item If we check blackhole sizes consecutively, what is the expected time of the first blackhole found?
+\item How do we reduce search field for the BruteForce stage and avoid repeatedly checking the same combinations for different sizes?
+\end{itemize}
 
-%
-\section{Algorithm Design}
+Let's consider the following example:
+
+Graph consisting of $(10^6)$ nodes, all of them form a single SCC. It means that there is only one blackhole - the whole graph.
+Unless we do not know anything about the graph structrure and do not have large enough cluster, we have one to a million chance
+to efficiently choose target size of blackhole. In general case, we have to check $10^6$ possible sizes.
+
+Another example is common SmallWorld\cite{watts1999networks} graph. It has most of its nodes in one large SCC, couple of leaves and couple of roots.
+With iBlackhole algorithm and consequtive approach we will quickly discover all the small blackholes and then for a long time will try to discover
+blackholes inside an SCC.
+
+\begin{figure}[H]
+	\begin{center}
+		\begin{algorithm}[H]
+			\SetAlgoLined
+			\SetKwInOut{Input}{Input}
+			\SetKwInOut{Output}{Output}
+			\Input{$G$ -- directed graph \\
+                                $V$ -- set of all nodes \\
+                                $E$ -- set of all edges \\
+                                $n$ -- max number of nodes each blackhole can contain}
+			\Output{$Blackhole$ -- 1 to n-node blackhole set of G}
+
+                        $Blackhole = \emptyset$ \\
+                        \For{$i = 1\ to\ n$} {
+                            $P_i = \{v | d_{out}(v) < i\}$ \\
+                            \ForEach{$v \in P_i$} {
+                                \If{$v \notin C_{i-1}$} {
+                                    \If{$at\ least\ one\ of\ v's\ directed\ successors\ are\ not\ in\ P_i$} {
+                                        $remove\ v\ from\ P_i$\\
+                                        $remove\ all\ v's\ predecessors\ from\ P_i$\\
+                                    }
+                                }
+                            }
+                        }
+                        $C_i = P_i$\\
+                        \ForEach{$v \in C_i$} {
+                            \If{$|Closure(v)| == i$} {
+                                $Blackhole = Blackhole \cup Closure(v)$
+                            }
+                            \If{$|Closure(v)| >= i$} {
+                                $remove\ v\ from\ C_i$ \\
+                                $remove\ all\ v's\ predecessors\ from\ C_i$ \\
+                            }
+                        }
+                        $F_i = C_i$\\
+                        \ForEach{$B \in F_i(i)$} {
+                            \If{$G(B)\ is\ weakly\ connected$} {
+                                \If{$d_{out}(B) == 0$} {
+                                    $Blackhole = Blackhole \cup B$ \\
+                                }
+                            }
+                        }
+                        \Return $Blackhole$
+			\label{alg:iblackhole}
+			\caption{iBlackHole}
+		\end{algorithm}
+	\end{center}
+\end{figure}
+
+\section{Algorithm design}
 We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
@@
 The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in a restricted time.
-%
 
 %
-\subsection{Graph Preprocessing}
-\FIXME{не забыть описать в KnownIssues}
-
+\subsection{Graph preprocessing}
 \FIXME{не забыть вставить ссылки SmallWorld \cite{watts1999networks}}
 
-As it was described in the Known Issues section, large SCCs waste lots of computational time without adding any new blackholes. 
+As it was described in the Known issues section, large SCCs waste lots of computational time without adding any new blackholes. 
 For certain graph types (RMAT, SmallWorld\cite{watts1999networks}) it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
@@
 On the preprocessing stage of the algorithm we reduce every SCC to a single node. The algorithm is not new and known as a graph condensation.
-
 More formally:
-
-\begin{definition}
-Set of nodes in a directed graph is called a Strongly Connected Component (SCC) if every node is reachable from every other node.
-\end{definition}
 
@@
 $v$ comes before $u$ in the ordering.
-\end{definition}
-
-\begin{definition}
-Closure of node v is a set of all nodes reachable from v, including v.
 \end{definition}
@@
 \begin{definition}{Blackhole basis. \\}
-Set of all roots of blackhole is called balckhole basis
+Set of all roots of blackhole is called blackhole basis
 \end{definition}
@@
 \begin{definition}{Simple blackhole. \\}
-A blackhole is called simple if its basis has consist of a single node, i.e. it has single root.
+A blackhole is called simple if its basis consists of a single node, i.e. it has single root.
 \end{definition}
 
-\begin{definition}{Comples blackhole. \\}
+\begin{definition}{Complex blackhole. \\}
 A blackhole is called complex if its basis consists of more than one node.
@@
 
-\FIXME{определить $G=(V, E)$}
-\FIXME{определить v[i]}
-
 Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
@@
 in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
-conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsort[i]$. We will call such nodes special.
+conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder[i]$. We will call such nodes special.
 
@@
 
-\begin{figure}[H]
-	\begin{center}
-		\begin{algorithm}[H]
-			\SetAlgoLined
-			\SetKwInOut{Input}{Input}
-			\SetKwInOut{Output}{Output}
-			\Input{$G$ -- directed graph \\
-                                $V$ -- set of all nodes \\
-                                $E$ -- set of all edges \\
-                                $n$ -- max number of nodes each blackhole can contain}
-			\Output{$Blackhole$ -- 1 to n-node blackhole set of G}
-
-                        $Blackhole = \emptyset$ \\
-                        \For{$i = 1\ to\ n$} {
-                            $P_i = \{v | d_{out}(v) < i\}$ \\
-                            \ForEach{$v \in P_i$} {
-                                \If{$v \notin C_{i-1}$} {
-                                    \If{$at\ least\ one\ of\ v's\ directed\ successors\ are\ not\ in\ P_i$} {
-                                        $remove\ v\ from\ P_i$\\
-                                        $remove\ all\ v's\ predecessors\ from\ P_i$\\
-                                    }
-                                }
-                            }
-                        }
-                        $C_i = P_i$\\
-                        \ForEach{$v \in C_i$} {
-                            \If{$|Closure(v)| == i$} {
-                                $Blackhole = Blackhole \cup Closure(v)$
-                            }
-                            \If{$|Closure(v)| >= i$} {
-                                $remove\ v\ from\ C_i$ \\
-                                $remove\ all\ v's\ predecessors\ from\ C_i$ \\
-                            }
-                        }
-                        $F_i = C_i$\\
-                        \ForEach{$B \in F_i(i)$} {
-                            \If{$G(B)\ is\ weakly\ connected$} {
-                                \If{$d_{out}(B) == 0$} {
-                                    $Blackhole = Blackhole \cup B$ \\
-                                }
-                            }
-                        }
-                        \Return $Blackhole$
-			\label{alg:iblackhole}
-			\caption{iBlackHole}
-		\end{algorithm}
-	\end{center}
-\end{figure}
 %
@@
 Size $i$ means that graph has $2^i$ nodes and $32*V$ edges.
-Actual count of edges to process can differ from these ideal values, as duplicate edges and self-loops are omitted when reading graph from file.
+Actual count of edges to process can differ from these ideal values, as duplicate edges and self-loops are omitted when readinggraph from file.
 Configuration of the considered graphs is presented in Table \ref{tabular:graphs}.
@@
 
+As it is shown in the Tables \ref{tabular:tableresults}, \ref{tabular:tableresults2}, our TopSort approach to blackhole search shows better perfoermance on small RMAT graphs.
+And it stays in advantage up to RMAT-22. Apparently, condensation preprocessing takes more time on larger graphs and increases time spent before finding
+the first blackhole. Nevertheless our algorithm shows significantly better performance in those cases when preprocessing saves more time. Also, we suppose
+that total working time on large graphs for our algorithm will be less, and this is the point for future survey.
+
+\FIXME{Это, вообще говоря, странно}
+SSCA2 graphs in our survey can be characterized with absolute absence of SCCs consisting of more than one node. Preprocessing here doesn't give any
+certain advantage, but only takes time. As we can see, on smaller graphs our approach beats iBlackhole algorithm, but slows down on larger graphs.
+We guess that the reason is less efficient order of search for our algorithm. One used in iBlackHole will faster print smaller balckholes, and later will try
+to find larger ones, meanwhile our approach will check blackholes of all sizes.
+
+Finally, we should point out, that our preprocessing dramatically speeds up blackhole search on the graphs, where no blackholes present at all. Brute force is
+avoided completely in this case, making calculation time impressively small in comparison to iBlackhole algorithm.
+
 \begin{table}[]
@@
 \end{table}
+\pagebreak
 
-\pagebreak
-%
-
-%
 \section{Conclusion}
@@
 \textit{Comment for the reviewer}. In the final paper revision we add parallel implementation of the TopSort algorithm and present much detailed evaluation on the larger graphs.
+
+In this paper we formulated and addressed some issues in application of already known iBlackhole algorithm. We introduced and tested two different ideas of optimization, combined into
+our TopSort algorithm. Those are: graph condensation preprocessing and topsort bruteforce optimization. We tested out approach on three different types of graphs and number of scales.
+In case of large SCCs we demostrate definite advantage over iBlackhole algorithm. Then, we can relatively fast detect total absence of blackholes, but sometimes quite costly preprocessing
+deminishes any boost of productivity and out performance degrades.
 

### 2019-12-05 blackholes/pavt/article.tex
Section 1.1 is almost done

@@
 
-\section{Blackhole search in directed graph}
+\section{Blackhole Search in Directed Graph}
 applications Anti-Money Laundering \cite{semenov2017survey}
 
-\subsection{Task statement and Preliminaries}
-In this section we provide some basic notations and definitions, which will be used in this paper.
+\subsection{Problem Statement and Preliminaries}
+In this section we provide some basic notations and definitions, which will be used in this paper. In this paper we consider simplified problem formulation, applied to directed unweighted graph.
 
-Previously team of University of New Jersey gave \cite{li2010detecting} general and simplified problem formulation for blackhole search
-in directed graph. In this paper we consider simplified problem formulation, applied to directed unweighted graph.
+First, consider a directed graph $G = G(V,E)$, where $V$ is a set of all nodes and $E$ is a set if all edges of the graph. 
+\FIXME{Мне кажется, это лишнее предложение, так не пишут в статьях:} Assume that we deal with the general case of graph, without any special restrictions, unless the opposite is stated.
 
-First, consider a directed graph G = G(V,E), where $V$ is the set of all nodes and E is the set if all edges.
-Assume, that we deal with the general case of graph, without any special restrictions, unless the opposite is stated.
-
-\begin{definition}{Blackhole}
-Subset of nodes $B \in V$ is called a blackhole if there is no such pair of nodes $(u,v)$ that $u \in B$ and $v \in V \backslash B$
-and $(u,v) \in E$. 
-$G'(B, E')$ must form a weakly connected subgraph of G(V,E).
+\begin{definition}
+Subset of nodes $B \in V$ is called a blackhole if  the following two conditions are satisfied: 
+1) Subgraph $G'(B, E')$ induced by $B$ is weakly connected, 
+%$|B| \geq 2$,
+and
+2) there is no such pair of nodes $(u,v)$ that $u \in B$, $v \in V \backslash B$ and $(u,v) \in E$.
 \end{definition}
 
-The task is to find in unweighted directed graph as many blackholes as possible in restricted time.
+The goal is to find in unweighted directed graph as many blackholes as possible in restricted time.
 
@@
 
+\FIXME{Удалил возможные кратные ребра в определении 2, это важно?}
 \begin{definition}
-Given a directed graph $G(V,E),\ v_i \in V,\ e_i \in E$, where $e_i = (v_{i-1}, v_i)$. We say that sequence $v_0e_1v_1e_2v_2...e_{k}v_{k}$
-forms a directed path from $v_0$ to $v_k$, if $v_i \neq v_j$ for all $0 <= i, j <= k, i \neq j$. The length of this directed path is $k$.
+For a given directed graph $G(V,E)$ we say that sequence $v_0v_1v_2...v_{k}$ of $v_i \in V, 0 \leq i \leq k$ forms a directed path from $v_0$ to $v_k$, if there are edges $(v_{i-1}, v_i) \in E, 1 \leq i \leq k$ and $v_i \neq v_j$ for all $0 \leq i, j \leq k, i \neq j$.
+The length of this directed path is $k$.
 \end{definition}
@@
 \begin{definition}
-Given a directed graph $G(V,E), u,v \in V$. We say that $v$ is reachable from $u$ if there is a directed path that starts from $u$ and ends at $v$.
+For a given directed graph $G(V,E)$ we say that $v \in V$ is reachable from $u \in V$ if there is a directed path that starts from $u$ and ends at $v$.
 \end{definition}
@@
 \begin{definition}
-Given a directed graph $G(V,E), u,v \in V$. If $v$ is reachable from $u$, then we say $u$ is a predecessor of $v$ and $v$ is a successor if $u$.
+For a given directed graph $G(V,E)$ if $v \in V$ is reachable from $u \in V$, then we say $u$ is a predecessor of $v$ and $v$ is a successor if $u$.
 If there is an edge from $u$ to $v$, then $u$ is a direct predecessor of $v$, and $v$ is a direct successor if $u$.
@@
 \begin{definition}
-Set of nodes in a directed graph is called a Strongly Connected Component(SCC) if every node is reachable from every other node.
+Set of nodes in a directed graph $G$ is called a Strongly Connected Component (SCC) if every node of $G$ is reachable from every other node of $G$.
 \end{definition}
@@
 \begin{definition}
-Given a directed graph $G(V,E), v \in V$. The closure of $v$, denoted as $Closure(v)$, is defined as $Closure(v) = \{u\ |\ there\ is\ a\ directed\ path\ from\ v\ to\ s\} \cup \{v\}$.
-In other words: closure of node $v$ is a set of all nodes reachable from $v$, including $v$.
+For a given directed graph $G(V,E)$. The closure of $v \in V$ is $Closure(v) = \{u\ |\ there\ is\ a\ directed\ path\ from\ v\ to\ u\} \cup \{v\}$.
+In other words, a closure of node $v$ is a set of all nodes reachable from $v$, including $v$.
 \end{definition}
 
-The following lemmas will be given without proof. It is provided in \cite{li2010detecting}.
+The following lemmas are proved in \cite{li2010detecting}.
 
 \begin{lemma}
-If a node $v \in B$, where $B \subseteq V$ is a blackhole, then all the successors of $v$ are all in $B$.
+If a node $v \in B$, where $B \subseteq V$ is a blackhole, then all the successors of $v$ are in $B$.
 \end{lemma}
@@
 \begin{lemma}
-If a node $v \in B$, where $B \subseteq V$ is a blackhole, then closure of $v$ is a subset of $B$.
+If a node $v \in B$, where $B \subseteq V$ is a blackhole, then the closure of $v$ is a subset of $B$.
 \end{lemma}
@@
 \subsection{Related Work}
+
+% Previously 
+Authors from the University of New Jersey gave \cite{li2010detecting} general and simplified problem formulation for blackhole search
+in directed graph.
+
 The main contributor to the topic is the New Jersey University. They both formulated the task and provided an algorithm for
@@
 
-Another example is common SmallWorld\cite{watts1999networks} graph. It has most of its nodes in one large SCC, couple of leaves and couple of roots.
+Another example is common SmallWorld \cite{watts1999networks} graph. It has most of its nodes in one large SCC, couple of leaves and couple of roots.
 With iBlackhole algorithm and consequtive approach we will quickly discover all the small blackholes and then for a long time will try to discover

### 2019-12-06 blackholes/pavt/article.tex
Section 1.1 is done

@@
 
+Semenov et al. \cite{semenov2017survey}
+in 2017 provided a survey of common approaches to anti - money laundering by graph mining.
+
+The online detection of black holes/volcanos can timely
+reflect anomalous events, such as disasters, catastrophic accidents,
+and therefore help keep public safety \cite{hong2015detecting}.
+
 \subsection{Problem Statement and Preliminaries}
@@
 \begin{lemma}
-Given graph $G(V,E)$, blackhole $B \subseteq V$, SCC $S \subseteq V$ and $v \in V : v \in S$. Then $\forall s \in S$ will be true $s \in B$.
-If any node in SCC belongs to blackhole, then all the nodes of this SCC belong to the same blackhole.
+For a given graph $G(V,E)$, SCC $S \subseteq V,\ blackhole\ B \subseteq V$ if $\ \exists v \in S: v \in B$,  then $\forall s \in S$ will be true $s \in B$.
+In other words, if any node in SCC belongs to a blackhole, then all the nodes of this SCC belong to the same blackhole.
 \end{lemma}
 \begin{proof}
-    In SCC any node is reachable from any other node. This means, that any node in SCC contains the whole SCC in its closure.
+    In SCC any node is reachable from any other node. This means, that any node $v$ in SCC contains the whole SCC in its closure. By Lemma 3 the whole SCC belongs to the same blackhole as $v$. 
 \end{proof}
@@
 % Previously 
-Authors from the University of New Jersey gave \cite{li2010detecting} general and simplified problem formulation for blackhole search
+%Authors from the University of New Jersey gave \cite{li2010detecting} general and simplified problem formulation for blackhole search
 in directed graph.
 
-The main contributor to the topic is the New Jersey University. They both formulated the task and provided an algorithm for
-blackhole detection in case of unweighted directed graph \cite{li2010detecting}. Two years later they published an algorithm
-for approximate blackholes search in general case (weighted graph)\cite{li2012mining,li2014mining}. Hong et al \cite{hong2015detecting} had
-an unusual application for blackhole mining in the city environment. They introduced original algorithm. Semenov et al. \cite{semenov2017survey}
-in 2017 provided a survey of common approaches to anti - money laundering by graph mining.
+The main contributor to the blackhole detecting problem is the authors from the University of New Jersey.
+They both formulated the task and provided an algorithm for blackhole detection in case of unweighted directed graph \cite{li2010detecting}.
+Two years later they published an algorithm for approximate blackhole search in general case (weighted graph) \cite{li2012mining,li2014mining}. 
+Hong et al \cite{hong2015detecting} had
+an unusual application for blackhole mining in the city environment. They introduced an original algorithm.
 

### 2019-12-06 blackholes/pavt/article.tex
Graphs added
fix
introduction

@@
 \section{Blackhole Search in Directed Graph}
-applications Anti-Money Laundering \cite{semenov2017survey}
-
-Semenov et al. \cite{semenov2017survey}
-in 2017 provided a survey of common approaches to anti - money laundering by graph mining.
-
-The online detection of black holes/volcanos can timely
-reflect anomalous events, such as disasters, catastrophic accidents,
-and therefore help keep public safety \cite{hong2015detecting}.
+The task of pattern mining in directed graphs finds its aplications in different areas. In 2010 Li et al.\cite{li2010mining} state
+that governments are interested in detecting cases of financial fraud. They formulate the task for finding cases of illegal collaboration, and introduce dual patterns:
+blackhole and volcano. Those appear when a group of traders perform operations only between each other in order to manipulate market.
+Continuing study of different financial fraud casese, in 2017 Semenov et al. \cite{semenov2017survey} provided a survey of common approaches to anti - money laundering by graph mining.
+There are other applications. For example, the online detection of black holes/volcanos can timely 
+reflect anomalous events, such as disasters, catastrophic accidents, and therefore help keep public safety \cite{hong2015detecting}.
 
@@
 
+\begin{figure}[p]
+    \includegraphics[width=\linewidth]{rmat.png}
+    \caption{RMAT}
+    \label{fig:rmat}
+\end{figure}
+\begin{figure}[p]
+    \includegraphics[width=\linewidth]{ssca2.png}
+    \caption{SSCA2}
+    \label{fig:ssca2}
+\end{figure}
+\begin{figure}[p]
+    \includegraphics[width=\linewidth]{ur.png}
+    \caption{UR}
+    \label{fig:ur}
+\end{figure}
+
 \end{document}

### 2019-12-07 blackholes/pavt/article.tex
Next modifications

@@
 \subsection{Known issues}
-In this section we shortly discuss applicability of iBlackhole algorithm.
-iBlackhole \ref{alg:iblackhole} algorithm is designed for searching blackholes of fixed size. It has a great resource of parallelism, which is demonstrated
-by tests. However, survey by Li et al, doesn't cover case of large scale graphs, as they only run their algorithm on 1500 node networks, which seem quite
+In this section we shortly discuss applicability of the iBlackhole algorithm, proposed in \cite{li2010detecting}.
+The iBlackhole algorithm (see Algorithm \ref{alg:iblackhole}) is designed for searching blackholes of fixed size.
+Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which are quite
 a small size.
-Also, it is worth mentioning that there are graphs, which has a small amount of blackholes in relation to number of nodes. They consist
-of several large SCCs and relatively small amount of independent leaves (or roots). Here arise some questions considering applicability of
+Also, there are graphs, which has a small number of blackholes in relation to number of nodes. They consist
+of several large SCCs and relatively small number of independent leaves (or roots). Here arise some questions considering applicability of
 iBlackhole algorithm:
 \begin{itemize}
-\item How do we know if there are blackholes of certain size? We have to check all the possible sizes, which can take too long even in parallel.
-\item If we check blackhole sizes consecutively, what is the expected time of the first blackhole found?
-\item How do we reduce search field for the BruteForce stage and avoid repeatedly checking the same combinations for different sizes?
+\item How do we know if there are blackholes of certain size? We have to check all the possible sizes, which can take too much time even in parallel.
+\item If we check blackhole sizes consecutively, what is the expected time of the first blackhole found? \FIXME{мы отвечаем на этот вопрос?}
+\item How do we reduce the search space for the brute-force stage and avoid repeatedly checking the same combinations for different candidate sizes?
 \end{itemize}
@@
 
-Graph consisting of $(10^6)$ nodes, all of them form a single SCC. It means that there is only one blackhole - the whole graph.
+Graph consisting of $(10^6)$ nodes, all of them form a single SCC. It means that there is only one blackhole -- the whole graph.
 Unless we do not know anything about the graph structrure and do not have large enough cluster, we have one to a million chance
@@
 
-Another example is common SmallWorld \cite{watts1999networks} graph. It has most of its nodes in one large SCC, couple of leaves and couple of roots.
-With iBlackhole algorithm and consequtive approach we will quickly discover all the small blackholes and then for a long time will try to discover
-blackholes inside an SCC.
+Another example is well-known small-world graphs \cite{watts1999networks}. It has most of its nodes in one large SCC, couple of leaves and couple of roots.
+We claim that with looking from small to large size candidates in the iBlackhole algorithm we will quickly discover all the small blackholes and then for a long time will try to discover blackholes inside an SCC.
 
@@
 			\SetKwInOut{Output}{Output}
-			\Input{$G$ -- directed graph \\
-                                $V$ -- set of all nodes \\
-                                $E$ -- set of all edges \\
+			\Input{$G(v, E)$ -- directed graph, $V$ -- set of nodes, $E$ -- set of edges, \\
                    $n$ -- max number of nodes each blackhole can contain}
-			\Output{$Blackhole$ -- 1 to n-node blackhole set of G}
+			\Output{$Blackholes$ -- 1 to n-node blackhole set of G}
 
-                        $Blackhole = \emptyset$ \\
+                        $Blackholes = \emptyset, C_0 = \emptyset$ \\
                         \For{$i = 1\ to\ n$} {
-                            $P_i = \{v | d_{out}(v) < i\}$ \\
+                            $P_i = \{v | d_{out}(v) < i\}$  
+                            \tcp*[f]{$d_{out}(v)$ is an out degree of $v$} \\
                             \ForEach{$v \in P_i$} {
                                 \If{$v \notin C_{i-1}$} {
-                                    \If{$at\ least\ one\ of\ v's\ directed\ successors\ are\ not\ in\ P_i$} {
+                                    \If{$at\ least\ one\ of\ direct\  successors\ of\ v\ are\ not\ in\ P_i$} {
                                         $remove\ v\ from\ P_i$\\
@@
                         }
-                        $C_i = P_i$\\
+                        $C_i = P_i$  \tcp*[f]{$C_i$ -- list of candidates} \\
                         \ForEach{$v \in C_i$} {
                             \If{$|Closure(v)| == i$} {
-                                $Blackhole = Blackhole \cup Closure(v)$
+                                $Blackholes = Blackholes \cup Closure(v)$
                             }
@@
                         }
-                        $F_i = C_i$\\
-                        \ForEach{$B \in F_i(i)$} {
+                        \tcc{Apply the brute-force algorithm to $C_(i)$ --  set of all possible subsets containing $i$ nodes}
+                        \ForEach{$B \in C_i(i)$} {
                             \If{$G(B)\ is\ weakly\ connected$} {
                                 \If{$d_{out}(B) == 0$} {
-                                    $Blackhole = Blackhole \cup B$ \\
+                                    $Blackholes = Blackholes \cup B$ \\
                                 }
@@
                         }
-                        \Return $Blackhole$
+                        \Return $Blackholes$
 			\label{alg:iblackhole}
@@
 
-\section{Algorithm design}
+\FIXME{Мы реализовали этот алгоритм}
+
+\section{Algorithm Design}
 We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
+\FIXME{У нас есть ограничение на суммарное время фазу препроцессинга?}
 The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
@@
 %
-\subsection{Graph preprocessing}
-\FIXME{не забыть вставить ссылки SmallWorld \cite{watts1999networks}}
-
-As it was described in the Known issues section, large SCCs waste lots of computational time without adding any new blackholes. 
-For certain graph types (RMAT, SmallWorld\cite{watts1999networks}) it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
+\subsection{Graph Preprocessing}
+As it was described above, large SCCs waste lots of computational time without adding any new blackholes. 
+For small-world graphs it can be crucial, because one SCC can contain up to 100\% of all graph's nodes. 
 In such a case, it would be nice to consider large SCC as a single node, aggregating all incoming and outcoming peripheral edges
@@
 \begin{definition}
-If every SCC of directed graph is contracted to a single node, the resulting graph is directed acyclic graph, which is call the condensation of G.
+If every SCC of a directed graph $G$ is contracted to a single node, the resulting graph is a directed acyclic graph, which is called the condensation of $G$.
 \end{definition}
 
-In this article we use the following algorithm for finding graph condensation. \ref{alg:condensation}
-It was independently proposed by Kosaraju and Sharir back in 1979. \FIXME{Ссылку на эту статью}
+In this paper we use the algorithm of Sharir for finding graph condensation \cite{sharir1981strong}.
 
@@
 $v$ comes before $u$ in the ordering.
-\end{definition}
-
-\begin{definition}
-
 \end{definition}
@@
 %
-\subsection{Blackholes Search}
+\subsection{Blackhole Search}
 When the graph is prepared, we can move on to searching blackholes. The problem is combinatorial in its nature, therefore, we aim to
@@
 
-\begin{definition}{Root of blackhole.\\}
-Root of blackhole(subgraph) is a node, which has no incoming edges from other nodes in blackhole.
+\begin{definition}
+Root of a blackhole is a node, which has no incoming edges from other nodes in the blackhole.
 \end{definition}
 
-\begin{definition}{Blackhole basis. \\}
-Set of all roots of blackhole is called blackhole basis
+\begin{definition}
+Set of all roots of a blackhole is called the blackhole basis.
 \end{definition}
 
-\begin{definition}{Simple blackhole. \\}
-A blackhole is called simple if its basis consists of a single node, i.e. it has single root.
-\end{definition}
-
-\begin{definition}{Complex blackhole. \\}
-A blackhole is called complex if its basis consists of more than one node.
-\end{definition}
+Note, that a blackhole basis can consist of a single node or more than one node.
 

### 2019-12-07 blackholes/pavt/article.tex
New results and images in pdf

--- 2019-12-06 blackholes/pavt/article.tex
@@
 \hline
-RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
-RMAT & 06 & NewJersey & 1 & 0.0121347 & TopSort & 1 & 0.00034812 \\
-RMAT & 08 & NewJersey & 1 & 1.66685 & TopSort & 1 & 0.00146486 \\
-RMAT & 10 & NewJersey & 6 & 1200 & TopSort & 10 & 0.00684045 \\
-RMAT & 12 & NewJersey & 29 & 1200 & TopSort & 971738 & 1200 \\
-RMAT & 14 & NewJersey & 187 & 1200 & TopSort & 735128 & 1200 \\
-RMAT & 16 & NewJersey & 1192 & 1200 & TopSort & 373705 & 1200 \\
-RMAT & 18 & NewJersey & 5176 & 1200 & TopSort & 10286 & 1200 \\
-RMAT & 20 & NewJersey & 617 & 1200 & TopSort & 48867 & 1200 \\
-RMAT & 22 & NewJersey & 98 & 1200 & TopSort & 0 & 1200 \\
-SSCA2 & 04 & NewJersey & 16 & 0.154763 & TopSort & 16 & 0.000189925 \\
-SSCA2 & 06 & NewJersey & 11 & 1200 & TopSort & 70 & 0.00121528 \\
-SSCA2 & 08 & NewJersey & 42 & 1200 & TopSort & 263314 & 1200 \\
-SSCA2 & 10 & NewJersey & 85 & 1200 & TopSort & 114411 & 1200 \\
-SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
+RMAT 04 & chinese & 1 & 0.00039521 & topsort & 1 & 0.000115912 \\
+RMAT & 06 & chinese & 1 & 0.0040142 & topsort & 1 & 0.00034812 \\
+RMAT & 08 & chinese & 1 & 0.0637587 & topsort & 1 & 0.00146486 \\
+RMAT & 10 & chinese & 6 & 1200 & topsort & 10 & 0.00684045 \\
+RMAT & 12 & chinese & 29 & 1200 & topsort & 971738 & 1200 \\
+RMAT & 14 & chinese & 187 & 1200 & topsort & 735128 & 1200 \\
+RMAT & 16 & chinese & 1192 & 1200 & topsort & 373705 & 1200 \\
+RMAT & 18 & chinese & 5161 & 1200 & topsort & 10286 & 1200 \\
+RMAT & 20 & chinese & 609 & 1200 & topsort & 48867 & 1200 \\
+RMAT & 22 & chinese & 108 & 1200 & topsort & 0 & 1200 \\
+SSCA2 & 04 & chinese & 16 & 0.255101 & topsort & 16 & 0.000189925 \\
+SSCA2 & 06 & chinese & 11 & 1200 & topsort & 70 & 0.00121528 \\
+SSCA2 & 08 & chinese & 42 & 1200 & topsort & 263314 & 1200 \\
+SSCA2 & 10 & chinese & 85 & 1200 & topsort & 114411 & 1200 \\
+SSCA2 & 12 & chinese & 343 & 1200 & topsort & 5870 & 1200 \\
 \end{tabular}
@@
 Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
-SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
-SSCA2 & 16 & NewJersey & 5829 & 1200 & TopSort & 893376 & 1200 \\
-SSCA2 & 18 & NewJersey & 23179 & 1200 & TopSort & 224567 & 1200 \\
-SSCA2 & 20 & NewJersey & 92347 & 1200 & TopSort & 24199 & 1200 \\
-SSCA2 & 22 & NewJersey & 370770 & 1200 & TopSort & 0 & 1200 \\
-UR & 04 & NewJersey & 1 & 0.000223768 & TopSort & 1 & 0.000115029 \\
-UR & 06 & NewJersey & 1 & 0.0170713 & TopSort & 1 & 0.000331235 \\
-UR & 08 & NewJersey & 1 & 2.30256 & TopSort & 1 & 0.00147483 \\
-UR & 10 & NewJersey & 1 & 240.822 & TopSort & 1 & 0.00670724 \\
-UR & 12 & NewJersey & 0 & 1200 & TopSort & 1 & 0.034011 \\
-UR & 14 & NewJersey & 0 & 1200 & TopSort & 1 & 0.47257 \\
-UR & 16 & NewJersey & 0 & 1200 & TopSort & 1 & 3.05003 \\
-UR & 18 & NewJersey & 0 & 1200 & TopSort & 1 & 14.5788 \\
-UR & 20 & NewJersey & 0 & 1200 & TopSort & 1 & 67.7176 \\
-UR & 22 & NewJersey & 0 & 1200 & TopSort & 1 & 373.244
+SSCA2 & 14 & chinese & 1433 & 1200 & topsort & 323671 & 1200 \\
+SSCA2 & 16 & chinese & 5829 & 1200 & topsort & 893376 & 1200 \\
+SSCA2 & 18 & chinese & 23179 & 1200 & topsort & 224567 & 1200 \\
+SSCA2 & 20 & chinese & 92347 & 1200 & topsort & 24199 & 1200 \\
+SSCA2 & 22 & chinese & 370770 & 1200 & topsort & 0 & 1200 \\
+UR & 04 & chinese & 1 & 0.000263955 & topsort & 1 & 0.000115029 \\
+UR & 06 & chinese & 1 & 0.00222731 & topsort & 1 & 0.000331235 \\
+UR & 08 & chinese & 1 & 0.0868632 & topsort & 1 & 0.00147483 \\
+UR & 10 & chinese & 1 & 1.93527 & topsort & 1 & 0.00670724 \\
+UR & 12 & chinese & 1 & 46.1463 & topsort & 1 & 0.034011 \\
+UR & 14 & chinese & 1 & 1169.28 & topsort & 1 & 0.47257 \\
+UR & 16 & chinese & 0 & 1200 & topsort & 1 & 3.05003 \\
+UR & 18 & chinese & 0 & 1200 & topsort & 1 & 14.5788 \\
+UR & 20 & chinese & 0 & 1200 & topsort & 1 & 67.7176 \\
+UR & 22 & chinese & 0 & 1200 & topsort & 1 & 373.244 \\
 \end{tabular}
@@
 \begin{figure}[p]
-    \includegraphics[width=\linewidth]{rmat.png}
+    \includegraphics[width=\linewidth]{rmat.pdf}
     \caption{RMAT}
@@
 \begin{figure}[p]
-    \includegraphics[width=\linewidth]{ssca2.png}
+    \includegraphics[width=\linewidth]{ssca2.pdf}
     \caption{SSCA2}
@@
 \begin{figure}[p]
-    \includegraphics[width=\linewidth]{ur.png}
+    \includegraphics[width=\linewidth]{ur.pdf}
     \caption{UR}

### 2019-12-07 blackholes/pavt/plot.py
plotting utility

@@ -1,0 +1,94 @@
+import matplotlib.pyplot as plt
+
+class Res:
+    def __init__(self, sz, bh, t):
+        self.sz = sz
+        self.bh = bh
+        self.t = t
+
+def PlotOne(figid, graph, name, results, form):
+    time = list()
+    bh_found = list()
+    size = list()
+    for res in results:
+        sz = res.sz
+        bh = res.bh
+        t = res.t
+
+        size.append(sz)
+        bh_found.append(bh)
+        time.append(t)
+
+    fig = plt.figure(figid)
+    fig.tight_layout()
+    top = fig.add_subplot(2, 1, 1)
+    bot = fig.add_subplot(2, 1, 2)
+
+    top.set_yscale('symlog')
+    top.plot(size, time, form, label = name)
+    top.grid(True)
+    top.legend()
+    top.set_xlabel('Graph scale')
+    top.set_ylabel('Time (s)')
+
+    if graph != 'UR':
+        bot.set_yscale('symlog')
+    bot.plot(size, bh_found, form, label = name)
+    bot.grid(True)
+    bot.legend()
+    bot.set_xlabel('Graph scale')
+    bot.set_ylabel('Number of blackholes')
+
+
+def PlotGraph(graph, res, figid):
+    f = ['b', 'r']
+    keys = list(res.keys());
+    for i in range(len(keys)):
+        alg = keys[i]
+        PlotOne(figid, graph, alg, res[alg], f[i])
+
+if __name__ == "__main__":
+    
+
+    rmat = dict()
+
+    rmat["iBlackhole"] = list()
+    rmat["TopSort"] = list()
+
+    ssca = dict()
+
+    ssca["iBlackhole"] = list()
+    ssca["TopSort"] = list()
+
+    ur = dict()
+
+    ur["iBlackhole"] = list()
+    ur["TopSort"] = list()
+
+    with open("results.txt", 'r') as file:
+        for line in file:
+            graph, sz, alg, bh, t = [x.strip() for x in line.strip().split()]
+            sz = int(sz)
+            t = float(t)
+            bh = int(bh)
+    
+            if graph == "RMAT":
+                rmat[alg].append(Res(sz, bh, t))
+            if graph == "SSCA2":
+                ssca[alg].append(Res(sz, bh, t))
+            if graph == "UR":
[…]

### 2019-12-07 blackholes/pavt/article.tex
Fix algo name

@@
 \hline
-RMAT 04 & chinese & 1 & 0.00039521 & topsort & 1 & 0.000115912 \\
-RMAT & 06 & chinese & 1 & 0.0040142 & topsort & 1 & 0.00034812 \\
-RMAT & 08 & chinese & 1 & 0.0637587 & topsort & 1 & 0.00146486 \\
-RMAT & 10 & chinese & 6 & 1200 & topsort & 10 & 0.00684045 \\
-RMAT & 12 & chinese & 29 & 1200 & topsort & 971738 & 1200 \\
-RMAT & 14 & chinese & 187 & 1200 & topsort & 735128 & 1200 \\
-RMAT & 16 & chinese & 1192 & 1200 & topsort & 373705 & 1200 \\
-RMAT & 18 & chinese & 5161 & 1200 & topsort & 10286 & 1200 \\
-RMAT & 20 & chinese & 609 & 1200 & topsort & 48867 & 1200 \\
-RMAT & 22 & chinese & 108 & 1200 & topsort & 0 & 1200 \\
-SSCA2 & 04 & chinese & 16 & 0.255101 & topsort & 16 & 0.000189925 \\
-SSCA2 & 06 & chinese & 11 & 1200 & topsort & 70 & 0.00121528 \\
-SSCA2 & 08 & chinese & 42 & 1200 & topsort & 263314 & 1200 \\
-SSCA2 & 10 & chinese & 85 & 1200 & topsort & 114411 & 1200 \\
-SSCA2 & 12 & chinese & 343 & 1200 & topsort & 5870 & 1200 \\
+RMAT 04 & iBlackhole & 1 & 0.00039521 & TopSort & 1 & 0.000115912 \\
+RMAT & 06 & iBlackhole & 1 & 0.0040142 & TopSort & 1 & 0.00034812 \\
+RMAT & 08 & iBlackhole & 1 & 0.0637587 & TopSort & 1 & 0.00146486 \\
+RMAT & 10 & iBlackhole & 6 & 1200 & TopSort & 10 & 0.00684045 \\
+RMAT & 12 & iBlackhole & 29 & 1200 & TopSort & 971738 & 1200 \\
+RMAT & 14 & iBlackhole & 187 & 1200 & TopSort & 735128 & 1200 \\
+RMAT & 16 & iBlackhole & 1192 & 1200 & TopSort & 373705 & 1200 \\
+RMAT & 18 & iBlackhole & 5161 & 1200 & TopSort & 10286 & 1200 \\
+RMAT & 20 & iBlackhole & 609 & 1200 & TopSort & 48867 & 1200 \\
+RMAT & 22 & iBlackhole & 108 & 1200 & TopSort & 0 & 1200 \\
+SSCA2 & 04 & iBlackhole & 16 & 0.255101 & TopSort & 16 & 0.000189925 \\
+SSCA2 & 06 & iBlackhole & 11 & 1200 & TopSort & 70 & 0.00121528 \\
+SSCA2 & 08 & iBlackhole & 42 & 1200 & TopSort & 263314 & 1200 \\
+SSCA2 & 10 & iBlackhole & 85 & 1200 & TopSort & 114411 & 1200 \\
+SSCA2 & 12 & iBlackhole & 343 & 1200 & TopSort & 5870 & 1200 \\
 \end{tabular}
@@
 Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
-SSCA2 & 14 & chinese & 1433 & 1200 & topsort & 323671 & 1200 \\
-SSCA2 & 16 & chinese & 5829 & 1200 & topsort & 893376 & 1200 \\
-SSCA2 & 18 & chinese & 23179 & 1200 & topsort & 224567 & 1200 \\
-SSCA2 & 20 & chinese & 92347 & 1200 & topsort & 24199 & 1200 \\
-SSCA2 & 22 & chinese & 370770 & 1200 & topsort & 0 & 1200 \\
-UR & 04 & chinese & 1 & 0.000263955 & topsort & 1 & 0.000115029 \\
-UR & 06 & chinese & 1 & 0.00222731 & topsort & 1 & 0.000331235 \\
-UR & 08 & chinese & 1 & 0.0868632 & topsort & 1 & 0.00147483 \\
-UR & 10 & chinese & 1 & 1.93527 & topsort & 1 & 0.00670724 \\
-UR & 12 & chinese & 1 & 46.1463 & topsort & 1 & 0.034011 \\
-UR & 14 & chinese & 1 & 1169.28 & topsort & 1 & 0.47257 \\
-UR & 16 & chinese & 0 & 1200 & topsort & 1 & 3.05003 \\
-UR & 18 & chinese & 0 & 1200 & topsort & 1 & 14.5788 \\
-UR & 20 & chinese & 0 & 1200 & topsort & 1 & 67.7176 \\
-UR & 22 & chinese & 0 & 1200 & topsort & 1 & 373.244 \\
+SSCA2 & 14 & iBlackhole & 1433 & 1200 & TopSort & 323671 & 1200 \\
+SSCA2 & 16 & iBlackhole & 5829 & 1200 & TopSort & 893376 & 1200 \\
+SSCA2 & 18 & iBlackhole & 23179 & 1200 & TopSort & 224567 & 1200 \\
+SSCA2 & 20 & iBlackhole & 92347 & 1200 & TopSort & 24199 & 1200 \\
+SSCA2 & 22 & iBlackhole & 370770 & 1200 & TopSort & 0 & 1200 \\
+UR & 04 & iBlackhole & 1 & 0.000263955 & TopSort & 1 & 0.000115029 \\
+UR & 06 & iBlackhole & 1 & 0.00222731 & TopSort & 1 & 0.000331235 \\
+UR & 08 & iBlackhole & 1 & 0.0868632 & TopSort & 1 & 0.00147483 \\
+UR & 10 & iBlackhole & 1 & 1.93527 & TopSort & 1 & 0.00670724 \\
+UR & 12 & iBlackhole & 1 & 46.1463 & TopSort & 1 & 0.034011 \\
+UR & 14 & iBlackhole & 1 & 1169.28 & TopSort & 1 & 0.47257 \\
+UR & 16 & iBlackhole & 0 & 1200 & TopSort & 1 & 3.05003 \\
+UR & 18 & iBlackhole & 0 & 1200 & TopSort & 1 & 14.5788 \\
+UR & 20 & iBlackhole & 0 & 1200 & TopSort & 1 & 67.7176 \\
+UR & 22 & iBlackhole & 0 & 1200 & TopSort & 1 & 373.244 \\
 \end{tabular}

### 2019-12-07 blackholes/pavt/article.tex
Small updates

--- 2019-12-07 blackholes/pavt/article.tex
@@
 \item How do we know if there are blackholes of certain size? We have to check all the possible sizes, which can take too much time even in parallel.
-\item If we check blackhole sizes consecutively, what is the expected time of the first blackhole found? \FIXME{мы отвечаем на этот вопрос?}
 \item How do we reduce the search space for the brute-force stage and avoid repeatedly checking the same combinations for different candidate sizes?
@@
 We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
-\FIXME{У нас есть ограничение на суммарное время фазу препроцессинга?}
+\FIXME{ограничение на суммарное время - да, 20 минут}
 The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
@@
 its vertices such that for every directed edge $(u,v)$ from node $u$ to node $v$,
-$v$ comes before $u$ in the ordering.
+$v$ comes before $u$ in the topological ordering.
 \end{definition}
@@
 
+\FIXME{избежать путаницы между brute-force в разных местах статьи}
+
 Note, that a blackhole basis can consist of a single node or more than one node.
@@
 
+\FIXME{не совсем понял логику в общем тексте}
 Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
-in preprocessing step, it still can result in lots of computation. It's worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
+in preprocessing step, it still can result in lots of computation. It is worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
 Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
@@
 
-Let's consider a closure of a single blackhole root. Let this closure to be acquired by depth-first search, which basically means, that
-we can iterate over these nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
-in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
-conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder[i]$. We will call such nodes special.
+Let's consider a closure of a single blackhole root $r$. Let this closure to be acquired by depth-first search, which basically means, that
+we can iterate by $i$ over these nodes in topologically sorted order $topsortOrder(r)[i]$. Also, while traversing this graph, we can calculate size of closure for each node
+in the given root closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
+conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder(r)[i]$. We will call such nodes special.
 
@@
 \hline
-RMAT 04 & chinese & 1 & 0.00039521 & topsort & 1 & 0.000115912 \\
-RMAT & 06 & chinese & 1 & 0.0040142 & topsort & 1 & 0.00034812 \\
-RMAT & 08 & chinese & 1 & 0.0637587 & topsort & 1 & 0.00146486 \\
-RMAT & 10 & chinese & 6 & 1200 & topsort & 10 & 0.00684045 \\
-RMAT & 12 & chinese & 29 & 1200 & topsort & 971738 & 1200 \\
-RMAT & 14 & chinese & 187 & 1200 & topsort & 735128 & 1200 \\
-RMAT & 16 & chinese & 1192 & 1200 & topsort & 373705 & 1200 \\
-RMAT & 18 & chinese & 5161 & 1200 & topsort & 10286 & 1200 \\
-RMAT & 20 & chinese & 609 & 1200 & topsort & 48867 & 1200 \\
-RMAT & 22 & chinese & 108 & 1200 & topsort & 0 & 1200 \\
-SSCA2 & 04 & chinese & 16 & 0.255101 & topsort & 16 & 0.000189925 \\
-SSCA2 & 06 & chinese & 11 & 1200 & topsort & 70 & 0.00121528 \\
-SSCA2 & 08 & chinese & 42 & 1200 & topsort & 263314 & 1200 \\
-SSCA2 & 10 & chinese & 85 & 1200 & topsort & 114411 & 1200 \\
-SSCA2 & 12 & chinese & 343 & 1200 & topsort & 5870 & 1200 \\
+RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
+RMAT & 06 & NewJersey & 1 & 0.0121347 & TopSort & 1 & 0.00034812 \\
+RMAT & 08 & NewJersey & 1 & 1.66685 & TopSort & 1 & 0.00146486 \\
+RMAT & 10 & NewJersey & 6 & 1200 & TopSort & 10 & 0.00684045 \\
+RMAT & 12 & NewJersey & 29 & 1200 & TopSort & 971738 & 1200 \\
+RMAT & 14 & NewJersey & 187 & 1200 & TopSort & 735128 & 1200 \\
+RMAT & 16 & NewJersey & 1192 & 1200 & TopSort & 373705 & 1200 \\
+RMAT & 18 & NewJersey & 5176 & 1200 & TopSort & 10286 & 1200 \\
+RMAT & 20 & NewJersey & 617 & 1200 & TopSort & 48867 & 1200 \\
+RMAT & 22 & NewJersey & 98 & 1200 & TopSort & 0 & 1200 \\
+SSCA2 & 04 & NewJersey & 16 & 0.154763 & TopSort & 16 & 0.000189925 \\
+SSCA2 & 06 & NewJersey & 11 & 1200 & TopSort & 70 & 0.00121528 \\
+SSCA2 & 08 & NewJersey & 42 & 1200 & TopSort & 263314 & 1200 \\
+SSCA2 & 10 & NewJersey & 85 & 1200 & TopSort & 114411 & 1200 \\
+SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
 \end{tabular}
@@
 Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
-SSCA2 & 14 & chinese & 1433 & 1200 & topsort & 323671 & 1200 \\
-SSCA2 & 16 & chinese & 5829 & 1200 & topsort & 893376 & 1200 \\
-SSCA2 & 18 & chinese & 23179 & 1200 & topsort & 224567 & 1200 \\
-SSCA2 & 20 & chinese & 92347 & 1200 & topsort & 24199 & 1200 \\
-SSCA2 & 22 & chinese & 370770 & 1200 & topsort & 0 & 1200 \\
-UR & 04 & chinese & 1 & 0.000263955 & topsort & 1 & 0.000115029 \\
-UR & 06 & chinese & 1 & 0.00222731 & topsort & 1 & 0.000331235 \\
-UR & 08 & chinese & 1 & 0.0868632 & topsort & 1 & 0.00147483 \\
-UR & 10 & chinese & 1 & 1.93527 & topsort & 1 & 0.00670724 \\
-UR & 12 & chinese & 1 & 46.1463 & topsort & 1 & 0.034011 \\
-UR & 14 & chinese & 1 & 1169.28 & topsort & 1 & 0.47257 \\
-UR & 16 & chinese & 0 & 1200 & topsort & 1 & 3.05003 \\
-UR & 18 & chinese & 0 & 1200 & topsort & 1 & 14.5788 \\
-UR & 20 & chinese & 0 & 1200 & topsort & 1 & 67.7176 \\
-UR & 22 & chinese & 0 & 1200 & topsort & 1 & 373.244 \\
+SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
+SSCA2 & 16 & NewJersey & 5829 & 1200 & TopSort & 893376 & 1200 \\
+SSCA2 & 18 & NewJersey & 23179 & 1200 & TopSort & 224567 & 1200 \\
+SSCA2 & 20 & NewJersey & 92347 & 1200 & TopSort & 24199 & 1200 \\
+SSCA2 & 22 & NewJersey & 370770 & 1200 & TopSort & 0 & 1200 \\
+UR & 04 & NewJersey & 1 & 0.000223768 & TopSort & 1 & 0.000115029 \\
+UR & 06 & NewJersey & 1 & 0.0170713 & TopSort & 1 & 0.000331235 \\
+UR & 08 & NewJersey & 1 & 2.30256 & TopSort & 1 & 0.00147483 \\
+UR & 10 & NewJersey & 1 & 240.822 & TopSort & 1 & 0.00670724 \\
+UR & 12 & NewJersey & 0 & 1200 & TopSort & 1 & 0.034011 \\
+UR & 14 & NewJersey & 0 & 1200 & TopSort & 1 & 0.47257 \\
+UR & 16 & NewJersey & 0 & 1200 & TopSort & 1 & 3.05003 \\
+UR & 18 & NewJersey & 0 & 1200 & TopSort & 1 & 14.5788 \\
+UR & 20 & NewJersey & 0 & 1200 & TopSort & 1 & 67.7176 \\
+UR & 22 & NewJersey & 0 & 1200 & TopSort & 1 & 373.244
 \end{tabular}
@@
 deminishes any boost of productivity and out performance degrades.
+\FIXME{можно написать в TODO}
+If we check blackhole sizes consecutively, what is the expected time of the first blackhole found?
 
@@
 \begin{figure}[p]
-    \includegraphics[width=\linewidth]{rmat.pdf}
+    \includegraphics[width=\linewidth]{rmat.png}
     \caption{RMAT}
@@
 \begin{figure}[p]
-    \includegraphics[width=\linewidth]{ssca2.pdf}
+    \includegraphics[width=\linewidth]{ssca2.png}
     \caption{SSCA2}
@@
 \begin{figure}[p]
-    \includegraphics[width=\linewidth]{ur.pdf}
+    \includegraphics[width=\linewidth]{ur.png}
     \caption{UR}

### 2019-12-08 blackholes/pavt/article.tex
New mods

@@
 \begin{abstract}
-In this paper we consider the task of finding so called Blackhole pattern in directed unweighed graphs.
+In this paper we consider the problem of finding so called Blackhole pattern in directed unweighted graphs.
 Firstly, we analyse already existing algorithms and approaches. Secondly, we describe how structure of a graph
@@
 and provide results of experimental comparison between our approach and previously developed algorithm.
-% We would like to encourage you to list your keywords within
-% the abstract section using the \keywords{...} command.
 \keywords{Directed networks $\cdot$ Subgraph mining}
@@
 of original SCC.
-On the preprocessing stage of the algorithm we reduce every SCC to a single node. The algorithm is not new and known as a graph condensation.
-More formally:
+On the preprocessing stage of the algorithm we reduce every SCC to a single node. The algorithm is known as a graph condensation.
 
@@
 
-First, we find topological sort of the initial graph..
-
-\begin{definition}
-Topological sort or topological ordering of a directed graph is a linear ordering of 
-its vertices such that for every directed edge $(u,v)$ from node $u$ to node $v$,
-$v$ comes before $u$ in the topological ordering.
-\end{definition}
-
-When we finally have an array, containing graph nodes in the topologically sorted order, we do the following:
-
-Define global array of used nodes. Initially, all nodes are unused.
-Then for every node from bottom to top:
-\begin{enumerate}
-    \item Acquire closure of unused nodes reachable from current node; 
-    \item Mark every node in the closure as used;
-    \item All the acquired nodes belong to the same SCC.
-\end{enumerate}
-
-The fact given in the last step can be shortly explained in the following way:
-topological sort guarantees that any node has index larger than indeces of her children.
-That is why, given that we process nodes from bottom to top, by the moment we process the node, all it's children are already marked used and no longer available for search.
-As all the ways down are prohibited and there's no way up, we will acquire an SCC.
-
-\begin{figure}[H]
-	\begin{center}
-		\begin{algorithm}[H]
-			\SetAlgoLined
-			\SetKwInOut{Input}{Input}
-			\SetKwInOut{Output}{Output}
-			\Input{$G(V,E)$ -- directed graph}
-			\Output{$G'(V',E')$ -- directed acyclic graph }
-                        $tsOrder = TopSort(G)$ \\
-                        $scc = [-1\ for\ i = 1\ to\ |V|] $ \\ 
-                        $component = 0$ \\
-                        \ForEach{$v \in tsOrder$} {
-                            \If{$scc[v] == -1$} {
-                                $C_v = Closure(v)$ \\
-                                \ForEach{$u \in C_v$} {
-                                    $scc[u] = component$ \\
-                                }
-                                $component = component + 1$ \\
-                            }
-                        }
-
-                        $E' = \emptyset$ \\
-                        $V' = \emptyset$ \\
-                        \ForEach{$(u,v) \in V \times V$} {
-                            \If{$scc[u] \neq scc[v]$} {
-                               $V' = V' \cup \{scc[u], scc[v]\}$ \\ 
-                               $E' = E' \cup (scc[u], scc[v])$ \\ 
-                            }
-                        }
-                        \Return $G'$
-			\label{alg:condensation}
-			\caption{Graph condensation. Kosarayu, Sharir}
-		\end{algorithm}
-	\end{center}
-\end{figure}
-
-As a result of the algorithm, described above, we have each node associated with certain SCC.
-SCCs itself form a graph with number of nodes less than or equal to one of the initial graph.
-Finally, we need to remove edges duplicates and the graph condesation is built.
-
-%
-
-%
 \subsection{Blackhole Search}
@@
 
+Further we use the concept of the topological order in a graph.
+
+\begin{definition}
+	Topological sort or topological ordering of a directed graph is a linear ordering of 
+	its vertices such that for every directed edge $(u,v)$ from node $u$ to node $v$,
+	$v$ comes before $u$ in the topological ordering.
+\end{definition}
+
 Let's consider a closure of a single blackhole root $r$. Let this closure to be acquired by depth-first search, which basically means, that
-we can iterate by $i$ over these nodes in topologically sorted order $topsortOrder(r)[i]$. Also, while traversing this graph, we can calculate size of closure for each node
+we can iterate by $i$ over these nodes in topologically sorted order $topsortOrder_r[i]$.
+
+Also, while traversing this graph, we can calculate size of closure for each node
 in the given root closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
-conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder(r)[i]$. We will call such nodes special.
+conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder_r[i]$. We will call such nodes special.
 
@@
 
-Let's consider two nodes in the closure of global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
-special one should be topologically lower than non-special. Otherwise it would be one node basis.
-To sum up: if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.
+%TODO: Осознать до конца
+Let's consider two nodes in the closure of a global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
+the special one should be topologically lower than non-special. Otherwise it would be one node basis.
+To summarize, if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.
 
-Imagine, we are given a graph and a set of nodes. What is the algorithm to decide whether these nodes form the basis of a blackhole? We shall decide if there's at least
-one non-basic node. It's done by \ref{alg:isbasis}
-\begin{enumerate}
-    \item First, we need to know what special nodes there are. This step can be precalculated. 
-    To find out, we will take every global root of condensated graph and find topoological sort of its closure.
-    Those vertices, which suffice the rule above, we will mark as "special". 
-    \item Now, we shall calculate how many special nodes for every root there is in the given set. If there is more than one special node for root - the set is to be called non-basis and
-    omitted.
-    \FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }
-    \item Also we shall check if there are any non-special nodes lower than the special ones in related root topsort order, and if those found - omit the node set.
-    \item When we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, omit this set. 
-    \FIXME{partially implemented: здесь условие выглядит несколько иначе: в упорядоченном наборе кандидатов специальная вершина может быть только первой. Впрочем, возможно, что тут баг. }
-    \item Otherwise, we have a valid basis of a blackhole.
+%We shall decide if there's at least
+%one non-basic node.
+
+Algorithm \ref{alg:isbasis} IsBasis tests a set of nodes as a blackhole basis. 
+First, we need to know what special nodes there are. This step can be precalculated. 
+\FIXME{как вычисляются global roots?}
+To find out, we take every global root of the condensated graph and find topological sort of its closure.
+Those vertices, which satisfy the rule above, we mark as special.
+
+\FIXME{зачем второй этап?} 
+Second, we calculate how many special nodes for every root there is in the given set.
+If there is more than one special node for root, the $Cand$ set is called non-basis and will be omitted in the high-level TopSort blackhole mining Algorithm \ref{alg:topsort}.
+
+%\FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }
+
+Also we check if there are any non-special nodes lower than the special ones in the related root topological order, and if those found, then omit the node set.
+
+Third, when we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, we omit this set.
+
+%\FIXME{partially implemented: здесь условие выглядит несколько иначе: в упорядоченном наборе кандидатов специальная вершина может быть только первой. Впрочем, возможно, что тут баг. }
+
+Otherwise, we have a valid basis of a blackhole.
 To acquire blackhole we should, according to definition of basis, unite all the closures of basic nodes.
-    Finally, we check for weak connectivity and if it is present - print blackhole.
-\end{enumerate}
+Finally, we check for weak connectivity and if it is present, then print the blackhole.
 
-\begin{remark}
-Of course, we still have to spend some computation time in order to decide, whether
+Clearly, we still have to spend some computation time in order to decide, whether
 the candidate is worth checking or not, and we understand, that sometimes it can make process even more slower than before, but experimental results demostrate its efficiency.
 Our goal is to avoid this costly procedure of acquiring blackhole if possible.
-\end{remark}
 
@@
 			\SetKwInOut{Output}{Output}
-			\Input{$G(V,E)$ -- directed acyclic graph (Graph condensation) \\
-                                $Cand(G)$ -- set of nodes; candidate to be basis}
-			\Output{$True/False$}
-                        // (1) precalc special nodes for every root \\
+			\Input{$G(V,E)$ -- directed acyclic graph (after graph condensation) \\
+                                $Cand \subset V$ -- set of nodes, candidate to be basis}
+			\Output{$True, if\ Cand\ is\ a\ basis, False\ otherwise$}
+                        \tcc{(1) calculate special nodes for every root} 
                         \ForEach{$r \in Roots(G)$} {
                             $Special_r = \emptyset$ \\
-                            $tsOrder_r = TopSort(r)$ \\
-                            \For{$i = 0; i<|tsOrder_r|; i=i + 1$} {
-                                $v = tsOrder_r[i]$ \\
+                            $build\ topsortOrder_r$ \\
+                            \For{$i = 0; i<|topsortOrder_r|; i=i + 1$} {
+                                $v = topsortOrder_r[i]$ \\
                                 $C_v = Closure(v)$ \\
@@
                         }
-			// (2) check for excluding specials \\
-			\ForEach{$v \in Cand(G)$} {
+			\tcc{(2) check for excluding specials} 
+			\ForEach{$v \in Cand$} {
                             \ForEach{$r \in Roots(G)$} {
@@
 			}
-                        // (3) Check for candidates reachable from one another \\
+                        \tcc{(3) check for candidates reachable from one another}
                         \ForEach{$(v, u) \in Cand(G) \times Cand(G)$} {
@@
 			\label{alg:isbasis}
-			\caption{BasisOrNot}
+			\caption{IsBasis}
 		\end{algorithm}
@@
 			\Input{$G(V,E)$ -- directed graph}
-			\Output{$Blackhole$ -- set of blackholes of different sizes}
+			\Output{$Blackholes$ -- set of blackholes of different sizes}
 
@@
                             \ForEach{$B_i \in V'(i)$} {
-                                \If{$BasisOrNot(Cond(V',E'),B_i)$} {
+                                \If{$IsBasis(Cond(V',E'),B_i)$} {
                                     \If{$B_i$\ is\ weakly\ connected} {
-                                        $BlackHole = Blackhole \cup B_i$ \\
+                                        $Blackholes = Blackholes \cup B_i$ \\
                                     }
@@
                         }
-                        \Return $Blackhole$
+                        \Return $Blackholes$
 			\label{alg:topsort}
@@
 
-As it is shown in the Tables \ref{tabular:tableresults}, \ref{tabular:tableresults2}, our TopSort approach to blackhole search shows better perfoermance on small RMAT graphs.
+As it is shown in the Tables \ref{tabular:tableresults}, \ref{tabular:tableresults2}, our TopSort approach to blackhole search shows better performance on small RMAT graphs.
 And it stays in advantage up to RMAT-22. Apparently, condensation preprocessing takes more time on larger graphs and increases time spent before finding
@@
 
-\begin{table}[]
-\caption{Experimental results}
-\label{tabular:tableresults}
-\begin{center}
-\begin{tabular}{c|c|c|c|c|c|c|c|c}
-Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
-\hline
-RMAT 04 & iBlackhole & 1 & 0.00039521 & TopSort & 1 & 0.000115912 \\
-RMAT & 06 & iBlackhole & 1 & 0.0040142 & TopSort & 1 & 0.00034812 \\
-RMAT & 08 & iBlackhole & 1 & 0.0637587 & TopSort & 1 & 0.00146486 \\
-RMAT & 10 & iBlackhole & 6 & 1200 & TopSort & 10 & 0.00684045 \\
-RMAT & 12 & iBlackhole & 29 & 1200 & TopSort & 971738 & 1200 \\
-RMAT & 14 & iBlackhole & 187 & 1200 & TopSort & 735128 & 1200 \\
-RMAT & 16 & iBlackhole & 1192 & 1200 & TopSort & 373705 & 1200 \\
-RMAT & 18 & iBlackhole & 5161 & 1200 & TopSort & 10286 & 1200 \\
-RMAT & 20 & iBlackhole & 609 & 1200 & TopSort & 48867 & 1200 \\
-RMAT & 22 & iBlackhole & 108 & 1200 & TopSort & 0 & 1200 \\
-SSCA2 & 04 & iBlackhole & 16 & 0.255101 & TopSort & 16 & 0.000189925 \\
-SSCA2 & 06 & iBlackhole & 11 & 1200 & TopSort & 70 & 0.00121528 \\
-SSCA2 & 08 & iBlackhole & 42 & 1200 & TopSort & 263314 & 1200 \\
-SSCA2 & 10 & iBlackhole & 85 & 1200 & TopSort & 114411 & 1200 \\
-SSCA2 & 12 & iBlackhole & 343 & 1200 & TopSort & 5870 & 1200 \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[]
-\caption{Experimental results. Part 2}
-\label{tabular:tableresults2}
-\begin{center}
-\begin{tabular}{c|c|c|c|c|c|c|c|c}
-Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
-SSCA2 & 14 & iBlackhole & 1433 & 1200 & TopSort & 323671 & 1200 \\
-SSCA2 & 16 & iBlackhole & 5829 & 1200 & TopSort & 893376 & 1200 \\
-SSCA2 & 18 & iBlackhole & 23179 & 1200 & TopSort & 224567 & 1200 \\
-SSCA2 & 20 & iBlackhole & 92347 & 1200 & TopSort & 24199 & 1200 \\
-SSCA2 & 22 & iBlackhole & 370770 & 1200 & TopSort & 0 & 1200 \\
-UR & 04 & iBlackhole & 1 & 0.000263955 & TopSort & 1 & 0.000115029 \\
-UR & 06 & iBlackhole & 1 & 0.00222731 & TopSort & 1 & 0.000331235 \\
-UR & 08 & iBlackhole & 1 & 0.0868632 & TopSort & 1 & 0.00147483 \\
-UR & 10 & iBlackhole & 1 & 1.93527 & TopSort & 1 & 0.00670724 \\
-UR & 12 & iBlackhole & 1 & 46.1463 & TopSort & 1 & 0.034011 \\
-UR & 14 & iBlackhole & 1 & 1169.28 & TopSort & 1 & 0.47257 \\
-UR & 16 & iBlackhole & 0 & 1200 & TopSort & 1 & 3.05003 \\
-UR & 18 & iBlackhole & 0 & 1200 & TopSort & 1 & 14.5788 \\
-UR & 20 & iBlackhole & 0 & 1200 & TopSort & 1 & 67.7176 \\
-UR & 22 & iBlackhole & 0 & 1200 & TopSort & 1 & 373.244 \\
-\end{tabular}
-\end{center}
-\end{table}
 \pagebreak
@@
 \section{Conclusion}
-
-\textit{Comment for the reviewer}. In the final paper revision we add parallel implementation of the TopSort algorithm and present much detailed evaluation on the larger graphs.
 
@@
 If we check blackhole sizes consecutively, what is the expected time of the first blackhole found?
+
+\textit{Comment for the reviewer}. In the final paper revision we add parallel implementation of the TopSort algorithm and present much detailed evaluation on the larger graphs.
 

### 2019-12-08 blackholes/pavt/sketch.tex
New mods

First, we find topological sort of the initial graph..

When we finally have an array, containing graph nodes in the topologically sorted order, we do the following:

Define global array of used nodes. Initially, all nodes are unused.
Then for every node from bottom to top:
\begin{enumerate}
	\item Acquire closure of unused nodes reachable from current node; 
	\item Mark every node in the closure as used;
	\item All the acquired nodes belong to the same SCC.
\end{enumerate}

The fact given in the last step can be shortly explained in the following way:
topological sort guarantees that any node has index larger than indeces of her children.
That is why, given that we process nodes from bottom to top, by the moment we process the node, all it's children are already marked used and no longer available for search.
As all the ways down are prohibited and there's no way up, we will acquire an SCC.

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$G(V,E)$ -- directed graph}
			\Output{$G'(V',E')$ -- directed acyclic graph }
			$tsOrder = TopSort(G)$ \\
			$scc = [-1\ for\ i = 1\ to\ |V|] $ \\ 
			$component = 0$ \\
			\ForEach{$v \in tsOrder$} {
				\If{$scc[v] == -1$} {
					$C_v = Closure(v)$ \\
					\ForEach{$u \in C_v$} {
						$scc[u] = component$ \\
					}
					$component = component + 1$ \\
				}
			}

			$E' = \emptyset$ \\
			$V' = \emptyset$ \\
			\ForEach{$(u,v) \in V \times V$} {
				\If{$scc[u] \neq scc[v]$} {
					$V' = V' \cup \{scc[u], scc[v]\}$ \\ 
					$E' = E' \cup (scc[u], scc[v])$ \\ 
				}
			}
			\Return $G'$
			\label{alg:condensation}
			\caption{Graph condensation. Kosarayu, Sharir}
		\end{algorithm}
	\end{center}
\end{figure}

As a result of the condensation algorithm we have each node associated with certain SCC.
SCCs itself form a graph with number of nodes less than or equal to one of the initial graph.
Finally, we need to remove edges duplicates and the graph condesation is built.

\begin{table}[]
	\caption{Experimental results}
	\label{tabular:tableresults}
	\begin{center}
		\begin{tabular}{c|c|c|c|c|c|c|c|c}
			Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
			\hline
			RMAT 04 & iBlackhole & 1 & 0.00039521 & TopSort & 1 & 0.000115912 \\
			RMAT & 06 & iBlackhole & 1 & 0.0040142 & TopSort & 1 & 0.00034812 \\
			RMAT & 08 & iBlackhole & 1 & 0.0637587 & TopSort & 1 & 0.00146486 \\
			RMAT & 10 & iBlackhole & 6 & 1200 & TopSort & 10 & 0.00684045 \\
			RMAT & 12 & iBlackhole & 29 & 1200 & TopSort & 971738 & 1200 \\
			RMAT & 14 & iBlackhole & 187 & 1200 & TopSort & 735128 & 1200 \\
			RMAT & 16 & iBlackhole & 1192 & 1200 & TopSort & 373705 & 1200 \\
			RMAT & 18 & iBlackhole & 5161 & 1200 & TopSort & 10286 & 1200 \\
			RMAT & 20 & iBlackhole & 609 & 1200 & TopSort & 48867 & 1200 \\
			RMAT & 22 & iBlackhole & 108 & 1200 & TopSort & 0 & 1200 \\
			SSCA2 & 04 & iBlackhole & 16 & 0.255101 & TopSort & 16 & 0.000189925 \\
			SSCA2 & 06 & iBlackhole & 11 & 1200 & TopSort & 70 & 0.00121528 \\
			SSCA2 & 08 & iBlackhole & 42 & 1200 & TopSort & 263314 & 1200 \\
			SSCA2 & 10 & iBlackhole & 85 & 1200 & TopSort & 114411 & 1200 \\
			SSCA2 & 12 & iBlackhole & 343 & 1200 & TopSort & 5870 & 1200 \\
		\end{tabular}
	\end{center}
\end{table}

\begin{table}[]
	\caption{Experimental results. Part 2}
	\label{tabular:tableresults2}
	\begin{center}
		\begin{tabular}{c|c|c|c|c|c|c|c|c}
			Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
			SSCA2 & 14 & iBlackhole & 1433 & 1200 & TopSort & 323671 & 1200 \\
			SSCA2 & 16 & iBlackhole & 5829 & 1200 & TopSort & 893376 & 1200 \\
			SSCA2 & 18 & iBlackhole & 23179 & 1200 & TopSort & 224567 & 1200 \\
			SSCA2 & 20 & iBlackhole & 92347 & 1200 & TopSort & 24199 & 1200 \\
			SSCA2 & 22 & iBlackhole & 370770 & 1200 & TopSort & 0 & 1200 \\
			UR & 04 & iBlackhole & 1 & 0.000263955 & TopSort & 1 & 0.000115029 \\
			UR & 06 & iBlackhole & 1 & 0.00222731 & TopSort & 1 & 0.000331235 \\
			UR & 08 & iBlackhole & 1 & 0.0868632 & TopSort & 1 & 0.00147483 \\
			UR & 10 & iBlackhole & 1 & 1.93527 & TopSort & 1 & 0.00670724 \\
			UR & 12 & iBlackhole & 1 & 46.1463 & TopSort & 1 & 0.034011 \\
			UR & 14 & iBlackhole & 1 & 1169.28 & TopSort & 1 & 0.47257 \\
			UR & 16 & iBlackhole & 0 & 1200 & TopSort & 1 & 3.05003 \\
			UR & 18 & iBlackhole & 0 & 1200 & TopSort & 1 & 14.5788 \\
			UR & 20 & iBlackhole & 0 & 1200 & TopSort & 1 & 67.7176 \\
			UR & 22 & iBlackhole & 0 & 1200 & TopSort & 1 & 373.244 \\
		\end{tabular}
	\end{center}
\end{table}

### 2019-12-08 blackholes/pavt/article.tex
Theorems for exclusion. Part 1

@@
 
-Let's consider a closure of a single blackhole root $r$. Let this closure to be acquired by depth-first search, which basically means, that
+\begin{definition}
+    Given graph $G(V, E)$, one of its roots $r \in Roots(G)$, topological sort of $Closure(r)$ $topSortOrder_r = TopologicalSort(Closure(r))$, and
+    ${|Closure(v)| \forall v \in Closure(r)}$ - closure sizes of all nodes reachable from $r$. We say that node $v$ is special under root $r$
+    $v \in Special_r \subseteq Closure(r)$, if its closure size $|Closure(v)|$ equals to its 1-based index in array $topSortOrder_r$.
+\end{definition}
+
+\begin{definition}
+    Given graph $G(V, E)$, one of its roots $r \in Roots(G)$, topological sort of $Closure(r)$ $topSortOrder_r = TopologicalSort(Closure(r))$,
+    $v \in Special_r$ and $u \in Closure(r)$. We say that $v$ dominates $u$ under root $r$, if  $v$ has larger index in $topSortOrder_r$.
+\end{definition}
+
+\begin{theorem}
+    Given graph $G(V,E)$, one of its roots $r \in Roots(G)$, and $v \in Special_r, u \in Closure(r)$.
+    If $v$ dominates $u$ under $r$ then $u$ is reachable from $v$.
+\end{theorem}
+\begin{proof}
+    By definition of special node $v$ has closure size equal to its index in $topSortOrder_r$. All the nodes of $Closure(v)$ has lesser indeces 
+    in $topSortOrder_r$ than $v$. Hence $\forall $u$ \in Closure(v)$ $u$ is reachable from $v$.
+\end{proof}
+
+\begin{theorem}
+   Given graph $G(V,E)$, one of its roots $r \in Roots(G)$, and $v \in Special_r$, $u \in Closure(r)$. 
+   If $v$ dominates $u$ under root $r$, then $Closure(v) \cup Closure(u) = Closure(v)$.
+\end{theorem}
+\begin{proof}
+    According to previous theorem, if $v$ dominates $u$ under root $r$, then $u$ is reachalbe from $v$. Then $u \in Closure(v)$,
+    which means that $Closure(u) \subset Closure(v)$.
+\end{proof}
+
+Let's consider a closure of a single global root $r$. Let this closure to be acquired by depth-first search, which basically means, that
 we can iterate by $i$ over these nodes in topologically sorted order $topsortOrder_r[i]$.

### 2019-12-08 blackholes/pavt/sketch.tex
Small updates

@@
 \end{table}
+
+\begin{table}[]
+	\caption{Configuration of experimental graphs}
+	\label{tabular:graphs}
+	\begin{center}
+		\begin{tabular}{c|c|c}
+			Graph Type & Nodes Count & Edges Count \\
+			\hline
+			RMAT.04 & 16 & 177 \\
+			RMAT.06 & 64 & 1270 \\
+			RMAT.08 & 256 & 6663 \\
+			RMAT.10 & 1024 & 30186 \\
+			RMAT.12 & 4096 & 126988 \\
+			RMAT.14 & 16384 & 517554 \\
+			RMAT.16 & 65536 & 2087072 \\
+			RMAT.18 & 262144 & 7892410 \\
+			RMAT.20 & 1048576 & 32993233 \\
+			RMAT.22 & 4194304 & 67099679 \\
+			SSCA2.04 & 16 & 120 \\
+			SSCA2.06 & 64 & 722 \\
+			SSCA2.08 & 256 & 2668 \\
+			SSCA2.10 & 1024 & 10529 \\
+			SSCA2.12 & 4096 & 43603 \\
+			SSCA2.14 & 16384 & 173477 \\
+			SSCA2.16 & 65536 & 677649 \\
+			SSCA2.18 & 262144 & 2724762 \\
+			SSCA2.20 & 1048576 & 10902287 \\
+			SSCA2.22 & 4194304 & 43580024 \\
+			UR.04 & 16 & 207 \\
+			UR.06 & 64 & 1614 \\
+			UR.08 & 256 & 7652 \\
+			UR.10 & 1024 & 32201 \\
+			UR.12 & 4096 & 130486 \\
+			UR.14 & 16384 & 523740 \\
+			UR.16 & 65536 & 2096570 \\
+			UR.18 & 262144 & 8388074 \\
+			UR.20 & 1048576 & 33553908 \\
+			UR.22 & 4194304 & 134217174
+		\end{tabular}
+	\end{center}
+\end{table}
+
+ In order to save time,
+the algorithms did not print out found blackholes, but printed counter on every found and not filtered (by any reason) blackhole.
+
+Please, mention, that "blackhole found" means that we have a set of vertex numbers in memory ready to be printed out and it's already validated.

### 2019-12-08 blackholes/pavt/article.tex
1st iteration is done

@@
 \section{Experimental Results}
-To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey \ref{alg:iblackhole} and TopSort ) on a 
-set of graphs.
 
-iBlackhole algorithm was implemented in accordance with the pseudocode in the original article. Divide and Conquer heuristic were not used for it.
-Our algorithm first applies condensation-based preprocessing and then uses BruteForce with partial reachability heuristic implemented via topsort.
-Both algorithms filter duplicates
+To demonstrate efficiency of our approach, we conducted a series of experiments for the iBlackhole and TopSort algorithms.
 
 For these experiments we chose Uniform Random (UR) \cite{random-uniform}, RMAT \cite{chakrabarti2004r},  and SSCA2 \cite{bader2005design} graph types. Those were generated for scales from 4 to 22
-with step 2.
-Size $i$ means that graph has $2^i$ nodes and $32*V$ edges.
-Actual count of edges to process can differ from these ideal values, as duplicate edges and self-loops are omitted when readinggraph from file.
-Configuration of the considered graphs is presented in Table \ref{tabular:graphs}.
-
-\begin{table}[]
-\caption{Configuration of experimental graphs}
-\label{tabular:graphs}
-\begin{center}
-\begin{tabular}{c|c|c}
-Graph Type & Nodes Count & Edges Count \\
-\hline
-RMAT.04 & 16 & 177 \\
-RMAT.06 & 64 & 1270 \\
-RMAT.08 & 256 & 6663 \\
-RMAT.10 & 1024 & 30186 \\
-RMAT.12 & 4096 & 126988 \\
-RMAT.14 & 16384 & 517554 \\
-RMAT.16 & 65536 & 2087072 \\
-RMAT.18 & 262144 & 7892410 \\
-RMAT.20 & 1048576 & 32993233 \\
-RMAT.22 & 4194304 & 67099679 \\
-SSCA2.04 & 16 & 120 \\
-SSCA2.06 & 64 & 722 \\
-SSCA2.08 & 256 & 2668 \\
-SSCA2.10 & 1024 & 10529 \\
-SSCA2.12 & 4096 & 43603 \\
-SSCA2.14 & 16384 & 173477 \\
-SSCA2.16 & 65536 & 677649 \\
-SSCA2.18 & 262144 & 2724762 \\
-SSCA2.20 & 1048576 & 10902287 \\
-SSCA2.22 & 4194304 & 43580024 \\
-UR.04 & 16 & 207 \\
-UR.06 & 64 & 1614 \\
-UR.08 & 256 & 7652 \\
-UR.10 & 1024 & 32201 \\
-UR.12 & 4096 & 130486 \\
-UR.14 & 16384 & 523740 \\
-UR.16 & 65536 & 2096570 \\
-UR.18 & 262144 & 8388074 \\
-UR.20 & 1048576 & 33553908 \\
-UR.22 & 4194304 & 134217174
-\end{tabular}
-\end{center}
-\end{table}
+with step 2. Size $i$ means that graph has $2^i$ nodes and approximately $32*V$ edges.
 
@@
 
-All runs were performed in a single-thread mode. We allowed up to 20 minutes of working time for every run. In order to save time,
-algorithms did not print out found blackholes, but printed counter on every found and not filtered (by any reason) blackhole.
+All runs were performed in a single-thread mode. We allowed up to 20 minutes of working time for every run. This time includes graph preprocessing for the TopSort algorithm.
 
-Please, mention, that "blackhole found" means that we have a set of vertex numbers in memory ready to be printed out and it's already validated.
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{rmat.pdf}
+	\caption{Algorithm results on RMAT graphs.}
+	\label{fig:rmat}
+\end{figure}
 
-As it is shown in the Tables \ref{tabular:tableresults}, \ref{tabular:tableresults2}, our TopSort approach to blackhole search shows better performance on small RMAT graphs.
-And it stays in advantage up to RMAT-22. Apparently, condensation preprocessing takes more time on larger graphs and increases time spent before finding
-the first blackhole. Nevertheless our algorithm shows significantly better performance in those cases when preprocessing saves more time. Also, we suppose
-that total working time on large graphs for our algorithm will be less, and this is the point for future survey.
+As it is shown in the Figures \ref{fig:rmat}, \ref{fig:ssca2} and \ref{fig:ur}, our TopSort approach to blackhole search shows better performance on small RMAT graphs and it stays in advantage up to RMAT-22.
+Apparently, condensation preprocessing takes more time on larger graphs and increases time spent before finding
+the first blackhole. Nevertheless, our algorithm shows significantly better performance in those cases when preprocessing saves more time. Also, we suppose
+that total working time on large graphs for our algorithm will be less, and this is the point for the future work.
 
 \FIXME{Это, вообще говоря, странно}
-SSCA2 graphs in our survey can be characterized with absolute absence of SCCs consisting of more than one node. Preprocessing here doesn't give any
-certain advantage, but only takes time. As we can see, on smaller graphs our approach beats iBlackhole algorithm, but slows down on larger graphs.
-We guess that the reason is less efficient order of search for our algorithm. One used in iBlackHole will faster print smaller balckholes, and later will try
+SSCA2 graphs in our survey can be characterized with absolute absence of SCCs consisting of more than one node. Preprocessing here does not allow to obtain any
+certain advantage, but only takes time. As we can see, on small graphs our approach beats the iBlackhole algorithm, but slows down on larger graphs.
+We guess that the reason is less efficient order of search for our algorithm. One used in iBlackHole will faster print smaller blackholes, and later will try
 to find larger ones, meanwhile our approach will check blackholes of all sizes.
 
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{ssca2.pdf}
+	\caption{Algorithm results on SSCA2 graphs.}
+	\label{fig:ssca2}
+\end{figure}
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{ur.pdf}
+	\caption{Algorithm results on Uniform Random graph.}
+	\label{fig:ur}
+\end{figure}
+
 Finally, we should point out, that our preprocessing dramatically speeds up blackhole search on the graphs, where no blackholes present at all. Brute force is
-avoided completely in this case, making calculation time impressively small in comparison to iBlackhole algorithm.
-
-\pagebreak
+avoided completely in this case, making calculation time impressively small in comparison to the iBlackhole algorithm.
 
@@
 
-In this paper we formulated and addressed some issues in application of already known iBlackhole algorithm. We introduced and tested two different ideas of optimization, combined into
-our TopSort algorithm. Those are: graph condensation preprocessing and topsort bruteforce optimization. We tested out approach on three different types of graphs and number of scales.
-In case of large SCCs we demostrate definite advantage over iBlackhole algorithm. Then, we can relatively fast detect total absence of blackholes, but sometimes quite costly preprocessing
-deminishes any boost of productivity and out performance degrades.
+In this paper we addressed some issues in application of the already known iBlackhole algorithm. We introduced and tested two different ideas of optimization, combined into
+our TopSort algorithm. Those are graph condensation preprocessing and topological sort of the brute-force optimization. We tested our approach on three different types of graphs and number of scales.
+
+In case of graph with large strong connected components we demostrate definite advantage over the iBlackhole algorithm.
+Then, we can relatively fast in the case of total blackhole absence, but sometimes quite costly preprocessing
+deminishes any performance boost and our advantage  degrades.
 \FIXME{можно написать в TODO}
@@
 
-\textit{Comment for the reviewer}. In the final paper revision we add parallel implementation of the TopSort algorithm and present much detailed evaluation on the larger graphs.
+\textit{Comment for the reviewer}. In the final paper revision we enhance our TopSort algorithm, add its parallel implementation and present much detailed evaluation on the larger graphs.
 
@@
 
-\begin{figure}[p]
-    \includegraphics[width=\linewidth]{rmat.pdf}
-    \caption{RMAT}
-    \label{fig:rmat}
-\end{figure}
-\begin{figure}[p]
-    \includegraphics[width=\linewidth]{ssca2.pdf}
-    \caption{SSCA2}
-    \label{fig:ssca2}
-\end{figure}
-\begin{figure}[p]
-    \includegraphics[width=\linewidth]{ur.pdf}
-    \caption{UR}
-    \label{fig:ur}
-\end{figure}
-
 \end{document}

### 2019-12-08 blackholes/pavt/article.tex
Small text changes + algo updated

--- 2019-12-08 blackholes/pavt/article.tex
@@
 \FIXME{не совсем понял логику в общем тексте}
-Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
+Of course, we could straightforward build the reachability matrix. But, it takes \FIXME{ Not sure $O(V^3)$ } time. Even though we decreased graph size
 in preprocessing step, it still can result in lots of computation. It is worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
@@
 
+\begin{remark}
+    Specialness of a node is decided independently under different roots. One node can be special for root $r1$ and non special for $r2$ at
+    the same time.
+\end{remark}
+
 \begin{definition}
@@
 \end{definition}
+
+\begin{remark}
+    If given two special nodes under the same root, ont of them is always dominated by the other.
+\end{remark}
 
@@
 
-Let's consider a closure of a single global root $r$. Let this closure to be acquired by depth-first search, which basically means, that
-we can iterate by $i$ over these nodes in topologically sorted order $topsortOrder_r[i]$.
-
-Also, while traversing this graph, we can calculate size of closure for each node
-in the given root closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
-conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder_r[i]$. We will call such nodes special.
-
-If there are two special nodes under the same root, union of their closures will be equal to the closure of topologically higher node.
-Hence, union of closures for any subset of special nodes under the same root will result into closure of the highest special node.
-
 %TODO: Осознать до конца
 Let's consider two nodes in the closure of a global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
-the special one should be topologically lower than non-special. Otherwise it would be one node basis.
-To summarize, if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.
+the special one should not dominate non-special. Otherwise non-special node can be omitted from basis.
+In general, if given a set of nodes, we shall remove all dominated ones in order for the rest to be a proper basis.
 
@@
 To find out, we take every global root of the condensated graph and find topological sort of its closure.
-Those vertices, which satisfy the rule above, we mark as special.
+Those vertices, which satisfy definition, we mark as special for according roots.
 
 \FIXME{зачем второй этап?}
-Second, we calculate how many special nodes for every root there is in the given set.
-If there is more than one special node for root, the $Cand$ set is called non-basis and will be omitted in the high-level TopSort blackhole mining Algorithm \ref{alg:topsort}.
-
 %\FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }
 
-Also we check if there are any non-special nodes lower than the special ones in the related root topological order, and if those found, then omit the node set.
+Second, if there are any dominated nodes (special or non-special) under any of the roots, we omit the node set.
 
-Third, when we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, we omit this set.
+Third, when we are out of cheap heuristics, we check if any of the candidate nodes are reachable from one another. If so, we omit this set.
 
@@
                             $Special_r = \emptyset$ \\
-                            $build\ topsortOrder_r$ \\
-                            \For{$i = 0; i<|topsortOrder_r|; i=i + 1$} {
-                                $v = topsortOrder_r[i]$ \\
+                            $build\ topSortOrder_r$ \\
+                            \For{$i = 0; i<|topSortOrder_r|; i=i + 1$} {
+                                $v = topSortOrder_r[i]$ \\
                                 $C_v = Closure(v)$ \\
@@
                                     $Special_r = Special_r \cup v$ \\
+                                    $maxSpecialIndex_r = max(maxSpecialIndex_r, i)$ \\
                                 }
@@
                         }
-			\tcc{(2) check for excluding specials} 
+			\tcc{(2) check for dominated nodes} 
+                        \ForEach{$r \in Roots(G)$} {
+                            $maxSpecialIndex_r = -1$ \\
+                        }
 			\ForEach{$v \in Cand$} {
@@
                                 \If{$v \in Special_r$} {
-                                    $Spec_r = Spec_r \cup v$ \\
-                                    \If{$|Spec_r| > 1$} {
+                                    $vTopSortIndex_r = arg_v (topSortOrder_r)$ \\
+                                    $maxSpecialIndex_r = max(maxSpecialIndex_r, vTopSortIndex_r)$ \\
+                                }
+                            }
+			}
+			\ForEach{$v \in Cand$} {
+                            \ForEach{$r \in Roots(G)$} {
+                                \If{$v \in Closure(r)$} {
+                                    $vTopSortIndex_r = arg_v (topSortOrder_r)$ \\
+                                    \If{$vTopSortIndex_r < maxSpecialIndex_r$} {
                                         \Return $False$ \\

### 2019-12-08 blackholes/pavt/article.tex
Start to edit Intro

--- 2019-12-08 blackholes/pavt/article.tex
@@
 %
-\title{Topological approach to finding blackholes in directed networks}
+\title{Topological Approach to Finding Blackholes in Directed Networks}
 %
-\titlerunning{Topological black holes mining}  % abbreviated title (for running head)
+\titlerunning{Topological Blackholes Mining}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 %
-\authorrunning{Ivanov, Semenov} % abbreviated author list (for running head)
+\authorrunning{D. Ivanov, A. Semenov} % abbreviated author list (for running head)
 %
@@
 \begin{abstract}
-In this paper we consider the problem of finding so called Blackhole pattern in directed unweighted graphs.
+In this paper we consider the problem of finding so called blackhole pattern in directed unweighted graphs.
 Firstly, we analyse already existing algorithms and approaches. Secondly, we describe how structure of a graph
 affects their efficiency. Then we introduce our approach to graph preprocessing. Finally, we describe topological sort based heuristic
-and provide results of experimental comparison between our approach and previously developed algorithm.
-\keywords{Directed networks $\cdot$ Subgraph mining}
+and provide results of experimental comparison between our approach and the previously developed algorithm.
+\keywords{Directed networks $\cdot$ Subgraph mining $\cdot$ Blackhole}
 \end{abstract}
@@
 \section{Blackhole Search in Directed Graph}
-The task of pattern mining in directed graphs finds its aplications in different areas. In 2010 Li et al.\cite{li2010mining} state
+The task of pattern mining in directed graphs finds its aplications in different areas. In 2010 Li et al.\cite{li2010detecting} state
 that governments are interested in detecting cases of financial fraud. They formulate the task for finding cases of illegal collaboration, and introduce dual patterns:
@@
 	\includegraphics[width=\linewidth]{rmat.pdf}
-	\caption{Algorithm results on RMAT graphs.}
+	\caption{Algorithm results on RMAT graphs}
 	\label{fig:rmat}
@@
 	\includegraphics[width=\linewidth]{ssca2.pdf}
-	\caption{Algorithm results on SSCA2 graphs.}
+	\caption{Algorithm results on SSCA2 graphs}
 	\label{fig:ssca2}
@@
 	\includegraphics[width=\linewidth]{ur.pdf}
-	\caption{Algorithm results on Uniform Random graph.}
+	\caption{Algorithm results on Uniform Random graph}
 	\label{fig:ur}

### 2019-12-08 blackholes/pavt/article.tex
Illustration for special nodes
Intro is almost done

@@
 
-\section{Blackhole Search in Directed Graph}
-The task of pattern mining in directed graphs finds its aplications in different areas. In 2010 Li et al.\cite{li2010detecting} state
-that governments are interested in detecting cases of financial fraud. They formulate the task for finding cases of illegal collaboration, and introduce dual patterns:
-blackhole and volcano. Those appear when a group of traders perform operations only between each other in order to manipulate market.
-Continuing study of different financial fraud casese, in 2017 Semenov et al. \cite{semenov2017survey} provided a survey of common approaches to anti - money laundering by graph mining.
-There are other applications. For example, the online detection of black holes/volcanos can timely 
-reflect anomalous events, such as disasters, catastrophic accidents, and therefore help keep public safety \cite{hong2015detecting}.
+\section{Introduction}
+
+One of the most relevant features of graphs representing real systems is community
+structure, or clustering. Detecting communities
+is of great importance in sociology, biology and computer science, disciplines where systems are
+often represented as graphs \cite{fortunato2010community}.
+
+In 2010 Li et al.\cite{li2010detecting} for the first time formulated a novel problem for finding blackhole and volcano patterns in a large directed graph.
+Specifically, a blackhole pattern is a group of graph nodes
+that there are only incoming edges to this group from the rest nodes in the graph. In contrast, a volcano pattern is a
+group which only has outgoing edges to the rest nodes in the graph.
+
+Both patterns can be observed in real world.
+For instance, in a trading network, a blackhole pattern may represent a group of traders who are
+manipulating the market. Also, blackhole and volcano patterns can represent the money laundering model \cite{semenov2017survey}.
+The online detection of blackholes and volcanos can timely 
+reflect anomalous events, such as disasters, catastrophic accidents, and therefore helps to keep public safety \cite{hong2015detecting}.
+
+The main contributor to the blackhole detecting problem is the authors from the University of New Jersey.
+They both formulated the task and proposed an algorithm for blackhole detection in case of unweighted directed graph \cite{li2010detecting}.
+Two years later they published an algorithm for approximate blackhole search in general case (weighted graph) \cite{li2012mining,li2014mining}. 
+Hong et al \cite{hong2015detecting} consider application for online blackhole mining in the city environment.
+They introduced an original algorithm an approximate solution of the problem for dynamic graphs.
+
+In this paper we consider simplified problem formulation, applied to a directed unweighted graph.
+
+\FIXME{переход к дальнейшему}
 
 \subsection{Problem Statement and Preliminaries}
-In this section we provide some basic notations and definitions, which will be used in this paper. In this paper we consider simplified problem formulation, applied to directed unweighted graph.
+In this section we provide some basic notations and definitions, which will be used in this paper.
 
-First, consider a directed graph $G = G(V,E)$, where $V$ is a set of all nodes and $E$ is a set if all edges of the graph. 
-\FIXME{Мне кажется, это лишнее предложение, так не пишут в статьях:} Assume that we deal with the general case of graph, without any special restrictions, unless the opposite is stated.
+First, consider a directed graph $G = G(V,E)$, where $V$ is a set of nodes and $E$ is a set if edges of the graph.
 
@@
 
-We will need the following statement to discuss issues of known approaches.
+We will need the following statement to discuss issues of the known approaches.
 \begin{lemma}
@@
 \end{proof}
-
-\subsection{Related Work}
-
-% Previously 
-%Authors from the University of New Jersey gave \cite{li2010detecting} general and simplified problem formulation for blackhole search
-in directed graph.
-
-The main contributor to the blackhole detecting problem is the authors from the University of New Jersey.
-They both formulated the task and provided an algorithm for blackhole detection in case of unweighted directed graph \cite{li2010detecting}.
-Two years later they published an algorithm for approximate blackhole search in general case (weighted graph) \cite{li2012mining,li2014mining}. 
-Hong et al \cite{hong2015detecting} had
-an unusual application for blackhole mining in the city environment. They introduced an original algorithm.
 
@@
 deminishes any performance boost and our advantage  degrades.
+
+%We originally evaluate blackhole algorithm Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which are quite a small size.
+
 \FIXME{можно написать в TODO}

### 2019-12-08 blackholes/pavt/article.tex
The paper is almost done

@@
 
-\FIXME{переход к дальнейшему}
-
-\subsection{Problem Statement and Preliminaries}
+\section{Preliminaries}
 In this section we provide some basic notations and definitions, which will be used in this paper.
 
+\subsection{Problem Statement} 
 First, consider a directed graph $G = G(V,E)$, where $V$ is a set of nodes and $E$ is a set if edges of the graph.
@@
 
+\subsection{Definitions}
 To support further discussion, we provide several more definitions.
@@
 
-We will need the following statement to discuss issues of the known approaches.
+\subsection{Known Issues}
+In this section we discuss applicability of the iBlackhole algorithm, proposed in \cite{li2010detecting}.
+The iBlackhole algorithm (see Algorithm \ref{alg:iblackhole}) is designed for searching blackholes of fixed size.
+Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which are quite
+a small size.
+
+We will need the following statement to discuss issues of the known approach.
 \begin{lemma}
@@
 
-\subsection{Known issues}
-In this section we shortly discuss applicability of the iBlackhole algorithm, proposed in \cite{li2010detecting}.
-The iBlackhole algorithm (see Algorithm \ref{alg:iblackhole}) is designed for searching blackholes of fixed size.
-Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which are quite
-a small size.
 Also, there are graphs, which has a small number of blackholes in relation to number of nodes. They consist
-of several large SCCs and relatively small number of independent leaves (or roots). Here arise some questions considering applicability of
+of several large SCCs and relatively small number of independent leaves (or roots). Here arise some questions considering applicability of the
 iBlackhole algorithm:
@@
 
-Let's consider the following example:
-
-Graph consisting of $(10^6)$ nodes, all of them form a single SCC. It means that there is only one blackhole -- the whole graph.
+Let's consider the following example. A graph is consist of $(10^6)$ nodes, all of them form a single SCC. It means that there is only one blackhole -- the whole graph.
 Unless we do not know anything about the graph structrure and do not have large enough cluster, we have one to a million chance
@@
                         }
-                        \tcc{Apply the brute-force algorithm to $C_(i)$ --  set of all possible subsets containing $i$ nodes}
+                        \tcc{Apply the brute-force algorithm to $C_i(i)$ --  set of all possible subsets of $C_i$ containing $i$ nodes}
                         \ForEach{$B \in C_i(i)$} {
@@
 			\label{alg:iblackhole}
-			\caption{iBlackHole}
+			\caption{iBlackhole}
 		\end{algorithm}
@@
 
-\FIXME{Мы реализовали этот алгоритм}
-
-\section{Algorithm Design}
+\section{Topological Algorithm Design}
 We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
-\FIXME{ограничение на суммарное время - да, 20 минут}
 The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
@@
 
-In this paper we use the algorithm of Sharir for finding graph condensation \cite{sharir1981strong}.
+In this work we use the algorithm of Sharir for finding graph condensation \cite{sharir1981strong}.
 
@@
 
-\FIXME{избежать путаницы между brute-force в разных местах статьи}
-
 Note, that a blackhole basis can consist of a single node or more than one node.
@@
 %TODO: осознать
-In the BruteForce algorithm we iterate over all possible unordered sets of nodes. 
+In the brute-force approach we iterate over all possible unordered sets of nodes. 
 Then for every node we acquire its closure. Union of all such closures is a potential blackhole.
@@
 
-With the BruteForce algorithm it is possible to find the same blackhole several times.
+With the brute-force approach it is possible to find the same blackhole several times.
 We aim to avoid at least some of the duplicates and introduce heuristic for it.
@@
 According to the definitions above if some node is not a root of a blackhole, it can be omitted, as set of roots uniquely identify a blackhole.
-Every non-root node in a blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.
-
-\FIXME{не совсем понял логику в общем тексте}
-Of course, we could straightforward build the reachability matrix. But, it takes \FIXME{ Not sure $O(V^3)$ } time. Even though we decreased graph size
+Every non-root node in a blackhole is reachable from one of the roots.
+Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.
+Of course, we could straightforward build the reachability matrix.
+But, it takes too much ($O(V^3)$) time.
+Even though we decreased graph size
 in preprocessing step, it still can result in lots of computation. It is worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
@@
 \begin{definition}
-    Given graph $G(V, E)$, one of its roots $r \in Roots(G)$, topological sort of $Closure(r)$ $topSortOrder_r = TopologicalSort(Closure(r))$, and
-    ${|Closure(v)| \forall v \in Closure(r)}$ - closure sizes of all nodes reachable from $r$. We say that node $v$ is special under root $r$
-    $v \in Special_r \subseteq Closure(r)$, if its closure size $|Closure(v)|$ equals to its 1-based index in array $topSortOrder_r$.
+	Given graph $G(V, E)$, a root $r \in Roots(G)$, $topSortOrder_r$ -- topological sort of $Closure(r)$ and 
+	$|Closure(v)|,\ \forall v \in Closure(r)$ --
+	closure sizes of all nodes reachable from $r$.
+We say that node $v$ is special under root $r$, if its closure size  equals to its 1-based index in array $topSortOrder_r$.
+Note that set of special nodes $Special_r \subseteq Closure(r)$.
 \end{definition}
@@
 \begin{remark}
-    Specialness of a node is decided independently under different roots. One node can be special for root $r1$ and non special for $r2$ at
-    the same time.
+    Specialness of a node is decided independently under different roots. One node can be special for root $r1$ and non-special for $r2$ at
+    the same time. It is illustrated in Figure \ref{fig:specialnodes}.
 \end{remark}
 
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{specialnodes.pdf}
+	\caption{Example of special nodes in a graph with two roots}
+	\label{fig:specialnodes}
+\end{figure}
+
 \begin{definition}
-    Given graph $G(V, E)$, one of its roots $r \in Roots(G)$, topological sort of $Closure(r)$ $topSortOrder_r = TopologicalSort(Closure(r))$,
-    $v \in Special_r$ and $u \in Closure(r)$. We say that $v$ dominates $u$ under root $r$, if  $v$ has larger index in $topSortOrder_r$.
+    Given graph $G(V, E)$, one of its roots $r \in Roots(G)$, $topSortOrder_r$ -- topological sort of $Closure(r)$,
+    $v \in Special_r$ and $u \in Closure(r)$.
+We say that $v$ dominates $u$ under root $r$, if $v$ has larger index than $u$ in $topSortOrder_r$.
+In other words, $v$ is on the right from $u$ in the $topSortOrder_r$ array.
 \end{definition}
@@
 \begin{remark}
-    If given two special nodes under the same root, ont of them is always dominated by the other.
+    If given two special nodes under the same root, one of them is always dominated by the other.
 \end{remark}
@@
     Given graph $G(V,E)$, one of its roots $r \in Roots(G)$, and $v \in Special_r, u \in Closure(r)$.
-    If $v$ dominates $u$ under $r$ then $u$ is reachable from $v$.
+    If $v$ dominates $u$ under $r$, then $u$ is reachable from $v$.
 \end{theorem}
+
+\FIXME{посмотреть еще раз внимательно}
 \begin{proof}
-    By definition of special node $v$ has closure size equal to its index in $topSortOrder_r$. All the nodes of $Closure(v)$ has lesser indeces 
+    By definition of special node $v$ has closure size equal to its index in $topSortOrder_r$. All the nodes of $Closure(v)$ has lesser indices 
     in $topSortOrder_r$ than $v$. Hence $\forall $u$ \in Closure(v)$ $u$ is reachable from $v$.
@@
 \begin{proof}
-    According to previous theorem, if $v$ dominates $u$ under root $r$, then $u$ is reachalbe from $v$. Then $u \in Closure(v)$,
+    According to the previous theorem, if $v$ dominates $u$ under root $r$, then $u$ is reachalbe from $v$. Then $u \in Closure(v)$,
     which means that $Closure(u) \subset Closure(v)$.
@@
 
-Third, when we are out of cheap heuristics, we check if any of the candidate nodes are reachable from one another. If so, we omit this set.
+Third, when we are out of cheap heuristic, we check if any of the candidate nodes are reachable from one another. If so, we omit this set in the high-level TopSort (Algorithm \ref{alg:topsort}).
 
@@
 
-Otherwise, we have a valid basis of a blackhole.
+The TopSort algorithm is based on looking over blackhole basises. Thus we have a valid basis of a blackhole.
 To acquire blackhole we should, according to definition of basis, unite all the closures of basic nodes.
@@
 
-Finally, we should point out, that our preprocessing dramatically speeds up blackhole search on the graphs, where no blackholes present at all. Brute force is
+Finally, we should point out, that our preprocessing dramatically speeds up blackhole search on the graphs, where no blackholes present at all. Brute-force is
 avoided completely in this case, making calculation time impressively small in comparison to the iBlackhole algorithm.
@@
 
-\FIXME{можно написать в TODO}
-If we check blackhole sizes consecutively, what is the expected time of the first blackhole found?
+%TODO: If we check blackhole sizes consecutively, what is the expected time of the first blackhole found?
 

### 2019-12-08 blackholes/pavt/article.tex
Smalll remarks and explanations.

@@
 Hong et al \cite{hong2015detecting} consider application for online blackhole mining in the city environment.
-They introduced an original algorithm an approximate solution of the problem for dynamic graphs.
+They introduced an original algorithm approximately solving the problem for dynamic graphs.
 
@@
 \subsection{Problem Statement} 
-First, consider a directed graph $G = G(V,E)$, where $V$ is a set of nodes and $E$ is a set if edges of the graph.
+First, consider a directed graph $G = G(V,E)$, where $V$ is a set of nodes and $E$ is a set of edges of the graph.
 
@@
 The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
-Decreased graph scale would definitely speed up the the calculation, as it can fall into a brute-force search in most cases.
+Decreased graph scale would definitely speed up the calculation, as it can fall into a brute-force search in most cases.
 The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in a restricted time.
@@
 Algorithm \ref{alg:isbasis} IsBasis tests a set of nodes as a blackhole basis.
+This algorthm takes a set of nodes as input. It returns $True$, if given set forms a proper basis,
+and  $False$ otherwise. This algortihm is used as a part of TopSort \ref{alg:topsort} algorithm, which is looking over
+blackhole basis candidates and sorts unproper ones out. If $IsBasis$ returns False on a basis candidate, this candidate is to be omitted.
+
+IsBasis algorithm does the following.
 First, we need to know what special nodes there are. This step can be precalculated. 
 \FIXME{как вычисляются global roots?}
+\begin{remark}
+    Those nodes of graph G(V,E) that have no incoming edges are called $Global\ roots$. 
+\end{remark}
 To find out, we take every global root of the condensated graph and find topological sort of its closure.
@@
 \FIXME{зачем второй этап?}
+\FIXME{Ответ: тут случаи отсеиваются за счет эвристики. Это дешево.}
 %\FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }
@@
 
-Third, when we are out of cheap heuristic, we check if any of the candidate nodes are reachable from one another. If so, we omit this set in the high-level TopSort (Algorithm \ref{alg:topsort}).
+Third, when we are out of cheap heuristics, we check if any of the candidate nodes are reachable from one another. 
+If so, we omit this set.
 
@@
 
-The TopSort algorithm is based on looking over blackhole basises. Thus we have a valid basis of a blackhole.
+If all previous checks did not return $False$ (omit), then $IsBasis$ returns $True$.
+Thus we return to TopSort algorithm and we have a valid basis of a blackhole.
 To acquire blackhole we should, according to definition of basis, unite all the closures of basic nodes.
@@
 Clearly, we still have to spend some computation time in order to decide, whether
-the candidate is worth checking or not, and we understand, that sometimes it can make process even more slower than before, but experimental results demostrate its efficiency.
+the candidate is worth checking or not, and we understand, that sometimes it can make process even more slower than before,
+but experimental results demostrate its efficiency.
 Our goal is to avoid this costly procedure of acquiring blackhole if possible.

### 2019-12-08 blackholes/pavt/article.tex
Typos

@@
 \begin{definition}
-Subset of nodes $B \in V$ is called a blackhole if  the following two conditions are satisfied: 
+Subset of nodes $B \subseteq V$ is called a blackhole if  the following two conditions are satisfied: 
 1) Subgraph $G'(B, E')$ induced by $B$ is weakly connected, 
@@
 For a given directed graph $G(V,E)$ if $v \in V$ is reachable from $u \in V$, then we say $u$ is a predecessor of $v$ and $v$ is a successor if $u$.
-If there is an edge from $u$ to $v$, then $u$ is a direct predecessor of $v$, and $v$ is a direct successor if $u$.
+If there is an edge from $u$ to $v$, then $u$ is a direct predecessor of $v$, and $v$ is a direct successor of $u$.
 \end{definition}
@@
 The iBlackhole algorithm (see Algorithm \ref{alg:iblackhole}) is designed for searching blackholes of fixed size.
-Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which are quite
+Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which is quite
 a small size.
@@
 	Given graph $G(V, E)$, a root $r \in Roots(G)$, $topSortOrder_r$ -- topological sort of $Closure(r)$ and 
-	$|Closure(v)|,\ \forall v \in Closure(r)$ --
+	$|Closure(v)|\ \forall v \in Closure(r)$ --
 	closure sizes of all nodes reachable from $r$.
@@
 \begin{remark}
-    Specialness of a node is decided independently under different roots. One node can be special for root $r1$ and non-special for $r2$ at
+    Specialness of a node is decided independently under different roots. One node can be special for root $r_1$ and non-special for $r_2$ at
     the same time. It is illustrated in Figure \ref{fig:specialnodes}.
@@
     By definition of special node $v$ has closure size equal to its index in $topSortOrder_r$. All the nodes of $Closure(v)$ has lesser indices 
-    in $topSortOrder_r$ than $v$. Hence $\forall $u$ \in Closure(v)$ $u$ is reachable from $v$.
+    in $topSortOrder_r$ than $v$. Hence $\forall u \in Closure(v)$ $u$ is reachable from $v$.
 \end{proof}

### 2019-12-08 blackholes/pavt/article.tex
The 1st version sent to the PCT Conference

@@
 %%%%%% DEBUG section %%%%%%
-\newcommand*{\DEBUG}{}
+%\newcommand*{\DEBUG}{}
 
@@
 affects their efficiency. Then we introduce our approach to graph preprocessing. Finally, we describe topological sort based heuristic
-and provide results of experimental comparison between our approach and the previously developed algorithm.
-\keywords{Directed networks $\cdot$ Subgraph mining $\cdot$ Blackhole}
+and present results of experimental comparison between our approach and the previously developed algorithm.
+\keywords{Directed networks $\cdot$ Subgraph mining $\cdot$ Blackhole pattern}
 \end{abstract}
@@
 \begin{proof}
-    According to the previous theorem, if $v$ dominates $u$ under root $r$, then $u$ is reachalbe from $v$. Then $u \in Closure(v)$,
+    According to the previous theorem, if $v$ dominates $u$ under root $r$, then $u$ is reachable from $v$. Then $u \in Closure(v)$,
     which means that $Closure(u) \subset Closure(v)$.
@@
 %TODO: Осознать до конца
+\begin{definition}
+	Those nodes of graph G(V,E) that have no incoming edges are called $Global\ roots$. 
+\end{definition}
+
 Let's consider two nodes in the closure of a global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
@@
 and  $False$ otherwise. This algortihm is used as a part of TopSort \ref{alg:topsort} algorithm, which is looking over
-blackhole basis candidates and sorts unproper ones out. If $IsBasis$ returns False on a basis candidate, this candidate is to be omitted.
+blackhole basis candidates and sorts unproper ones out. If $IsBasis$ returns False on a basis candidate, this candidate is omitted.
 
@@
 \FIXME{как вычисляются global roots?}
-\begin{remark}
-    Those nodes of graph G(V,E) that have no incoming edges are called $Global\ roots$. 
-\end{remark}
+
 To find out, we take every global root of the condensated graph and find topological sort of its closure.
@@
 If all previous checks did not return $False$ (omit), then $IsBasis$ returns $True$.
-Thus we return to TopSort algorithm and we have a valid basis of a blackhole.
+Thus we return to the TopSort algorithm and we have a valid basis of a blackhole.
 To acquire blackhole we should, according to definition of basis, unite all the closures of basic nodes.
@@
 the candidate is worth checking or not, and we understand, that sometimes it can make process even more slower than before,
-but experimental results demostrate its efficiency.
+but experimental results demonstrate its efficiency.
 Our goal is to avoid this costly procedure of acquiring blackhole if possible.
@@
 \FIXME{Это, вообще говоря, странно}
-SSCA2 graphs in our survey can be characterized with absolute absence of SCCs consisting of more than one node. Preprocessing here does not allow to obtain any
+SSCA2 graphs in our evaluation can be characterized with absolute absence of SCCs consisting of more than one node. Preprocessing here does not allow to obtain any
 certain advantage, but only takes time. As we can see, on small graphs our approach beats the iBlackhole algorithm, but slows down on larger graphs.
@@
 
-In this paper we addressed some issues in application of the already known iBlackhole algorithm. We introduced and tested two different ideas of optimization, combined into
-our TopSort algorithm. Those are graph condensation preprocessing and topological sort of the brute-force optimization. We tested our approach on three different types of graphs and number of scales.
+In this paper we addressed some issues in application of the already known iBlackhole algorithm. We proposed and tested two different ideas of optimization, combined into
+our TopSort algorithm.
+Those are graph condensation preprocessing and topological sort of the brute-force optimization.
+We tested our approach on three different types of graphs and number of scales.
+We originally consider the blackhole detection problem on graphs with number of nodes larger than 1 million.
 

### 2019-12-10 hpc_mod_and_tech/lab2/bg_old.py
Before HPC run

@@ -1,0 +1,17 @@
+
+
+if __name__ == "__main__":
+    proc = [(8, 16), (16, 16)]
+    grid = [8192, 16384, 32768]
+    for p in proc:
+        totalProc = p[0] * p[1]
+        for g in grid:
+            outfile = "outputs/bluegene." + str(totalProc) + "." + str(g)
+            errfile = "errors/bluegene." + str(totalProc) + "." + str(g)
+            params = "{0} {1} {2} {3} {4} {5} {6} {7}".format(g, g, p[0], p[1], 20, 1.0, 1.0, 1)
+            if totalProc == 128:
+                mins = 15
+            else:
+                mins = 10
+            cmd = "mpisubmit.pl -n {0} -m smp -w {4}:00 --stderr {1} --stdout {2} ./bluegene -- {3}".format(totalProc, outfile, errfile, params, mins)
+            print(cmd)

### 2019-12-10 hpc_mod_and_tech/lab2/bluegene.py

@@ -1,0 +1,18 @@
+
+
+if __name__ == "__main__":
+    proc = [(8, 16), (16, 16)]
+    grid = [8192, 16384, 32768]
+    for ompNumProc in range(1, 4):
+        for p in proc:
+            totalProc = p[0] * p[1]
+            for g in grid:
+                outfile = "outputs/bluegene." + str(totalProc) + "." + str(g)
+                errfile = "errors/bluegene." + str(totalProc) + "." + str(g)
+                params = "{0} {1} {2} {3} {4} {5} {6} {7}".format(g, g, p[0], p[1], 20, 1.0, 1.0, ompNumProc)
+                if totalProc == 128:
+                    mins = 15
+                else:
+                    mins = 10
+                cmd = "mpisubmit.pl -n {0} -m smp -w {4}:00 --stderr {1} --stdout {2} ./bluegene -- {3}".format(totalProc, outfile, errfile, params, mins)
+                print(cmd)

### 2019-12-10 hpc_mod_and_tech/lab2/main.cpp

@@ -1,8 +1,12 @@
+#define _USE_MATH_DEFINES
+#include <cmath>
+
 #include <cstdlib>
 #include <iostream>
 #include <vector>
 #include <string>
 #include <mpi.h>
+#include <omp.h>
 
 using namespace std;
 
@@ -30,9 +34,9 @@ public:
     Calc(int localYSize, double hx, double hy, double ht,  int i, int j) : hx(hx), hy(hy), ht(ht), i(i), j(j) {
         int storedYSize = localYSize + 2; 
 
-        tl = storedYSize * (i - 1) + j;
-        t = storedYSize * (i - 1) + 1 + j;
-        tr = storedYSize * (i - 1) + 1 + j + 1;
+        tl = storedYSize * (i) + j;
+        t = storedYSize * (i) + 1 + j;
+        tr = storedYSize * (i) + 1 + j + 1;
 
         l = storedYSize * (i + 1) + j;
         m = storedYSize * (i + 1) + 1 + j;
@@ -46,33 +50,15 @@ public:
     void CalcNextV1(double* next1, double* v1, double* v2) {
         double laplas1 = Laplas(v1);      
         double g1 = Grad1(v1, v2);
-        next1[m] = ht * (v1[m] + laplas1) + g1;
+        next1[m] = ht * (g1 + laplas1) + v1[m];
     }
 
     void CalcNextV2(double* next2, double* v1, double* v2) {
         double laplas2 = Laplas(v2);
         double g2 = Grad2(v1, v2);
-        next2[m] = ht * (v2[m] + laplas2) + g2;
+        next2[m] = ht * (g2 + laplas2) + v2[m];
     }
 };
-
-// t = 0 Initial condition
-double Phi1(double x, double y) {
-   return x + y;
-}
-
-// t = 0 Initial condition
-double Phi2(double x, double y) {
-   return x + y;
-}
-
-// var 10
-// v1(x = 0, x = xmax) = 0
-// v1(y) periodic
-// v2(x) 2nd
-// v2(y) 2nd
-
-
 
 class Solution {
 private:
@@ -81,7 +67,7 @@ private:
     // solution grid size 
     int gridN, gridM, gridT;
     // constraints
-    double xmax, ymax, tmax;
+    double xmax, ymax;
     // grid step
     double hx, hy, ht;
     // mpi params
@@ -97,6 +83,7 @@ private:
     double* recvBuf;
 
     int Neighbor[8];
+    int Tag[8];
     int CommBufShift[8];
     int CommBufSize[8];
 
[…]

### 2019-12-10 hpc_mod_and_tech/lab2/polus.py

@@ -1,0 +1,13 @@
+
+
+if __name__ == "__main__":
+    proc = [(1, 1), (2, 5), (4, 5), (5, 8)]
+    grid = [2048, 4096, 8192]
+    for p in proc:
+        totalProc = p[0] * p[1]
+        for g in grid:
+            outfile = "outputs/polus." + str(totalProc) + "." + str(g)
+            errfile = "errors/polus." + str(totalProc) + "." + str(g)
+            params = "{0} {1} {2} {3} {4} {5} {6} {7}".format(g, g, p[0], p[1], 20, 1.0, 1.0, 1)
+            cmd = "mpisubmit.pl -p {0} -w 60:00 --stderr {1} --stdout {2} ./polus -- {3}".format(totalProc, outfile, errfile, params)
+            print(cmd)

### 2019-12-10 hpc_mod_and_tech/lab2/main.cpp
HPC Lab2 fixed

@@ -96,7 +96,7 @@ private:
 
     void Init(int argc, char** argv) {
         if (argc < 9) {
-            cerr << "use ./solution mpiN mpiM gridN gridM gridT xmax ymax printDebug" << endl;
+            cerr << "use ./solution mpiN mpiM gridN gridM gridT xmax ymax ompNumThreads printDebug" << endl;
             return;
         }
         // mpi grid size
@@ -134,28 +134,29 @@ private:
             return;
         }
 
-        mpiX = mpiRank / mpiN;
-        mpiY = mpiRank % mpiN;
+        mpiX = mpiRank / mpiM;
+        mpiY = mpiRank % mpiM;
+
+        //printf("rank %d mpiX %d mpiY %d\n", mpiRank, mpiX, mpiY);
 
         bool isXMax = mpiX == mpiN - 1;  
         bool isYMax = mpiY == mpiM - 1;
 
-        localXSize = isXMax ? gridN % mpiN : gridN / mpiN;
-        if (localXSize == 0) {
-            localXSize = gridN / mpiN;
-        }
-        localYSize = isYMax ? gridM % mpiM : gridM / mpiM;
-        if (localYSize == 0) {
-            localYSize = gridM / mpiM;
-        }
+        int maxLocalXSize = gridN / mpiN;
+        localXSize = isXMax ? maxLocalXSize + gridN % maxLocalXSize : maxLocalXSize;
+
+        int maxLocalYSize = gridM / mpiM;
+        localYSize = isYMax ? maxLocalYSize + gridM % maxLocalYSize : maxLocalYSize;
+
+        //printf("rank %d localXSize %d localYSize %d\n", mpiRank, localXSize, localYSize);
         storedXSize = localXSize + 2;
         storedYSize = localYSize + 2;
 
         sliceSize = storedXSize * storedYSize;
         commBufSize = 2 * (localXSize + localYSize) + 4;
 
-        xStart = mpiX * localXSize;
-        yStart = mpiY * localYSize;
+        xStart = mpiX * maxLocalXSize;
+        yStart = mpiY * maxLocalYSize;
 
         gridT += 1;
         v1.reserve(gridT);
@@ -181,6 +182,18 @@ private:
         Neighbor[5] = GetDest(-1, 1);
         Neighbor[6] = GetDest(1, -1);
         Neighbor[7] = GetDest(1, 1);
+
+        /*printf("rank %d neigbors %d %d %d %d %d %d %d %d\n",
[…]

### 2019-12-11 hpc_mod_and_tech/lab2/bluegene.py
Memory optimization for bluegene 512

@@ -1,18 +1,29 @@
-
+import sys 
+import os
 
 if __name__ == "__main__":
-    proc = [(8, 16), (16, 16)]
+    
+    submit = False
+    if len(sys.argv) == 2:
+        submit = sys.argv[1] == "submit"
+
+    proc = [(8, 16), (16, 16), (16, 32)]
     grid = [8192, 16384, 32768]
-    for ompNumProc in range(1, 4):
+    for ompNumProc in range(1, 4): 
         for p in proc:
             totalProc = p[0] * p[1]
             for g in grid:
-                outfile = "outputs/bluegene." + str(totalProc) + "." + str(g)
-                errfile = "errors/bluegene." + str(totalProc) + "." + str(g)
-                params = "{0} {1} {2} {3} {4} {5} {6} {7}".format(g, g, p[0], p[1], 20, 1.0, 1.0, ompNumProc)
+                filename = "bluegene." + str(totalProc) + "." + str(g) + "." + str(ompNumProc)
+                errfile = "errors/" + filename
+                outfile = "outputs/" + filename
+                params = str(p[0]) + " " + str(p[1]) + " " + str(g) + " " + str(g) + " " + str(20) + " " + str(1.0) + " " + str(1.0) + " " + str(ompNumProc)
                 if totalProc == 128:
                     mins = 15
-                else:
+                if totalProc == 256:
                     mins = 10
-                cmd = "mpisubmit.pl -n {0} -m smp -w {4}:00 --stderr {1} --stdout {2} ./bluegene -- {3}".format(totalProc, outfile, errfile, params, mins)
+                if totalProc == 512:
+                    mins = 5 
+                cmd = "mpisubmit.bg -n " + str(totalProc) + " -m smp -w " + str(mins) + ":00 --stdout " + outfile + " --stderr " + errfile + " ./bluegene -- " + params
                 print(cmd)
+                if submit:
+                    os.system(cmd)

### 2019-12-11 hpc_mod_and_tech/lab2/main.cpp

@@ -107,6 +107,7 @@ private:
         gridN = atoi(argv[3]);
         gridM = atoi(argv[4]);
         gridT = atoi(argv[5]);
+        gridT += 1;
 
         // constraints
         xmax = atof(argv[6]);
@@ -137,8 +138,6 @@ private:
         mpiX = mpiRank / mpiM;
         mpiY = mpiRank % mpiM;
 
-        //printf("rank %d mpiX %d mpiY %d\n", mpiRank, mpiX, mpiY);
-
         bool isXMax = mpiX == mpiN - 1;  
         bool isYMax = mpiY == mpiM - 1;
 
@@ -148,7 +147,6 @@ private:
         int maxLocalYSize = gridM / mpiM;
         localYSize = isYMax ? maxLocalYSize + gridM % maxLocalYSize : maxLocalYSize;
 
-        //printf("rank %d localXSize %d localYSize %d\n", mpiRank, localXSize, localYSize);
         storedXSize = localXSize + 2;
         storedYSize = localYSize + 2;
 
@@ -158,11 +156,10 @@ private:
         xStart = mpiX * maxLocalXSize;
         yStart = mpiY * maxLocalYSize;
 
-        gridT += 1;
-        v1.reserve(gridT);
-        v2.reserve(gridT);
+        v1.reserve(2);
+        v2.reserve(2);
         
-        for (int i = 0; i < gridT; i++) {
+        for (int i = 0; i < 2; i++) {
             v1.push_back(new double[sliceSize]);
             v2.push_back(new double[sliceSize]);
         }
@@ -182,18 +179,6 @@ private:
         Neighbor[5] = GetDest(-1, 1);
         Neighbor[6] = GetDest(1, -1);
         Neighbor[7] = GetDest(1, 1);
-
-        /*printf("rank %d neigbors %d %d %d %d %d %d %d %d\n",
-            mpiRank,
-            Neighbor[0],
-            Neighbor[1],
-            Neighbor[2],
-            Neighbor[3],
-            Neighbor[4],
-            Neighbor[5],
-            Neighbor[6],
-            Neighbor[7]
-        ); */
 
         Tag[0] = 1;
         Tag[1] = 0;
@@ -291,14 +276,12 @@ private:
         while (res >= mod) {
             res -= mod;
         }
-        //printf("rank %d val %d d %d mod %d\n", mpiRank, val, d, mod);
         return res;
     }
 
[…]

### 2019-12-11 hpc_mod_and_tech/lab2/report/report.tex
Report in tex

\documentclass[12pt,a4paper]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{physics}

\begin{document}

\begin{titlepage}

    \begin{figure}[h]
        \centering
        \includegraphics[scale=0.5]{gzlogo_new.png}
    \end{figure}

    \begin{center}
        \large
        Московский государственный университет имени М.В. Ломоносова

        \vfill

        \LARGE
        \textbf{Задание 2. Численное решение параболическоого уравнения}

        \vfill

        \hfill\begin{minipage}{0.6\textwidth}
            \normalsize
            \textbf{Факультет:} Вычислительной математики и кабернетики \\
            \textbf{Кафедра:} Суперкомпьютеров и квантовой информатики \\
            \textbf{Группа:} М-218 \\
            \textbf{Студент:} Иванов Денис Евгеньевич \\
            \textbf{Предмет:} Суперкомпьютерное Моделирование и Технологии \\
        \end{minipage}
        \vfill

    \end{center}

    \begin{center}
        \large
        Москва, 2019
    \end{center}
\end{titlepage}

\Large
Математическая постановка дифференциальной задачи

\normalsize
В двумерной замкнутой области:

\begin{equation}
    \Omega = [0 \leq x \leq L_x] \times [0 \leq y \leq L_y]
\end{equation}

Для $0 < t \leq T$ требуется найти решение $u(x,y,t)$ равнения в частных производных

\begin{equation}
    \label{eq:derivative}
    \frac{\partial u}{\partial t} = \Delta u + \nabla div u
\end{equation}

Заметим, что функция $u$ - это вектор функция $u=(u_1, u_2)$ поэтому уравнения ()
можно записать в виде

\begin{equation}
    \begin{cases} 
    \label{eq:decompose}
        \frac{\partial{u_1}}{\partial t} = 2 * \pdv[2]{u_1}{x} + \pdv[2]{u_1}{y} + \pdv{u_2}{x}{y} \\ 
        \frac{\partial{u_2}}{\partial t} = \pdv[2]{u_2}{x} + 2 * \pdv[2]{u_2}{y} + \pdv{u_1}{x}{y} \\ 
    \end{cases}
\end{equation}

Дополним систему (\ref{eq:derivative}) начальными условиями

\begin{equation}
    u|_{t = 0} = \phi(x, y)
\end{equation}

На границах области заданы граничные условия первого рода, второго рода или периодические
\newpage

\Large
Численный метод решения задачи

\normalsize
Для численного решения задачи введем на области $\Omega$ равномерную решетку
$\Omega_h = \bar{\omega_h} \times \omega_{\tau}$, где

\begin{equation}
\begin{array}{lcl}
    \bar{\omega_{h}} & = & \{u(x_i = ih_x, jh_y),i,j = 0,1,...,N, h_xN = L_x, h_y N = L_y\} \\
    \omega_{\tau} & = & \{t_n = n\tau, n = 0,1,...,K, \tau K = T\} \\
\end{array}
\end{equation}

 - множество внутренних узлов
 - множество граничных узлов сетки.

    Для аппроксимации исходного уравнения (1) с однородными граничными условиями и начальными условиями (2) воспользуемся следующей системой уравнений

Здесь через .... обозначен дискретный аналог дифференциального решения ... . А запись
... означает значение численного решения в точке ... .

 - пятиточечный разностный аналог оператора Лапласа:

... - девятиточечный аналог оператора ... :

приведенная выше разностная схема является явной - значение ... на ... шаге по времени явным образом выразить через значения на предыдущих слоях

Для параболических задач такое соотношение имеет вид .... , где ... - константа, которая вычисляется аналитически или подбирается экспериментально.

Для начала счета используем условие (2)

 В случае однородных граничных условий первого рода:

В случае однородных условий второго рода, используя более точную аппроксимацию (пример производной по переменной ... )

Разностная аппроксимация для периодических граничных условий выглядит следующим образом:

\end{document}

### 2019-12-11 hpc_mod_and_tech/lab2/report/report.tex
Task statement and numerical solution

@@
 
-\Large
-Математическая постановка дифференциальной задачи
+\section{Математическая постановка дифференциальной задачи}
 
-\normalsize
 В двумерной замкнутой области:
@@
 
-Заметим, что функция $u$ - это вектор функция $u=(u_1, u_2)$ поэтому уравнения ()
+Заметим, что функция $u$ - это вектор функция $u=(u_1, u_2)$ поэтому уравнения (\ref{eq:derivative})
 можно записать в виде
@@
 
-Дополним систему (\ref{eq:derivative}) начальными условиями
+Дополним систему (\ref{eq:decompose}) начальными условиями
 
 \begin{equation}
+    \label{eq:initialcondition}
     u|_{t = 0} = \phi(x, y)
@@
 
-\Large
-Численный метод решения задачи
+\section{Численный метод решения задачи}
 
-\normalsize
 Для численного решения задачи введем на области $\Omega$ равномерную решетку
@@
 \begin{array}{lcl}
-    \bar{\omega_{h}} & = & \{u(x_i = ih_x, jh_y),i,j = 0,1,...,N, h_xN = L_x, h_y N = L_y\} \\
+    \bar{\omega_{h}} & = & \{u(x_i = ih_x, y_j = jh_y),i,j = 0,1,...,N, h_xN = L_x, h_y N = L_y\} \\
     \omega_{\tau} & = & \{t_n = n\tau, n = 0,1,...,K, \tau K = T\} \\
@@
 
- - множество внутренних узлов
- - множество граничных узлов сетки.
+$\omega_h$ - множество внутренних узлов
 
-    Для аппроксимации исходного уравнения (1) с однородными граничными условиями и начальными условиями (2) воспользуемся следующей системой уравнений
+$\omega_{\gamma}$ - множество граничных узлов сетки.
 
-Здесь через .... обозначен дискретный аналог дифференциального решения ... . А запись
-... означает значение численного решения в точке ... .
+    Для аппроксимации исходного уравнения (\ref{eq:decompose}) с однородными граничными условиями и начальными условиями (\ref{eq:initialcondition}) воспользуемся следующей системой уравнений
 
- - пятиточечный разностный аналог оператора Лапласа:
+\begin{equation}
+    \label{eq:system}
+        \frac{v_{ij}^{n+1} + v_{ij}^n}{\tau} = \Delta^hv^n + \nabla^h{div}^hv^n, (x_i, y_j) \in \omega_h, n = 1,2,...,K-1
+\end{equation}
 
-... - девятиточечный аналог оператора ... :
+Здесь через $v = (v_1, v_2)$ обозначен дискретный аналог дифференциального решения $u = (u_1, u_2)$ . А запись
+$v_{ij}^n$ означает значение численного решения в точке $(x_i,y_j,t_n) \in \bar{\omega_h} \times \omega_{\tau}$.
 
-приведенная выше разностная схема является явной - значение ... на ... шаге по времени явным образом выразить через значения на предыдущих слоях
+$\Delta^h$ - пятиточечный разностный аналог оператора Лапласа:
 
-Для параболических задач такое соотношение имеет вид .... , где ... - константа, которая вычисляется аналитически или подбирается экспериментально.
+\begin{equation}
+    \label{eq:laplas}
+        \Delta^hv^n = \frac{v_{i-1j}^n - 2v_{ij}^n + v_{i+1j}^n}{h_x^2} + \frac{v_{ij-1}^n - 2v_{ij}^n + v_{ij+1}^n}{h_y^2}
+\end{equation}
 
-Для начала счета используем условие (2)
+$\nabla^hdiv^h$ - девятиточечный аналог оператора $\nabla div$:
+\begin{equation}
+    \label{eq:nabladiv}
+        \nabla^hdiv^hv^n  = 
+        \left(
+            \begin{array}{lcr} 
+                \frac{v1_{i-1j}^n - 2v1_{ij}^n + v1_{i+1j}^n}{h_x^2} & + & \frac{v2_{i+1j+1}^n - v2_{i+1j-1}^n - v2_{i-1j+1}^n + v2_{i-1j-1}^n}{4h_xh_y} \\
+                \frac{v2_{ij-1}^n - 2v2_{ij}^n + v2_{ij+1}^n}{h_y^2} & + & \frac{v2_{i+1j+1}^n - v2_{i+1j-1}^n - v2_{i-1j+1}^n + v2_{i-1j-1}^n}{4h_xh_y} 
+     \end{array}\right)
+\end{equation}
+
+приведенная выше разностная схема является явной - значение $v_{ij}^{n+1}$ на $(n+1)$ шаге по времени можно явным образом выразить через значения на предыдущих слоях
+
+Для параболических задач такое соотношение имеет вид $\tau < \gamma(h_x^2, h_y^2)$ , где $\gamma$ - константа, которая вычисляется аналитически или подбирается экспериментально.
+
+Для начала счета используем условие (\ref{eq:initialcondition}):
+
+\begin{equation}
+    \label{eq:begincalc}
+        v_{ij}^0 = \phi(x_i,y_j), (x_i,y_j) \in \omega_h
+\end{equation}
 
@@
 
-В случае однородных условий второго рода, используя более точную аппроксимацию (пример производной по переменной ... )
+\begin{equation}
+    \label{eq:secondtype}
+        v_{ij}^n = 0, (x_i,y_j) \in \gamma_h
+\end{equation}
+
+В случае однородных условий второго рода, используем более точную аппроксимацию (пример производной по переменной $x$):
+
+\begin{equation}
+    \label{eq:consistentsecond}
+    \begin{array}{lcl}
+        \frac{\partial v}{\partial x}|_{x = 0} & \approx & \frac{-3v_{0j} + 4v_{1j} - v_{2j}}{2h_x}, (x_i, y_j) \in \omega_h \\
+        \frac{\partial v}{\partial x}|_{x = L_x} & \approx & \frac{v_{N-2j} - 4v{N-1j} + 3v_{Nj}}{2h_x}, (x_i, y_j) \in \omega_h \\
+    \end{array}
+\end{equation}
 
@@
 
+\begin{equation}
+    \label{eq:periodicapprox}
+    \begin{array}{lcl}
+        v_{0j}^{n+1} = v_{Nj}^{n+1} & , & v_{1j}^{n+1} = v_{N+1j}^{n+1},j=0,1,...,N \\
+        v_{i0}^{n+1} = v_{iN}^{n+1} & , & v_{i1}^{n+1} = v_{iN+1}^{n+1},j=0,1,...,N \\
+    \end{array}
+\end{equation}
+
 \end{document}

### 2019-12-11 hpc_mod_and_tech/lab2/report/report.tex
results and report v1

@@
 
+\section{Описание реализации}
+
+Решение задачи было реализовано на языке С++ с использованием библиотек MPI и OpenMP.
+
+Узлы суперкомпьютера расположены в порядке прямоугольной сетки. Эта сетка накладывается на расчетную сетку,
+и каждый узел получает соответствующий участок. Если по какой-то из координат $x,y$ расчетная сетка не делится нацело
+на количество узлов, то остаток отдается последнему узлу по данной координате. Это создает незначительный дисбаланс нагрузки.
+
+Каждый узел хранит в памяти свой расчетный участок и дополнительный слой элементов толщиной 1, т.н. гало. Гало - пограничные элементы,
+рассчитывается соседними узлами и передается на текущий.
+Таким образом, у данного узла есть достаточно данных, чтобы рассчитать свою область на текущей итерации.
+
+Для реализации циклических условий, расчетная область расширяется таким образом, чтобы можно было вычислить крайний элемент $N$.
+Далее этот элемент пересылается на противоположную сторону сетки и помещается в индекс $0$. И, наоборот, первый элемент сетки
+пересылается, чтобы расположиться в $N+1$ позиции. Т.к. крайнему узлу не требуется гало с той стороны, где сетка заканчивается,
+то гало ячейки используются для расширения сетки.
+
+OpenMP был сразу же включен в программный код. В данной работе считаем, что запуск OpenMP в один поток - это запуск без OpenMP.
+За количество потоков отвечает отдельный параметр.
+
+Программа запускается командой
+
+\begin{center}
+\small
+\textit{mpirun -n X ./solution mpiN mpiM gridN gridM gridT xmax ymax ompNumThreads}
+\end{center}
+
+Где:
+\begin{itemize}
+    \item X - число MPI узлов
+    \item mpiN, mpiM - размер сетки MPI
+    \item gridN, gridM - размер вычислительной сетки
+    \item gridT - количество шагов по времени
+    \item xmax, ymax - ограничения по $x, y$
+    \item ompNumThreads - число потоков OpenMP
+\end{itemize}
+
+\newpage
+
+\section{Экспериментальные результаты}
+
+\begin{center}
+\begin{table}[h]
+\begin{tabular}{l|lllllll}
+\label{tabular:bluegene}
+HPC & Nodes & Grid & OMP & Wall time & Delta & Speedup & Efficiency \\
+\hline
+bluegene & 128 & 8192 & 1 & 4.96e+00 & 8.24e-08 & 1.00e+00 & 1.00e+00 \\
+bluegene & 128 & 8192 & 2 & 2.49e+00 & 8.24e-08 & 1.99e+00 & 9.94e-01 \\
+bluegene & 128 & 8192 & 3 & 1.67e+00 & 8.24e-08 & 2.96e+00 & 9.87e-01 \\
+bluegene & 128 & 16384 & 1 & 2.09e+01 & 2.06e-08 & 1.00e+00 & 1.00e+00 \\
+bluegene & 128 & 16384 & 2 & 1.05e+01 & 2.06e-08 & 2.00e+00 & 9.98e-01 \\
+bluegene & 128 & 16384 & 3 & 6.98e+00 & 2.06e-08 & 2.99e+00 & 9.98e-01 \\
+bluegene & 128 & 32768 & 1 & 8.32e+01 & 5.15e-09 & 1.00e+00 & 1.00e+00 \\
+bluegene & 128 & 32768 & 2 & 4.17e+01 & 5.15e-09 & 1.99e+00 & 9.97e-01 \\
+bluegene & 128 & 32768 & 3 & 2.79e+01 & 5.15e-09 & 2.99e+00 & 9.95e-01 \\
+bluegene & 256 & 8192 & 1 & 2.48e+00 & 8.24e-08 & 2.00e+00 & 9.98e-01 \\
+bluegene & 256 & 8192 & 2 & 1.25e+00 & 8.24e-08 & 3.95e+00 & 9.88e-01 \\
+bluegene & 256 & 8192 & 3 & 8.45e-01 & 8.24e-08 & 5.86e+00 & 9.77e-01 \\
+bluegene & 256 & 16384 & 1 & 1.04e+01 & 2.06e-08 & 2.00e+00 & 9.99e-01 \\
+bluegene & 256 & 16384 & 2 & 5.23e+00 & 2.06e-08 & 3.99e+00 & 9.97e-01 \\
+bluegene & 256 & 16384 & 3 & 3.50e+00 & 2.06e-08 & 5.96e+00 & 9.94e-01 \\
+bluegene & 256 & 32768 & 1 & 4.17e+01 & 5.15e-09 & 2.00e+00 & 9.98e-01 \\
+bluegene & 256 & 32768 & 2 & 2.12e+01 & 5.15e-09 & 3.93e+00 & 9.83e-01 \\
+bluegene & 256 & 32768 & 3 & 1.42e+01 & 5.15e-09 & 5.88e+00 & 9.80e-01 \\
+bluegene & 512 & 8192 & 1 & 1.24e+00 & 8.24e-08 & 3.99e+00 & 9.96e-01 \\
+bluegene & 512 & 8192 & 2 & 6.36e-01 & 8.24e-08 & 7.80e+00 & 9.75e-01 \\
+bluegene & 512 & 8192 & 3 & 4.33e-01 & 8.24e-08 & 1.14e+01 & 9.54e-01 \\
+bluegene & 512 & 16384 & 1 & 5.22e+00 & 2.06e-08 & 4.00e+00 & 1.00e+00 \\
+bluegene & 512 & 16384 & 2 & 2.63e+00 & 2.06e-08 & 7.94e+00 & 9.93e-01 \\
+bluegene & 512 & 16384 & 3 & 1.76e+00 & 2.06e-08 & 1.19e+01 & 9.88e-01 \\
+bluegene & 512 & 32768 & 1 & 2.08e+01 & 5.15e-09 & 3.99e+00 & 9.98e-01 \\
+bluegene & 512 & 32768 & 2 & 1.05e+01 & 5.15e-09 & 7.92e+00 & 9.90e-01 \\
+bluegene & 512 & 32768 & 3 & 7.03e+00 & 5.15e-09 & 1.18e+01 & 9.87e-01 \\
+\hline
+\end{tabular}
+\caption{Результаты запусков на Bluegene.}
+\end{table}
+\end{center}
+
+\newpage
+
+\begin{center}
+\begin{table}[h]
+\begin{tabular}{l|lllllll}
+\label{tabular:polus}
+
+HPC & Nodes & Grid & OMP & Wall time & Delta & Speedup & Efficiency \\
+\hline
+polus & 1 & 2048 & 1 & 8.797578 & 1.31902e-06 & 1.0 & 1.0 \\
+polus & 1 & 4096 & 1 & 33.57414 & 3.295944e-07 & 1.0 & 1.0 \\
+polus & 1 & 8192 & 1 & 79.26881 & 8.23785e-08 & 1.0 & 1.0 \\
+polus & 10 & 2048 & 1 & 29.14442 & 1.31902e-06 & 0.301861488408 & 0.0301861488408 \\
+polus & 10 & 4096 & 1 & 20.4802 & 3.295944e-07 & 1.63934629545 & 0.163934629545 \\
+polus & 10 & 8192 & 1 & 61.29066 & 8.23785e-08 & 1.29332609569 & 0.129332609569 \\
+polus & 20 & 2048 & 1 & 14.50379 & 1.31902e-06 & 0.606570972139 & 0.030328548607 \\
+polus & 20 & 4096 & 1 & 18.49763 & 3.295944e-07 & 1.81505090112 & 0.0907525450558 \\
+polus & 20 & 8192 & 1 & 51.10186 & 8.23785e-08 & 1.5511922658 & 0.07755961329 \\
+polus & 40 & 2048 & 1 & 30.35062 & 1.31902e-06 & 0.289864852843 & 0.00724662132108 \\
+polus & 40 & 4096 & 1 & 29.18828 & 3.295944e-07 & 1.15026099517 & 0.0287565248792 \\
+polus & 40 & 8192 & 1 & 50.30223 & 8.23785e-08 & 1.57585081218 & 0.0393962703045 \\
+\hline
+
+\end{tabular}
+\caption{Результаты запусков на Polus}
+\end{table}
+\end{center}
+
 \end{document}

### 2019-12-11 hpc_mod_and_tech/lab2/results_bluegene.py

@@ -1,0 +1,63 @@
+import os
+import sys
+
+def sc(x):
+    return '{:.2e}'.format(x)
+
+if __name__ == "__main__":
+    
+    resFile = sys.argv[1]
+
+    base = dict()
+    with open(resFile, 'r') as file:
+        for line in file:
+            a = line.strip().split()
+
+            pc = a[0]
+            nodes = int(a[1])
+            grid = int(a[2])
+            omp = int(a[4])
+            time = float(a[6])
+            delta = float(a[14]) 
+
+            if nodes == 128 and omp == 1:
+                base[grid] = time
+
+
+
+    with open(resFile, 'r') as file:
+        delim = " & "
+        s = "HPC" + delim 
+        s += "Nodes" + delim
+        s += "Grid" + delim
+        s += "OMP Threads" + delim
+        s += "Wall time" + delim
+        s += "Delta" + delim
+        s += "Speedup" + delim
+        s += "Efficiency" + r" \\"
+        print(s)
+        for line in file:
+            a = line.strip().split()
+            pc = a[0]
+            nodes = int(a[1])
+            grid = int(a[2])
+            omp = int(a[4])
+            time = float(a[6])
+            delta = float(a[14]) 
+            speedup = base[grid] / time
+            expected = (nodes / 128) * omp
+            efficiency = speedup  / expected
+
+            lst = [pc, nodes, grid, omp, time, delta, speedup, efficiency]
+            delim = " & "
+            s = ""
+            s += pc + delim 
+            s += str(nodes) + delim
+            s += str(grid) + delim
+            s += str(omp) + delim
+            s += sc(time) + delim
+            s += sc(delta) + delim
+            s += sc(speedup) + delim
+            s += sc(efficiency) + r" \\"
+
+            print(s)

### 2019-12-11 hpc_mod_and_tech/lab2/results_polus.py

@@ -1,0 +1,62 @@
+import os
+import sys
+
+def sc(x):
+    return str(x) #'{:.2e}'.format(x)
+
+if __name__ == "__main__":
+    
+    resFile = sys.argv[1]
+
+    base = dict()
+    with open(resFile, 'r') as file:
+        for line in file:
+            a = line.strip().split()
+
+            pc = a[0]
+            nodes = int(a[1])
+            grid = int(a[2])
+            omp = int(a[4])
+            time = float(a[6])
+            delta = float(a[14]) 
+
+            if nodes == 1 and omp == 1:
+                base[grid] = time
+
+
+    with open(resFile, 'r') as file:
+        delim = " & "
+        s = "HPC" + delim 
+        s += "Nodes" + delim
+        s += "Grid" + delim
+        s += "OMP Threads" + delim
+        s += "Wall time" + delim
+        s += "Delta" + delim
+        s += "Speedup" + delim
+        s += "Efficiency" + r" \\"
+        print(s)
+        for line in file:
+            a = line.strip().split()
+            pc = a[0]
+            nodes = int(a[1])
+            grid = int(a[2])
+            omp = int(a[4])
+            time = float(a[6])
+            delta = float(a[14]) 
+            speedup = base[grid] / time
+            expected = nodes * omp
+            efficiency = speedup  / expected
+
+            lst = [pc, nodes, grid, omp, time, delta, speedup, efficiency]
+            delim = " & "
+            s = ""
+            s += pc + delim 
+            s += str(nodes) + delim
+            s += str(grid) + delim
+            s += str(omp) + delim
+            s += sc(time) + delim
+            s += sc(delta) + delim
+            s += sc(speedup) + delim
+            s += sc(efficiency) + r" \\"
+
+            print(s)

### 2019-12-11 specsem/2019_12_24/report.tex
Scientific actvity report. Autumn 2019

\documentclass[12pt,a4paper]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}

\usepackage{graphicx}
\usepackage{amsmath}

\begin{document}

\Large

\begin{center}
\textbf{Осенний семестр 2019 года. \\
Краткий отчет по НИР.}
\end{center}

\hfill\begin{minipage}{0.6\textwidth}
    \normalsize
    \textbf{Факультет:} Вычислительной математики и кабернетики \\
    \textbf{Кафедра:} Суперкомпьютеров и квантовой информатики \\
    \textbf{Группа:} М-218 \\
    \textbf{Студент:} Иванов Денис Евгеньевич \\
    \textbf{Научный руководитель:} к.т.н. Семенов Александр Сергеевич \\
\end{minipage}
\vfill

\normalsize
В мае этого года была представлена работа, в которой были описаны две различныe идеи для
поиска черных дыр в графах. Для взвешенного случая предлагалось использовать муравьиный алгоритм,
а для невзвешенного сжатие графа и топологическая сортивка.

В этом семестре алгоритм поиска черных дыр в невзвешенном направленном графе с использованием топологической сортировки получил
свое развитие. Было предложено разделить поиск черных дыр на две стадии: предобработка графа и поиск самих дыр.
На первом этапе мы, как и раньше, используем сжатие графа на основе конденсации компонент сильной связности.
Тем самым, задача упрощается, а также исключаются некоторые "неудобные" для полного перебора случаи.
Особенность второго этапа состоит в том, что здесь комбинаторная природа решаемой задачи проявляется во 
всей красе. Поэтому основной задачей тут мы выбрали сокращение перебора.
Для этого мы используем алгоритм, основанный на топологической структуре графа, который позволяет
значительно ускорить перебор за счет пропуска не уникальных черных дыр.

Описанный выше двухшаговый алгоритм был реализован в однопоточном режиме.
Для демонстрации его сильных и слабых сторон, мы также реализовали алгоритм iBlackhole, который в 2010 году был
предложен исследователями университета New Jersey. Результаты запусков показали, что в ряде проблемных 
случаев наш алгоритм значительно превосходит iBlackhole, однако имеет и свои слабые места за счет более дорогой предобработки графа,
а также накладных расходов, связанных с вычислением необходимых данных для применения эвристики.

Детали работы алгоритма и проведенного исследования изложены в статье поданной на конференцию ПАВТ 2020.

На данный момент алгоритм был реализован не полностью, опущена часть сложных случаев, но даже при этом было
продемонстрировано значительное преимущество в ряде задач. Планируется реализацию закончить и провести испытания
на полноценной версии. Есть предположение, что применимость предложенного алгоритма будет значительно зависеть от размера графа и временных ограничений
в конкретных приложениях. Также планируется реализовать более общий случай данного алгоритма и провести сравнительный анализ применимости
различных версий.

\end{document}

### 2019-12-12 specsem/2019_12_24/report.tex
Отчет подправлен

@@
 \begin{center}
-\textbf{Осенний семестр 2019 года. \\
-Краткий отчет по НИР.}
+\textbf{Алгоритмы поиска подграфов типа <<черная дыра>> в ориентированном графе}\\
+	Осенний семестр 2019 года \\
+Краткий отчет по НИР 
 \end{center}
@@
     \textbf{Студент:} Иванов Денис Евгеньевич \\
-    \textbf{Научный руководитель:} к.т.н. Семенов Александр Сергеевич \\
+    \textbf{Научный руководитель:} к.т.н. А.С. Семенов \\
 \end{minipage}
@@
 \normalsize
-В мае этого года была представлена работа, в которой были описаны две различныe идеи для
-поиска черных дыр в графах. Для взвешенного случая предлагалось использовать муравьиный алгоритм,
-а для невзвешенного сжатие графа и топологическая сортивка.
+В конце 2018/2019 уч. года была представлена работа, в которой были описаны две различныe идеи для
+поиска черных дыр в графах. Для случая взвешенного графа предлагалось использовать муравьиный алгоритм, а для невзвешенного -- сжатие графа и топологическую сортировку.
 
-В этом семестре алгоритм поиска черных дыр в невзвешенном направленном графе с использованием топологической сортировки получил
-свое развитие. Было предложено разделить поиск черных дыр на две стадии: предобработка графа и поиск самих дыр.
+В осеннем семестре 2019/2020 уч.
+года алгоритм поиска черных дыр для невзвешенного ориентированного графа с использованием топологической сортировки получил свое развитие.
+Было предложено разделить поиск черных дыр на две стадии: предобработка графа и поиск самих дыр.
 На первом этапе мы, как и раньше, используем сжатие графа на основе конденсации компонент сильной связности.
 Тем самым, задача упрощается, а также исключаются некоторые "неудобные" для полного перебора случаи.
-Особенность второго этапа состоит в том, что здесь комбинаторная природа решаемой задачи проявляется во 
-всей красе. Поэтому основной задачей тут мы выбрали сокращение перебора.
+Особенность второго этапа состоит в необходимости комбинаторого перебора. Поэтому основной задачей выбрана оптимизация этого перебора.
 Для этого мы используем алгоритм, основанный на топологической структуре графа, который позволяет
-значительно ускорить перебор за счет пропуска не уникальных черных дыр.
+значительно ускорить перебор за счет пропуска неуникальных черных дыр.
 
 Описанный выше двухшаговый алгоритм был реализован в однопоточном режиме.
-Для демонстрации его сильных и слабых сторон, мы также реализовали алгоритм iBlackhole, который в 2010 году был
+Для демонстрации его сильных и слабых сторон мы также реализовали алгоритм iBlackhole, который в 2010 году был
 предложен исследователями университета New Jersey. Результаты запусков показали, что в ряде проблемных 
@@
 
-Детали работы алгоритма и проведенного исследования изложены в статье поданной на конференцию ПАВТ 2020.
-
-На данный момент алгоритм был реализован не полностью, опущена часть сложных случаев, но даже при этом было
-продемонстрировано значительное преимущество в ряде задач. Планируется реализацию закончить и провести испытания
+На данный момент алгоритм был реализован не полностью, опущена часть сложных случаев, но даже при этом продемонстрировано значительное преимущество в ряде задач.
+Планируется реализацию закончить и провести испытания
 на полноценной версии. Есть предположение, что применимость предложенного алгоритма будет значительно зависеть от размера графа и временных ограничений
@@
 
+Детали работы алгоритма и проведенного исследования изложены в полной статье на английском языке, поданной на конференцию ПАВТ 2020 (см. приложение).
+
 \end{document}

### 2019-12-13 hpc_mod_and_tech/lab2/report/report.tex
Final version, I hope

@@
 
-OpenMP был сразу же включен в программный код. В данной работе считаем, что запуск OpenMP в один поток - это запуск без OpenMP.
-За количество потоков отвечает отдельный параметр.
+На Bluegene код выполнялся с использованием OpenMP, в один, два или три потока. 
+Запуск в один OpenMP поток считаем равносильным запуску без OpenMP. Производительность различных конфигураций
+считалась относительно запусков на 128 узлах в один поток. Для сборки кода на bluegene была использована следующая команда:
+
+\begin{center}
+\small
+mpixlcxxi\_r -O2 main.cpp -o bluegene -qsmp=omp
+\end{center}
+
+На Polus запуск с одной OpenMP нитью почти не показывал роста производительности
+при увеличении числа узлов. Однако, если убрать OpenMP директивы из кода программы и скомпилировать без оптимизаций,
+то прирост производительности становится очевиден. Для сборки кода на polus была использована следующая команда:
+
+\begin{center}
+\small
+mpixlC -O0 main.cpp -o polus
+\end{center}
 
@@
 \small
-\textit{mpirun -n X ./solution mpiN mpiM gridN gridM gridT xmax ymax ompNumThreads}
+mpirun -n X ./solution mpiN mpiM gridN gridM gridT xmax ymax ompNumThreads
 \end{center}
@@
     \item xmax, ymax - ограничения по $x, y$
-    \item ompNumThreads - число потоков OpenMP
+    \item ompNumThreads - число потоков OpenMP (отсутствует для Polus)
 \end{itemize}
@@
 
-\section{Экспериментальные результаты}
+\section{Результаты}
 
@@
 \label{tabular:polus}
-
-HPC & Nodes & Grid & OMP & Wall time & Delta & Speedup & Efficiency \\
+HPC & Nodes & Grid & Wall time & Delta & Speedup & Efficiency \\
 \hline
-polus & 1 & 2048 & 1 & 8.797578 & 1.31902e-06 & 1.0 & 1.0 \\
-polus & 1 & 4096 & 1 & 33.57414 & 3.295944e-07 & 1.0 & 1.0 \\
-polus & 1 & 8192 & 1 & 79.26881 & 8.23785e-08 & 1.0 & 1.0 \\
-polus & 10 & 2048 & 1 & 29.14442 & 1.31902e-06 & 0.301861488408 & 0.0301861488408 \\
-polus & 10 & 4096 & 1 & 20.4802 & 3.295944e-07 & 1.63934629545 & 0.163934629545 \\
-polus & 10 & 8192 & 1 & 61.29066 & 8.23785e-08 & 1.29332609569 & 0.129332609569 \\
-polus & 20 & 2048 & 1 & 14.50379 & 1.31902e-06 & 0.606570972139 & 0.030328548607 \\
-polus & 20 & 4096 & 1 & 18.49763 & 3.295944e-07 & 1.81505090112 & 0.0907525450558 \\
-polus & 20 & 8192 & 1 & 51.10186 & 8.23785e-08 & 1.5511922658 & 0.07755961329 \\
-polus & 40 & 2048 & 1 & 30.35062 & 1.31902e-06 & 0.289864852843 & 0.00724662132108 \\
-polus & 40 & 4096 & 1 & 29.18828 & 3.295944e-07 & 1.15026099517 & 0.0287565248792 \\
-polus & 40 & 8192 & 1 & 50.30223 & 8.23785e-08 & 1.57585081218 & 0.0393962703045 \\
+polus & 1 & 2048 & 1.48e+01 & 1.32e-06 & 1.00e+00 & 1.00e+00 \\
+polus & 1 & 4096 & 5.60e+01 & 3.30e-07 & 1.00e+00 & 1.00e+00 \\
+polus & 1 & 8192 & 2.20e+02 & 8.24e-08 & 1.00e+00 & 1.00e+00 \\
+polus & 10 & 2048 & 1.57e+00 & 1.32e-06 & 9.41e+00 & 9.41e-01 \\
+polus & 10 & 4096 & 5.94e+00 & 3.30e-07 & 9.43e+00 & 9.43e-01 \\
+polus & 10 & 8192 & 2.34e+01 & 8.24e-08 & 9.41e+00 & 9.41e-01 \\
+polus & 20 & 2048 & 1.09e+00 & 1.32e-06 & 1.35e+01 & 6.77e-01 \\
+polus & 20 & 4096 & 3.58e+00 & 3.30e-07 & 1.56e+01 & 7.82e-01 \\
+polus & 20 & 8192 & 1.38e+01 & 8.24e-08 & 1.60e+01 & 8.00e-01 \\
+polus & 40 & 2048 & 1 & IN QUEUE \\
+polus & 40 & 4096 & 1 & IN QUEUE \\
+polus & 40 & 8192 & 1 & IN QUEUE \\
 \hline

### 2019-12-13 hpc_mod_and_tech/lab2/main.cpp

@@ -12,6 +12,54 @@
 
 using namespace std;
 
+double u1(double x, double y, double t) {
+    return (t * t  + 1) * sin(2 * M_PI * x) * sin(2 * M_PI * y);
+}
+
+double u1t(double x, double y, double t) {
+    return 2 * t * sin(2 * M_PI * x) * sin(2 * M_PI * y);
+}
+
+double u1xx(double x, double y, double t) {
+    return (t * t  + 1) * sin(2 * M_PI * x) * sin(2 * M_PI * y) * (-4 * M_PI * M_PI);
+}
+
+double u1yy(double x, double y, double t) {
+    return (t * t  + 1) * sin(2 * M_PI * x) * sin(2 * M_PI * y) * (-4 * M_PI * M_PI);
+}
+
+double u1xy(double x, double y, double t) {
+    return (t * t  + 1) * cos(2 * M_PI * x) * cos(2 * M_PI * y) * (4 * M_PI * M_PI);
+}
+        
+double u2(double x, double y, double t) {
+    return (t * t + 1) * (cos(2 * M_PI * x) - 1) * (cos(2 * M_PI * y) - 1);
+}
+
+double u2t(double x, double y, double t) {
+    return 2 * t * (cos(2 * M_PI * x) - 1) * (cos(2 * M_PI * y) - 1);
+}
+
+double u2xx(double x, double y, double t) {
+    return (t * t + 1) * cos(2 * M_PI * x) * (cos(2 * M_PI * y) - 1) * (-4 * M_PI * M_PI);
+}
+
+double u2yy(double x, double y, double t) {
+    return (t * t + 1) * cos(2 * M_PI * y) * (cos(2 * M_PI * x) - 1) * (-4 * M_PI * M_PI);
+}
+
+double u2xy(double x, double y, double t) {
+    return (t * t + 1) * sin(2 * M_PI * x) * sin(2 * M_PI * y) * (4 * M_PI * M_PI);
+}
+
+double f1(double x, double y, double t) {
+    return u1t(x,y,t) - 2 * u1xx(x,y,t) - u1yy(x,y,t) - u2xy(x,y,t);
+}
+
+double f2(double x, double y, double t) {
+    return u2t(x,y,t) - u2xx(x,y,t) - 2 * u2yy(x,y,t) - u1xy(x,y,t);
+}
+
 class Calc {
 private:
     int tl, t, tr;
@@ -19,6 +67,7 @@ private:
     int bl, b, br;
[…]

### 2019-12-13 hpc_mod_and_tech/lab2/results_polus.py

@@ -2,7 +2,7 @@ import os
 import sys
 
 def sc(x):
-    return str(x) #'{:.2e}'.format(x)
+    return '{:.2e}'.format(x)
 
 if __name__ == "__main__":
     
@@ -29,7 +29,6 @@ if __name__ == "__main__":
         s = "HPC" + delim 
         s += "Nodes" + delim
         s += "Grid" + delim
-        s += "OMP Threads" + delim
         s += "Wall time" + delim
         s += "Delta" + delim
         s += "Speedup" + delim
@@ -40,20 +39,18 @@ if __name__ == "__main__":
             pc = a[0]
             nodes = int(a[1])
             grid = int(a[2])
-            omp = int(a[4])
             time = float(a[6])
             delta = float(a[14]) 
             speedup = base[grid] / time
-            expected = nodes * omp
+            expected = nodes
             efficiency = speedup  / expected
 
-            lst = [pc, nodes, grid, omp, time, delta, speedup, efficiency]
+            lst = [pc, nodes, grid, time, delta, speedup, efficiency]
             delim = " & "
             s = ""
             s += pc + delim 
             s += str(nodes) + delim
             s += str(grid) + delim
-            s += str(omp) + delim
             s += sc(time) + delim
             s += sc(delta) + delim
             s += sc(speedup) + delim

### 2020-04-19 blackholes/pavt/results.py
cleanup

@@ -1,0 +1,24 @@
+import os
+
+
+if __name__ == "__main__":
+    for filename in os.listdir("outputs"):
+        path = os.path.join("outputs", filename)
+        vals = [x.strip() for x in filename.split('.')]
+        graph, size, algo = vals
+        res = 0
+        time = str(60 * 20)
+        with open(path, 'r') as file:
+            for line in file:
+                line = line.strip()
+                vals = line.split()
+                if len(vals) == 2:
+                    s, cnt = vals
+                    if s == "bhcount":
+                        res = max(res, int(cnt))
+                else:
+                    s = vals[0]
+                    if s == "It":
+                        time = vals[3]
+        out = graph + " " + str(size) + " " + algo + " " + str(res) + " " + str(time)
+        print(out)

### 2020-04-19 blackholes/pavt/article.tex
Small changes for russian supercomputer days

@@
 %
-\title{Topological Approach to Finding Blackholes in Directed Networks}
+\title{Topological Approach to Finding Blackholes in Large Directed Networks}
 %
@@
 \begin{abstract}
-In this paper we consider the problem of finding so called blackhole pattern in directed unweighted graphs.
+In this paper we consider the problem of finding so called blackhole pattern in relatively large directed unweighted graphs.
 Firstly, we analyse already existing algorithms and approaches. Secondly, we describe how structure of a graph
@@
 Two years later they published an algorithm for approximate blackhole search in general case (weighted graph) \cite{li2012mining,li2014mining}.
+These works do not cover the case of large scale graphs, as the authors only run their algorithm on 1500 node networks, which is quite a small size.
 Hong et al \cite{hong2015detecting} consider application for online blackhole mining in the city environment.

### 2020-04-25 blackholes/chinese/main.cpp
consmetics

@@ -20,7 +20,7 @@ int bhCount = 0;
 int filtered = 0;
 
 int printDebugInfo = 0;
- 
+
 class Solution {
 public:
 
@@ -76,7 +76,6 @@ public:
         }
         return res;
     }
-                
 
     static void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<int>& tsOrder, int bhSize) {
         const int totalVertex = static_cast<int>(tsOrder.size());
@@ -293,7 +292,7 @@ void Print(const string& heading, const set<int>& closure) {
     cout << endl;
 }
 
-int main(int argc, char** argv) { 
+int main(int argc, char** argv) {
     const string inputFileName(argv[1]);
     int maxBHSize = atoi(argv[2]);
     if (argc >= 4) {
@@ -315,7 +314,7 @@ int main(int argc, char** argv) {
         } else {
             maxBHSize = min(maxBHSize, static_cast<int>(graph.size()));
         }
- 
+
         vector<TCandidates> candidates(maxBHSize + 1);
         for (int i = 1; i <= maxBHSize; i++) {
             cerr << "Iter: " << i << endl;
@@ -328,7 +327,7 @@ int main(int argc, char** argv) {
                 }
             }
             Print("P", candidates[i]);
-            
+
             TCandidates toRemove;
             for (int v : candidates[i]) {
                 if (toRemove.find(v) != toRemove.end()) {
@@ -337,12 +336,24 @@ int main(int argc, char** argv) {
                 if (candidates[i - 1].find(v) != candidates[i - 1].end()) {
                     continue;
                 }
+                bool remove = false;
+                for (int s : graph[v]) {
+                    if (candidates[i].find(s) == candidates[i].end() || toRemove.find(s) != toRemove.end()) {
+                        remove = true;
+                        break;
+                    }
+                }
+                if (remove) {
+                    TClosure revClosure = Solution::GetClosure(graphRev, v);
+                    toRemove.insert(revClosure.begin(), revClosure.end());
+                }
+               /* 
                 TClosure closure = Solution::GetClosure(graph, v);
                 Print("closure", closure);
                 if (!Solution::HasCandidates(closure, candidates[i])) {
                     TClosure revClosure = Solution::GetClosure(graphRev, v);
                     toRemove.insert(revClosure.begin(), revClosure.end());
-                } 
+                } */
             }
 
             for (int v : toRemove) {
[…]

### 2020-04-25 blackholes/experiment/results.py
changes

@@ -2,7 +2,6 @@ import os
 
 
 if __name__ == "__main__":
-    
     for filename in os.listdir("outputs"):
         path = os.path.join("outputs", filename)
         vals = [x.strip() for x in filename.split('.')]
@@ -21,5 +20,5 @@ if __name__ == "__main__":
                     s = vals[0]
                     if s == "It":
                         time = vals[3]
-        out = graph + "\t" + str(size) + "\t" + algo + "\t" + str(res) + "\t" + str(time)
-        print(out) 
+        out = graph + " " + str(size) + " " + algo + " " + str(res) + " " + str(time)
+        print(out)

### 2020-04-25 blackholes/experiment/run.py

@@ -5,7 +5,7 @@ class Submission:
     def __init__(self, typeStr1, typeStr2):
         self.typeStr1 = typeStr1
         self.typeStr2 = typeStr2
-        self.command = r"mpisubmit.pl = -g --stdout=" + self.typeStr1 + "." 
+        self.command = r"mpisubmit.pl = -g --stdout=" + self.typeStr1 + "."
         pass
 
 if __name__ == "__main__":
@@ -15,23 +15,21 @@ if __name__ == "__main__":
         for i in range(t):
             typeStr1, typeStr2 = file.readline().strip().split()
             sub.append(Submission(typeStr1, typeStr2))
-    
-    for s in sub:
-        for scale in range(a, b + 1):
-            scale_str = "{:02d}".format(scale) 
+
+    for scale in range(a, b + 1, 2):
+        print("echo scale = " + str(scale))
+        for s in sub:
+            scale_str = "{:02d}".format(scale)
             id_str = s.typeStr1 + "." + scale_str
             outfile = "outputs/" + id_str
             errfile = "errors/" + id_str
-            inputfile = "inputs/" + id_str
-            timeout = "5s"
+            inputfile = "graphs/" + id_str + ".bin"
 
-            cmd1 = "timeout 5s " + "../chinese/local " + inputfile + " -1 &> " + outfile + ".chinese" 
-            cmd2 = "timeout 5s " + "../topsort/local " + inputfile + " 1 &> " + outfile + ".topsort"
+            cmd1 = "timeout 20m " + "../chinese/local " + inputfile + " -1 &> " + outfile + ".chinese"
+            cmd2 = "timeout 20m " + "../topsort/local " + inputfile + " 1 &> " + outfile + ".topsort"
 
             print(cmd1)
-            os.system(cmd1);
             print(cmd2)
-            os.system(cmd2);
 
 #define graph
 #	./generator -s $(1) -directed -weighted -file inputs/$(1).$(2).graph.bin -type $(2)

### 2020-04-25 blackholes/experiment/run.sh

@@ -1,30 +1,70 @@
-timeout 5s ../chinese/local graphs/RMAT.04.bin -1 &> outputs/RMAT.04.chinese
-timeout 5s ../topsort/local graphs/RMAT.04.bin 1 &> outputs/RMAT.04.topsort
-timeout 5s ../chinese/local graphs/RMAT.08.bin -1 &> outputs/RMAT.08.chinese
-timeout 5s ../topsort/local graphs/RMAT.08.bin 1 &> outputs/RMAT.08.topsort
-timeout 5s ../chinese/local graphs/RMAT.12.bin -1 &> outputs/RMAT.12.chinese
-timeout 5s ../topsort/local graphs/RMAT.12.bin 1 &> outputs/RMAT.12.topsort
-timeout 5s ../chinese/local graphs/RMAT.16.bin -1 &> outputs/RMAT.16.chinese
-timeout 5s ../topsort/local graphs/RMAT.16.bin 1 &> outputs/RMAT.16.topsort
-timeout 5s ../chinese/local graphs/RMAT.20.bin -1 &> outputs/RMAT.20.chinese
-timeout 5s ../topsort/local graphs/RMAT.20.bin 1 &> outputs/RMAT.20.topsort
-timeout 5s ../chinese/local graphs/SSCA2.04.bin -1 &> outputs/SSCA2.04.chinese
-timeout 5s ../topsort/local graphs/SSCA2.04.bin 1 &> outputs/SSCA2.04.topsort
-timeout 5s ../chinese/local graphs/SSCA2.08.bin -1 &> outputs/SSCA2.08.chinese
-timeout 5s ../topsort/local graphs/SSCA2.08.bin 1 &> outputs/SSCA2.08.topsort
-timeout 5s ../chinese/local graphs/SSCA2.12.bin -1 &> outputs/SSCA2.12.chinese
-timeout 5s ../topsort/local graphs/SSCA2.12.bin 1 &> outputs/SSCA2.12.topsort
-timeout 5s ../chinese/local graphs/SSCA2.16.bin -1 &> outputs/SSCA2.16.chinese
-timeout 5s ../topsort/local graphs/SSCA2.16.bin 1 &> outputs/SSCA2.16.topsort
-timeout 5s ../chinese/local graphs/SSCA2.20.bin -1 &> outputs/SSCA2.20.chinese
-timeout 5s ../topsort/local graphs/SSCA2.20.bin 1 &> outputs/SSCA2.20.topsort
[…]

### 2020-04-25 blackholes/graph_generator/graph.py
changes

@@ -25,12 +25,12 @@ if __name__ == "__main__":
             inputfile = "inputs/" + id_str
 
             cmd1 = "./generator/generator -s " + scale_str + " -unweighted -file " + graphfile + " -type " + s.typeStr1
-            cmd2 = "./graph_to_file/graph_to_file " + graphfile + " " + inputfile
+            #cmd2 = "./graph_to_file/graph_to_file " + graphfile + " " + inputfile
             print(cmd1)
-            print(cmd2)
+            #print(cmd2)
 
             os.system(cmd1)
-            os.system(cmd2)
+            #os.system(cmd2)
 
 
 #define graph

### 2020-04-25 blackholes/graph_generator/graph.py
changes and gitignore

@@ -4,7 +4,7 @@ class Submission:
     def __init__(self, typeStr1, typeStr2):
         self.typeStr1 = typeStr1
         self.typeStr2 = typeStr2
-        self.command = r"mpisubmit.pl = -g --stdout=" + self.typeStr1 + "." 
+        self.command = r"mpisubmit.pl = -g --stdout=" + self.typeStr1 + "."
         pass
 
 if __name__ == "__main__":
@@ -14,10 +14,9 @@ if __name__ == "__main__":
         for i in range(t):
             typeStr1, typeStr2 = file.readline().strip().split()
             sub.append(Submission(typeStr1, typeStr2))
-    
     for s in sub:
         for scale in range(a, b + 1):
-            scale_str = "{:02d}".format(scale) 
+            scale_str = "{:02d}".format(scale)
             id_str = s.typeStr1 + "." + scale_str
             outfile = "outputs/" + id_str
             errfile = "errors/" + id_str

### 2020-04-25 blackholes/stats/Source.cpp
Graph statistics tool

@@ -1,0 +1,71 @@
+#include <iostream>
+#include <string>
+#include <vector>
+#include <set>
+#include <algorithm>
+#include <cmath>
+#include <iomanip>
+#include <chrono>
+#include <fstream>
+#include <stack>
+#include <queue>
+#include <map>
+
+using namespace std;
+
+using TUsed = vector<char>;
+using TGraph = vector<vector<int>>;
+using TComponents = vector<int>;
+using TSccMembers = vector<vector<int>>;
+using TClosure = set<int>;
+
+int bhCount = 0;
+int filtered = 0;
+int skipped = 0;
+
+class Solution {
+public:
+    static TGraph ReadGraphFromFile(const string& filename) {
+        fstream file(filename, ios::in | ios::binary);
+        int n;
+        long long m;
+        // read header
+        file.read((char*)(&n), sizeof(int));
+        file.read((char*)(&m), sizeof(long long));
+
+        TGraph graph(n);
+
+        vector<set<int>> tmp(n);
+        int e = 0;
+        for(long long i = 0; i < m; i++) {
+            int x, y;
+            file.read((char*)(&x), sizeof(int));
+            file.read((char*)(&y), sizeof(int));
+            if (x == y) {
+                continue;
+            }
+            tmp[x].insert(y);
+        }
+        for(int i = 0; i < tmp.size(); i++) {
+            const auto& t = tmp[i];
+            e += t.size();
+            for (int x : t) {
+                graph[i].push_back(x);
+            }
+        }
+        cout << filename << " " << n << " " << e << endl;
+        return graph;
+    }
+};
+
+int main(int argc, char * argv[]) {
+    if (argc < 2){
+        cout << "use ./main input.txt";
+    }
+
+    const string inputFilePath(argv[1]);
+
+    TGraph graph = Solution::ReadGraphFromFile(inputFilePath);
+
+    return 0;
+}

### 2020-04-25 blackholes/stats/stats.py

@@ -1,0 +1,8 @@
+import os
+
+
+if __name__ == "__main__":
+    folder = "../experiment/graphs"
+    for filename in os.listdir(folder):
+        path = os.path.join(folder, filename)
+        print("./local " + path)

### 2020-04-25 blackholes/stats/stats.sh

@@ -1,0 +1,30 @@
+./local ../experiment/graphs/RMAT.04.bin
+./local ../experiment/graphs/RMAT.06.bin
+./local ../experiment/graphs/RMAT.08.bin
+./local ../experiment/graphs/RMAT.10.bin
+./local ../experiment/graphs/RMAT.12.bin
+./local ../experiment/graphs/RMAT.14.bin
+./local ../experiment/graphs/RMAT.16.bin
+./local ../experiment/graphs/RMAT.18.bin
+./local ../experiment/graphs/RMAT.20.bin
+./local ../experiment/graphs/RMAT.22.bin
+./local ../experiment/graphs/SSCA2.04.bin
+./local ../experiment/graphs/SSCA2.06.bin
+./local ../experiment/graphs/SSCA2.08.bin
+./local ../experiment/graphs/SSCA2.10.bin
+./local ../experiment/graphs/SSCA2.12.bin
+./local ../experiment/graphs/SSCA2.14.bin
+./local ../experiment/graphs/SSCA2.16.bin
+./local ../experiment/graphs/SSCA2.18.bin
+./local ../experiment/graphs/SSCA2.20.bin
+./local ../experiment/graphs/SSCA2.22.bin
+./local ../experiment/graphs/UR.04.bin
+./local ../experiment/graphs/UR.06.bin
+./local ../experiment/graphs/UR.08.bin
+./local ../experiment/graphs/UR.10.bin
+./local ../experiment/graphs/UR.12.bin
+./local ../experiment/graphs/UR.14.bin
+./local ../experiment/graphs/UR.16.bin
+./local ../experiment/graphs/UR.18.bin
+./local ../experiment/graphs/UR.20.bin
+./local ../experiment/graphs/UR.22.bin

### 2020-04-26 blackholes/topsort/Source.cpp
refactor topsort

@@ -1,583 +1,0 @@
-#include <iostream>
-#include <string>
-#include <vector>
-#include <set>
-#include <algorithm>
-#include <cmath>
-#include <iomanip>
-#include <chrono>
-#include <fstream>
-#include <stack>
-#include <queue>
-#include <map>
-
-using namespace std;
-
-using TUsed = vector<char>;
-using TGraph = vector<vector<int>>;
-using TComponents = vector<int>;
-using TSccMembers = vector<vector<int>>;
-using TClosure = set<int>;
-
-int bhCount = 0;
-int filtered = 0;
-int skipped = 0;
-
-class Solution {
-public:
-    static TGraph BuildCondensedGraph(int compCnt, const TComponents& comp, const TGraph& graph) {
-        TGraph condensed(compCnt);
-        for (int v = 0; v < graph.size(); v++) {
-            for (int to : graph[v]) {
-                if(comp[v] == comp[to]) {
-                    continue;
-                }
-                condensed[comp[v]].push_back(comp[to]);
-            }
-        }
-        for (auto& cond : condensed) {
-            sort(cond.begin(), cond.end());
-            auto it = std::unique(cond.begin(), cond.end());
-            cond.resize(std::distance(cond.begin(), it));
-        }
-        return condensed;
-    }
-
-    static TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& graphRev) {
-        TComponents scc(graph.size());
-        vector<int> order = TopSort(graph);
-        vector<char> used(graphRev.size());
-        int cnt = 0;
-        for (int v : order) {
-            if (used[v]) {
-                continue;
-            }
-            markSingleSCC(graphRev, used, scc, v, cnt);
-            cnt++;
-        }
-        return scc;
-    }
-
-    static vector<int> TopSort(const TGraph& g){
-        vector<char> used(g.size());
-        vector<int> order;
-        for (int i = 0; i < g.size(); i++) {
-            if(used[i]) {
-                continue;
-            }
-            dfs(g, used, order, i);
-        }
-        reverse(order.begin(), order.end());
-        return order;
-    }
-
-    static vector<int> TopSort(const TGraph& g, const vector<int>& roots, vector<char>& special) {
-        vector<char> used(g.size());
-        vector<int> order;
-        vector<int> count(g.size());
-        for (int i = 0; i < roots.size(); i++) {
-            int v = roots[i];
-            dfs_recursive(g, used, order, v, count, special);
-        }
-        return order;
-    }
-
[…]

### 2020-04-26 blackholes/topsort/lib/algorithm/algorithm.cpp

@@ -1,0 +1,113 @@
+#include "algorithm.h"
+
+#include <graph.h>
+
+#include <algorithm>
+#include <stack>
+
+using namespace std;
+
+size_t dfs_recursive_with_special(const TGraph& g, vector<char>& used, vector<size_t>& order, size_t v, vector<size_t>& count, vector<char>& special) {
+    used[v] = 1;
+    for (size_t to : g[v]) {
+        if (!used[to]) {
+            count[v] += dfs_recursive_with_special(g, used, order, to, count, special);
+        }
+    }
+    count[v] += 1;
+    order.push_back(v);
+    special[v] = char(order.size() == count[v]);
+    return count[v];
+}
+
+vector<size_t> TopSort(const TGraph& g, const vector<size_t>& roots, vector<char>& special) {
+    vector<char> used(g.size());
+    vector<size_t> order;
+    vector<size_t> count(g.size());
+    for (size_t v : roots) {
+        dfs_recursive_with_special(g, used, order, v, count, special);
+    }
+    return order;
+}
+
+void dfs_non_recursive(const TGraph& g, vector<char>& used, vector<size_t>& order, size_t v) {
+    vector<size_t> pos(g.size());
+    stack<size_t> st;
+    st.push(v);
+    while (!st.empty()) {
+        v = st.top();
+        used[v] = 1;
+        if (pos[v] < g[v].size()) {
+            size_t to = g[v][pos[v]];
+            pos[v]++;
+            if (used[to]) {
+                continue;
+            }
+            st.push(to);
+            continue;
+        }
+        st.pop();
+        order.push_back(v);
+    }
+}
+
+vector<size_t> TopSort(const TGraph& g){
+    vector<char> used(g.size());
+    vector<size_t> order;
+    for (size_t i = 0; i < g.size(); i++) {
+        if(used[i]) {
+            continue;
+        }
+        dfs_non_recursive(g, used, order, i);
+    }
+    reverse(order.begin(), order.end());
+    return order;
+}
+
+TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& graphRev) {
+    auto&& markSingleSCC = [](const TGraph& g, vector<char>& used, TComponents& comp, size_t v, size_t compId) {
+        vector<size_t> pos(g.size());
+        stack<size_t> st;
+        st.push(v);
+        while (!st.empty()){
+            v = st.top();
+            used[v] = 1;
+            if (pos[v] < g[v].size()) {
[…]

### 2020-04-26 blackholes/topsort/lib/algorithm/algorithm.h

@@ -1,0 +1,11 @@
+#pragma once
+
+#include <common.h>
+
+#include <vector>
+#include <set>
+
+std::vector<size_t> TopSort(const TGraph& g, const std::vector<size_t>& roots, std::vector<char>& special);
+std::vector<size_t> TopSort(const TGraph& g);
+TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& graphRev);
+bool CheckConnectivity(const TGraph& graphUndir, const std::set<size_t>& bh);

### 2020-04-26 blackholes/topsort/lib/blackholes.cpp

@@ -1,0 +1,187 @@
+#include "blackholes.h"
+
+#include <algorithm.h>
+#include <graph.h>
+
+#include <iostream>
+
+using namespace std;
+
+void Solve(const TGraph& graphCond) {
+    TGraph graphCondRev = GetReverseGraph(graphCond);
+    TGraph graphCondUndir = GetUndirGraph(graphCond);
+    vector<size_t> roots = GetRoots(graphCondRev);
+    size_t maxid = 0;
+    size_t maxsize = 0;
+    for (size_t i = 0; i < roots.size(); i++) {
+        TClosure closure = GetClosure(graphCond, roots[i]);
+        if (closure.size() > maxsize) {
+            maxid = i;
+            maxsize = closure.size();
+        }
+    }
+    swap(roots[0], roots[maxid]);
+    // Solution::Print("roots", roots);
+    vector<char> special(graphCond.size());
+    vector<size_t> tsOrder = TopSort(graphCond, roots, special);
+//           Solution::Print("topsort", tsOrder);
+    BruteForce(graphCond, graphCondUndir, tsOrder, special);
+}
+
+void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size_t>& tsOrder) {
+    (void)graphUndir;
+    const size_t totalVertex = static_cast<size_t>(tsOrder.size());
+    vector<size_t> pos;
+    pos.reserve(totalVertex);
+    for (size_t sampleSize = 1; sampleSize <= totalVertex; sampleSize++) {
+        //cerr << "BF Sample size: " << sampleSize << endl;
+        pos.resize(sampleSize);
+        for (size_t i = 0; i < pos.size(); ++i) {
+            pos[i] = i;
+        }
+        while (true) {
+            set<size_t> bh = GetBlackHole(graph, tsOrder, pos);
+            FoundBlackHole(bh.size());
+            // cout << "BH count " << bhCount << endl;
+            // Print("BH", bh);
+            if (!BruteNext(pos, totalVertex)) {
+                break;
+            }
+        }
+    }
+}
+
+void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size_t>& tsOrder, const vector<char>& special) {
+    size_t filtered = 0;
+    size_t skipped = 0;
+    const size_t totalVertex = static_cast<size_t>(tsOrder.size());
+    const size_t maxSampleSize = totalVertex;
+    vector<size_t> pos;
[…]

### 2020-04-26 blackholes/topsort/lib/blackholes.h

@@ -1,0 +1,22 @@
+#pragma once
+
+#include <common.h>
+
+#include <cstddef>
+
+#include <vector>
+#include <set>
+#include <string>
+
+void Solve(const TGraph& graphCond);
+
+void BruteForce(const TGraph& graph, const TGraph& graphUndir, const std::vector<size_t>& tsOrder);
+void BruteForce(const TGraph& graph, const TGraph& graphUndir, const std::vector<size_t>& tsOrder, const std::vector<char>& special);
+bool BruteNext(std::vector<size_t>& pos, size_t vertexCount);
+
+void FoundBlackHole(size_t size);
+std::set<size_t> GetBlackHole(const TGraph& graph, const std::vector<size_t>& tsOrder, const std::vector<size_t>& pos);
+
+void Print(const std::string& tag, const std::set<size_t>& v);
+void Print(const std::string& tag, const std::vector<size_t>& v);
+void Print(const std::string& tag, const std::vector<std::vector<size_t>>& vv);

### 2020-04-26 blackholes/topsort/lib/common/common.h

@@ -1,0 +1,10 @@
+#pragma once
+#include <cstddef>
+#include <vector>
+#include <set>
+
+using TUsed = std::vector<char>;
+using TGraph = std::vector<std::vector<size_t>>;
+using TComponents = std::vector<size_t>;
+using TSccMembers = std::vector<std::vector<size_t>>;
+using TClosure = std::set<size_t>;

### 2020-04-26 blackholes/topsort/lib/graph/graph.cpp

@@ -1,0 +1,142 @@
+#include "graph.h"
+
+#include <algorithm>
+#include <map>
+#include <queue>
+
+using namespace std;
+
+TGraph GetReverseGraph(const TGraph& graph) {
+    TGraph graphRev(graph.size());
+    for (size_t v = 0; v < graph.size(); v++) {
+        for (size_t to : graph[v]) {
+            graphRev[to].push_back(v);
+        }
+    }
+    return graphRev;
+}
+
+TGraph GetUndirGraph(const TGraph& graph) {
+    TGraph graphUndir(graph.size());
+    for (size_t v = 0; v < graph.size(); v++) {
+        for (size_t to : graph[v]) {
+            graphUndir[to].push_back(v);
+            graphUndir[v].push_back(to);
+        }
+    }
+    return graphUndir;
+}
+
+TClosure GetClosure(const TGraph& g, size_t v, TUsed& used) {
+    return ClosureBFS(g, used, v);
+}
+
+TClosure GetClosure(const TGraph& g, size_t v) {
+    TUsed used;
+    used.resize(g.size());
+    return ClosureBFS(g, used, v);
+}
+
+TClosure ClosureBFS(const TGraph& g, TUsed& used, size_t v) {
+    queue<size_t> q;
+    TClosure closure;
+    if (!used[v]) {
+        q.push(v);
+        closure.insert(v);
+    }
+    used[v] += 1;
+    while(!q.empty()) {
+        v = q.front();
+        q.pop();
+        for (size_t to : g[v]) {
+            if (!used[to]) {
+                used[to] += 1;
+                q.push(to);
+                closure.insert(to);
+            } else {
+                used[to] += 1;
+            }
+        }
+    }
+    return closure;
+}
+
+TGraph CarveComponent(const TGraph& graph, const TClosure& component) {
+    TGraph weak;
+    weak.resize(component.size());
+    map<size_t, size_t> mapping;
+    size_t cnt = 0;
+    for (size_t v : component) {
+        mapping[v] = cnt;
+        cnt++;
+    }
+
+    for (const auto& p : mapping) {
+        size_t old = p.first;
+        size_t neww = p.second;
+        for (size_t v : graph[old]) {
+            if (component.find(v) != component.end()) {
+                weak[neww].push_back(mapping[v]);
+            }
+        }
+    }
+    return weak;
+}
+
[…]

### 2020-04-26 blackholes/topsort/lib/graph/graph.h

@@ -1,0 +1,15 @@
+#pragma once
+#include <common.h>
+
+#include <vector>
+
+TGraph GetReverseGraph(const TGraph& graph);
+TGraph GetUndirGraph(const TGraph& graph);
+TClosure GetClosure(const TGraph& g, size_t v, TUsed& used);
+TClosure GetClosure(const TGraph& g, size_t v);
+TClosure ClosureBFS(const TGraph& g, TUsed& used, size_t v);
+std::vector<TGraph> GetWeakComponents(const TGraph& graph, const TGraph& graphUndir);
+std::vector<size_t> GetRoots(const TGraph& graphRev);
+TGraph BuildCondensedGraph(size_t compCnt, const TComponents& comp, const TGraph& graph);
+size_t GetCompCnt(const TComponents& comp);
+TSccMembers GetSccMembers(const TComponents& comp);

### 2020-04-26 blackholes/topsort/lib/io/io.cpp

@@ -1,0 +1,59 @@
+#include "io.h"
+
+using namespace std;
+
+TGraph ReadGraphFromBinaryStream(std::istream& file) {
+    int n;
+    long long m;
+    file.read((char*)(&n), sizeof(int));
+    file.read((char*)(&m), sizeof(long long));
+    TGraph graph(n);
+    vector<set<int>> tmp(n);
+    for(long long i = 0; i < m; i++) {
+        int x, y;
+        file.read((char*)(&x), sizeof(int));
+        file.read((char*)(&y), sizeof(int));
+        if (x == y) { // omit self loops
+            continue;
+        }
+        tmp[x].insert(y);
+    }
+    for(size_t i = 0; i < tmp.size(); i++) {
+        const auto& t = tmp[i];
+        for (int x : t) {
+            graph[i].push_back(static_cast<size_t>(x));
+        }
+    }
+    return graph;
+}
+
+TGraph ReadGraphFromBinaryFile(const std::string& filename) {
+    ifstream file(filename, ios::in | ios::binary);
+    return ReadGraphFromBinaryStream(file);
+}
+
+TGraph ReadGraphFromTextStream(std::istream& stream) {
+    int n = 0;
+    long long m = 0;
+    stream >> n >> m;
+    TGraph graph(n);
+
+    vector<set<int>> tmp(n);
+    for(int i = 0; i < m; i++) {
+        int x, y;
+        stream >> x >> y;
+        tmp[x].insert(y);
+    }
+    for(size_t i = 0; i < tmp.size(); i++) {
+        const auto& t = tmp[i];
+        for (int x : t) {
+            graph[i].push_back(static_cast<size_t>(x));
+        }
+    }
+    return graph;
+}
+
+TGraph ReadGraphFromTextFile(const string& filename) {
+    ifstream stream(filename);
+    return ReadGraphFromTextStream(stream);
+}

### 2020-04-26 blackholes/topsort/lib/io/io.h

@@ -1,0 +1,13 @@
+#pragma once
+
+#include <common.h>
+
+#include <iostream>
+#include <fstream>
+#include <string>
+
+TGraph ReadGraphFromBinaryStream(std::istream& file);
+TGraph ReadGraphFromBinaryFile(const std::string& filename);
+
+TGraph ReadGraphFromTextStream(std::istream& stream);
+TGraph ReadGraphFromTextFile(const std::string& filename);

### 2020-04-26 blackholes/topsort/main.cpp

@@ -1,0 +1,79 @@
+#include <algorithm.h>
+#include <common.h>
+#include <blackholes.h>
+#include <graph.h>
+#include <io.h>
+
+#include <iostream>
+#include <string>
+//#include <vector>
+//#include <set>
+//#include <algorithm>
+//#include <cmath>
+//#include <iomanip>
+#include <chrono>
+//#include <fstream>
+//#include <stack>
+//#include <queue>
+//#include <map>
+
+using namespace std;
+
+int main(int argc, char * argv[]) {
+    if (argc < 3){
+        cout << "use ./main input.txt optimize";
+    }
+
+    const string inputFilePath(argv[1]);
+    const int optimize(atoi(argv[2]));
+
+    // #ifdef _DEBUG
+        //ofstream fout(outputFilePath, std::ifstream::out);
+    // #endif
+
+    using namespace std::chrono;
+    high_resolution_clock::time_point t1 = high_resolution_clock::now();
+    {
+        TGraph graph = ReadGraphFromBinaryFile(inputFilePath);
+        TGraph graphRev = GetReverseGraph(graph);
+
+        TComponents comp2Vertex = GetStrongConnectivityComponents(graph, graphRev);
+        TSccMembers compMembers = GetSccMembers(comp2Vertex);
+        // Print("components", compMembers);
+        //PrintStat("SCC component sizes", compMembers);
+        size_t compCnt = GetCompCnt(comp2Vertex);
+        TGraph graphCond = BuildCondensedGraph(compCnt, comp2Vertex, graph);
+        if (optimize) {
+            TGraph graphCondUndir = GetUndirGraph(graphCond);
+            vector<TGraph> weaks = GetWeakComponents(graphCond, graphCondUndir);
+            //PrintStat("Weakly connected component sizes", weaks);
+            //if (false)
+            for (const TGraph& weak : weaks) {
+                Solve(weak);
+            }
+        } else {
+            TGraph graphCondUndir = GetUndirGraph(graphCond);
+            vector<size_t> tsOrder = TopSort(graphCond);
+  //          Print("topsort", tsOrder);
+            BruteForce(graphCond, graphCondUndir, tsOrder);
+        }
+/*
+        int total = 0;
+        fout << inputFilePath << endl;
+        fout << "Nodes " << n << "Edges " << m << endl;
+        fout << "BlackHoles cnt: " << endl;
+        for( int i = 0 ; i < blackHoles.size(); i++ ){
[…]

### 2020-04-26 blackholes/blackholes/lib/blackholes.cpp
chinese refactor

@@ -121,7 +121,7 @@ bool BruteNext(vector<size_t>& pos, size_t vertexCount) {
             continue;
         } else {
             cur++;
-            for (size_t j = i; j < sampleSize; j++) {
+            for (size_t j = i; j < sampleSize + 1; j++) {
                 pos[j] = pos[j - 1] + 1;
             }
             break;
@@ -137,7 +137,8 @@ void FoundBlackHole(size_t size) {
     cout << "bh_count" << bhCount << endl;
 }
 
-set<size_t> GetBlackHole(const TGraph& graph, const vector<size_t>& tsOrder, const vector<size_t>& pos) {
+set<size_t> GetBlackHole(const TGraph& graph,
+                         const vector<size_t>& tsOrder, const vector<size_t>& pos) {
     set<size_t> blackhole;
     TUsed used;
     used.assign(graph.size(), 0);

### 2020-04-26 blackholes/chinese/main.cpp
remove old chinese

@@ -1,289 +1,0 @@
-#include <iostream>
-//#include <fstream>
-//#include <string>
-//#include <vector>
-//#include <set>
-//#include <queue>
-//#include <stack>
-//#include <algorithm>
-#include <chrono>
-
-#include <algorithm.h>
-#include <blackholes.h>
-#include <common.h>
-#include <graph.h>
-#include <io.h>
-
-
-using namespace std;
-
-//using TGraph = vector<vector<int>>;
-//using TClosure = set<int>;
-using TCandidates = set<int>;
-//using TUsed = vector<char>;
-
-//int bhCount = 0;
-int printDebugInfo = 0;
-
-void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size_t>& tsOrder, size_t bhSize) {
-    size_t filtered = 0;
-    const size_t totalVertex = static_cast<size_t>(tsOrder.size());
-    vector<size_t> pos;
-    pos.reserve(totalVertex);
-    for (size_t sampleSize = 1; sampleSize <= totalVertex; sampleSize++) {
-        //cerr << "BF Sample size: " << sampleSize << endl;
-        pos.resize(sampleSize);
-        for (size_t i = 0; i < pos.size(); ++i) {
-            pos[i] = i;
-        }
-        while (true) {
-            set<size_t> bh = GetBlackHole(graph, tsOrder, pos);
-            if (bh.size() == 0) {
-                filtered++;
-            }
-            if (bh.size() == bhSize && CheckConnectivity(graphUndir, bh)) {
-                FoundBlackHole(bh.size());
-            }
-            if (!BruteNext(pos, totalVertex)) {
-                break;
-            }
-        }
-        cout << "filtered " << filtered << endl;
-    }
-}
-/*
-    static void Closure(const TGraph& g, int v, TUsed& used) {
-        ClosureSize(g, v, used);
-        return;
-    }
-
-    static void Closure(const TGraph& g, int v) {
-        ClosureSize(g,v);
-        return;
-    } */
-/*
-    static int ClosureSize(const TGraph& g, int v, TUsed& used) {
-        return BFS(g, used, v);
-    } */
-/*
-    static int ClosureSize(const TGraph& g, int v) {
-        TUsed used;
-        used.resize(g.size());
-        return BFS(g, used, v);
-    } */
-
-int DegreeOut(const TClosure& closure, const TGraph& gd) {
-    long long dout = 0;
-    for (int v : closure) {
-        for (int to : gd[v]) {
[…]

### 2020-04-28 blackholes/blackholes/chinese/main.cpp
Removed commentaries, bugfixes and other

@@ -1,11 +1,5 @@
 #include <iostream>
-//#include <fstream>
-//#include <string>
-//#include <vector>
-//#include <set>
-//#include <queue>
-//#include <stack>
-//#include <algorithm>
+#include <queue>
 #include <chrono>
 
 #include <algorithm.h>
@@ -17,13 +11,29 @@
 
 using namespace std;
 
-//using TGraph = vector<vector<int>>;
-//using TClosure = set<int>;
-using TCandidates = set<int>;
-//using TUsed = vector<char>;
+using TCandidates = set<size_t>;
 
-//int bhCount = 0;
-int printDebugInfo = 0;
+bool printDebugInfo = false;
+
+set<size_t> GetBlackHoleChinese(const TGraph& graph, const vector<size_t>& tsOrder, const vector<size_t>& pos) {
+    set<size_t> blackhole;
+    TUsed used;
+    used.assign(graph.size(), 0);
+    for (size_t p : pos) {
+        size_t v = tsOrder[p];
+        TClosure closure = GetClosure(graph, v, used);
+        for (size_t x : closure) {
+            blackhole.insert(x);
+        }
+    }
+    for (size_t p : pos) {
+        size_t x = tsOrder[p];
+        if (used[x] >= 2) {
+            return {}; // means this blackhole has already been detected;
+        }
+    }
+    return blackhole;
+}
 
 void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size_t>& tsOrder, size_t bhSize) {
     size_t filtered = 0;
@@ -31,57 +41,26 @@ void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size
     vector<size_t> pos;
     pos.reserve(totalVertex);
     for (size_t sampleSize = 1; sampleSize <= totalVertex; sampleSize++) {
-        //cerr << "BF Sample size: " << sampleSize << endl;
         pos.resize(sampleSize);
         for (size_t i = 0; i < pos.size(); ++i) {
             pos[i] = i;
         }
         while (true) {
-            set<size_t> bh = GetBlackHole(graph, tsOrder, pos);
+            set<size_t> bh = GetBlackHoleChinese(graph, tsOrder, pos);
             if (bh.size() == 0) {
                 filtered++;
             }
             if (bh.size() == bhSize && CheckConnectivity(graphUndir, bh)) {
-                FoundBlackHole(bh.size());
+                FoundBlackHole(bh, printDebugInfo);
             }
[…]

### 2020-04-28 blackholes/blackholes/lib/algorithm/algorithm.cpp

@@ -64,7 +64,7 @@ vector<size_t> TopSort(const TGraph& g){
     return order;
 }
 
-TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& graphRev) {
+TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& graphRev, size_t& componentCount) {
     auto&& markSingleSCC = [](const TGraph& g, vector<char>& used, TComponents& comp, size_t v, size_t compId) {
         vector<size_t> pos(g.size());
         stack<size_t> st;
@@ -89,13 +89,13 @@ TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& g
     TComponents scc(graph.size());
     vector<size_t> order = TopSort(graph);
     vector<char> used(graphRev.size());
-    size_t cnt = 0;
+    componentCount = 0;
     for (size_t v : order) {
         if (used[v]) {
             continue;
         }
-        markSingleSCC(graphRev, used, scc, v, cnt);
-        cnt++;
+        markSingleSCC(graphRev, used, scc, v, componentCount);
+        componentCount++;
     }
     return scc;
 }

### 2020-04-28 blackholes/blackholes/lib/algorithm/algorithm.h

@@ -7,5 +7,5 @@
 
 std::vector<size_t> TopSort(const TGraph& g, const std::vector<size_t>& roots, std::vector<char>& special);
 std::vector<size_t> TopSort(const TGraph& g);
-TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& graphRev);
+TComponents GetStrongConnectivityComponents(const TGraph& graph, const TGraph& graphRev, size_t& componentCount);
 bool CheckConnectivity(const TGraph& graphUndir, const std::set<size_t>& bh);

### 2020-04-28 blackholes/blackholes/lib/blackholes.cpp

@@ -7,25 +7,25 @@
 
 using namespace std;
 
-void Solve(const TGraph& graphCond) {
-    TGraph graphCondRev = GetReverseGraph(graphCond);
-    TGraph graphCondUndir = GetUndirGraph(graphCond);
-    vector<size_t> roots = GetRoots(graphCondRev);
+void TopsortSolver(const TGraph& graph, bool printDebugInfo) {
+    TGraph graphRev = GetReverseGraph(graph);
+    TGraph graphUndir = GetUndirGraph(graph);
+    vector<size_t> roots = GetRoots(graphRev);
     size_t maxid = 0;
     size_t maxsize = 0;
     for (size_t i = 0; i < roots.size(); i++) {
-        TClosure closure = GetClosure(graphCond, roots[i]);
+        TClosure closure = GetClosure(graph, roots[i]);
         if (closure.size() > maxsize) {
             maxid = i;
             maxsize = closure.size();
         }
     }
     swap(roots[0], roots[maxid]);
-    // Solution::Print("roots", roots);
-    vector<char> special(graphCond.size());
-    vector<size_t> tsOrder = TopSort(graphCond, roots, special);
-//           Solution::Print("topsort", tsOrder);
-    BruteForce(graphCond, graphCondUndir, tsOrder, special);
+    Print("roots", roots, printDebugInfo);
+    vector<char> special(graph.size());
+    vector<size_t> tsOrder = TopSort(graph, roots, special);
+    Print("topsort", tsOrder, printDebugInfo);
+    BruteForce(graph, graphUndir, tsOrder, special);
 }
 
 void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size_t>& tsOrder) {
@@ -41,7 +41,7 @@ void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size
         }
         while (true) {
             set<size_t> bh = GetBlackHole(graph, tsOrder, pos);
-            FoundBlackHole(bh.size());
+            FoundBlackHole(bh);
             // cout << "BH count " << bhCount << endl;
             // Print("BH", bh);
             if (!BruteNext(pos, totalVertex)) {
@@ -51,7 +51,7 @@ void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size
     }
 }
 
-void BruteForce(const TGraph& graph, const TGraph& graphUndir, const vector<size_t>& tsOrder, const vector<char>& special) {
[…]

### 2020-04-28 blackholes/blackholes/lib/blackholes.h

@@ -1,5 +1,4 @@
-#pragma once
-
+#pragma once 
 #include <common.h>
 
 #include <cstddef>
@@ -8,15 +7,15 @@
 #include <set>
 #include <string>
 
-void Solve(const TGraph& graphCond);
+void TopsortSolver(const TGraph& graphCond, bool printDebugInfo = 0);
 
 void BruteForce(const TGraph& graph, const TGraph& graphUndir, const std::vector<size_t>& tsOrder);
-void BruteForce(const TGraph& graph, const TGraph& graphUndir, const std::vector<size_t>& tsOrder, const std::vector<char>& special);
+void BruteForce(const TGraph& graph, const TGraph& graphUndir, const std::vector<size_t>& tsOrder, const std::vector<char>& special, bool printDebugInfo = false);
 bool BruteNext(std::vector<size_t>& pos, size_t vertexCount);
 
-void FoundBlackHole(size_t size);
+void FoundBlackHole(const std::set<size_t>& bh, bool printDebugInfo = false);
 std::set<size_t> GetBlackHole(const TGraph& graph, const std::vector<size_t>& tsOrder, const std::vector<size_t>& pos);
 
-void Print(const std::string& tag, const std::set<size_t>& v);
-void Print(const std::string& tag, const std::vector<size_t>& v);
-void Print(const std::string& tag, const std::vector<std::vector<size_t>>& vv);
+void Print(const std::string& tag, const std::set<size_t>& v, bool printDebugInfo);
+void Print(const std::string& tag, const std::vector<size_t>& v, bool printDebugInfo);
+void Print(const std::string& tag, const std::vector<std::vector<size_t>>& vv, bool printDebugInfo);

### 2020-04-28 blackholes/blackholes/lib/graph/graph.cpp

@@ -36,17 +36,18 @@ TClosure GetClosure(const TGraph& g, size_t v) {
     used.resize(g.size());
     return ClosureBFS(g, used, v);
 }
-
-TClosure ClosureBFS(const TGraph& g, TUsed& used, size_t v) {
+/* // old version. only good for shrinked graph.
+    // Do not create localUsed
+TClosure ClosureBFS(const TGraph& g, TUsed& used, size_t s) {
     queue<size_t> q;
     TClosure closure;
-    if (!used[v]) {
-        q.push(v);
-        closure.insert(v);
+    if (!used[s]) {
+        q.push(s);
+        closure.insert(s);
     }
-    used[v] += 1;
+    used[s] += 1;
     while(!q.empty()) {
-        v = q.front();
+        size_t v = q.front();
         q.pop();
         for (size_t to : g[v]) {
             if (!used[to]) {
@@ -54,7 +55,34 @@ TClosure ClosureBFS(const TGraph& g, TUsed& used, size_t v) {
                 q.push(to);
                 closure.insert(to);
             } else {
+                used[to] += 1; // size_t(to != s);
+            }
+        }
+    }
+    return closure;
+} */
+
+TClosure ClosureBFS(const TGraph& g, TUsed& used, size_t s) {
+    TUsed localUsed(used.size(), 0);
+    queue<size_t> q;
+    TClosure closure;
+    if (!used[s]) {
+        q.push(s);
+        closure.insert(s);
+    }
+    used[s] += 1;
+    localUsed[s] += 1;
+    while(!q.empty()) {
+        size_t v = q.front();
+        q.pop();
+        for (size_t to : g[v]) {
+            if (!localUsed[to]) {
+                if (!used[to]) {
+                    q.push(to);
+                    closure.insert(to);
+                }
                 used[to] += 1;
+                localUsed[to] += 1;
             }
         }
     }

### 2020-04-28 blackholes/blackholes/topsort/main.cpp

@@ -6,73 +6,46 @@
 
 #include <iostream>
 #include <string>
-//#include <vector>
-//#include <set>
-//#include <algorithm>
-//#include <cmath>
-//#include <iomanip>
 #include <chrono>
-//#include <fstream>
-//#include <stack>
-//#include <queue>
-//#include <map>
 
 using namespace std;
 
 int main(int argc, char * argv[]) {
-    if (argc < 3){
-        cout << "use ./main input.txt optimize";
+    if (argc < 4){
+        cout << "use ./topsort_cli input.txt binaryInput divideAndConquer" << endl;
+        return 0;
     }
 
     const string inputFilePath(argv[1]);
-    const int optimize(atoi(argv[2]));
-
-    // #ifdef _DEBUG
-        //ofstream fout(outputFilePath, std::ifstream::out);
-    // #endif
+    const int binaryInput(atoi(argv[2]));
+    const int divideAndConquer(atoi(argv[3]));
 
     using namespace std::chrono;
     high_resolution_clock::time_point t1 = high_resolution_clock::now();
     {
-        TGraph graph = ReadGraphFromBinaryFile(inputFilePath);
+        TGraph graph;
+        if (binaryInput) {
+            graph = ReadGraphFromBinaryFile(inputFilePath);
+        } else {
+            graph = ReadGraphFromTextFile(inputFilePath);
+        }
         TGraph graphRev = GetReverseGraph(graph);
 
-        TComponents comp2Vertex = GetStrongConnectivityComponents(graph, graphRev);
-        TSccMembers compMembers = GetSccMembers(comp2Vertex);
-        // Print("components", compMembers);
-        //PrintStat("SCC component sizes", compMembers);
-        size_t compCnt = GetCompCnt(comp2Vertex);
+        size_t compCnt = 0;
+        TComponents comp2Vertex = GetStrongConnectivityComponents(graph, graphRev, compCnt);
         TGraph graphCond = BuildCondensedGraph(compCnt, comp2Vertex, graph);
-        if (optimize) {
+
+        if (divideAndConquer) {
             TGraph graphCondUndir = GetUndirGraph(graphCond);
             vector<TGraph> weaks = GetWeakComponents(graphCond, graphCondUndir);
-            //PrintStat("Weakly connected component sizes", weaks);
-            //if (false)
             for (const TGraph& weak : weaks) {
-                Solve(weak);
+                TopsortSolver(weak);
             }
         } else {
[…]
